---
title: "SDMs for CPO Sanctuaries"
author: "Owen Liu"
date: '2022-09-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cowplot) # for organizing multipanel plots
library(visreg) # for visualizing conditional environmental relationships
library(tictoc) # for timing
```

# Introduction

Running SDMs using `sdmTMB` for California Current elasmobranchs and flatfish, following Janelle Layton's cross-validation [model selection](https://docs.google.com/spreadsheets/d/1gTmQwU91oO_JewZJpgWHUXUdn9pyqxpm/edit#gid=510573475)

First, we load survey and environmental data used for fitting and projection.

# Load Data
```{r}
rmarkdown::render(here::here('scripts','sdmTMB data construction.Rmd'),quiet=TRUE)
```

Next, we outline the functions we will use to fit, project, and visualize model outputs.

# Required Functions

## Prepare Species Data

```{r}
prepare_species <- function(dat,spp){
  dat_sub <- dat %>% 
    filter(species==spp) %>% 
    
    # rescale depth, oxygen, and temp to be N(0,1)
    mutate(across(c(depth_trawl,mean_temp_roms_30,mean_oxygen_roms_30),~(scale(.) %>% as.vector()),.names="{.col}_norm")) %>% 
    
    # add a year indicator
    mutate(year=lubridate::year(date))
}
```

## Run SDM

This is a wrapper around `sdmTMB::sdmTMB()` that includes data construction, and allows for options on whether to include various covariates (depth, substrate) and how to model them in the mixed effects framework (e.g., GAM splines vs. quadratics)

```{r}
run_sdmTMB <- function(dat,spp,use_depth=F,time_vary=F,spatial_field=F,include_substrate=T,hab_spline=T,env_spline=T,spline_k=3){
  # filter data for species
  modeldat <- prepare_species(dat,spp=spp)
  
  # make spde
  spde <- make_mesh(modeldat,xy_cols = c('longitude','latitude'), 
                   cutoff = 20)
  
  # model formula
  formula <- paste0("cpue_kg_km2 ~ ")
  
  # substrate relationship
  substrate <- paste("prop_hard_mixed + I(prop_hard_mixed^2)")
  #wiggly habitat relationship?
  substrate <- ifelse(hab_spline, paste0("s(prop_hard_mixed,k=",spline_k,")"),
                      substrate)
  if(!include_substrate) substrate=""
  
  # make the environmental effects
  enviro <- paste("mean_temp_roms_30_norm + 
                  I(mean_temp_roms_30_norm^2) + 
                  mean_oxygen_roms_30_norm + 
                  I(mean_oxygen_roms_30_norm^2)")
  # wiggly environmental relationships?
  enviro <- ifelse(env_spline, paste0("s(mean_temp_roms_30_norm,k=",spline_k,") + ",
                                      "s(mean_oxygen_roms_30_norm,k=",spline_k,")"),
                   enviro)
  # if depth effect, add to model formula
  if(use_depth) {
    formula = paste0(formula, " + depth_trawl_norm + I(depth_trawl_norm^2)")
  }
  
  time_formula = "~ -1"
  if(time_vary) {
    time_formula = paste0(time_formula, " + ", substrate, " + ", enviro)
    time_varying = as.formula(time_formula)
    time = "year"
  } else {
    formula = paste0(formula, " + ", substrate, " + ", enviro)
    time_varying = NULL
    time = "year"
  }
  print('running model.')
  m <- try( sdmTMB(
    formula = as.formula(formula),
    time_varying = time_varying,
    mesh = spde,
    time = time,
    family = tweedie(link = "log"),
    data = modeldat,
    anisotropy = FALSE,
    spatial = spatial_field,
    #extra_time argument necessary for prediction
    extra_time=1980:2100),
  silent=F)

  if(class(m)=="try-error"){
    print(paste("Error."))
  }else{
    print(paste("Model for",spp,"complete."))
  }

  return(m)

}
```

```{r}
# m <- run_sdmTMB(dat=trawl_roms_utm,spp="longnose skate",use_depth=T,include_substrate=T) # all other options default
```

## Diagnostic Plots

QQ plot, spatiotemporal predictions, spatiotemporal residuals

```{r}
make_diagnostic_plots <- function(m){
  spp=unique(m$data$species)
  sppdat <- prepare_species(trawl_roms_utm,spp)
  resids <- residuals(m)[1:nrow(sppdat),1] %>% as_tibble()
  sppdat$resid=resids$value
  
  # qq plot
  qq <- ggplot(resids,aes(sample=value))+stat_qq()+stat_qq_line()+
    labs(x="Theoretical",y="Sample")

  # spatial and spatiotemporal predictions
  preds <- predict(m) %>% filter(year %in% unique(sppdat$year)) %>% pull('est')
  sppdat$pred <- preds
  preds_time_space <- pred_space(sppdat,X='longitude',Y='latitude',time='year') + coord_equal()
  preds_space <- pred_space(sppdat,X='longitude',Y='latitude',time='year',by_time = F)
  
  # spatiotemporal residuals
  resids_space <- resid_space(sppdat,X='longitude',Y='latitude',time='year') + coord_equal()

  # make and return the plot
  diagplots <- plot_grid(preds_time_space,resids_space,preds_space,qq,nrow=2,rel_heights = c(1,0.6))
  
  diagplots
}
```

```{r,fig.height=8,fig.width=8}
# make_diagnostic_plots(m)
```

## Make Projections

Choose a gcm (Hadley=`hadl`, GFDL=`gfdl`,IPSL=`ipsl`)

```{r}
make_predictions <- function(modelobj,gcm="hadl"){

  original_model_data <- modelobj$data
  # need to use the original (hindcast) environmental data to scale the projected data
  mean_t <- mean(original_model_data$mean_temp_roms_30,na.rm=T)
  sd_t <- sd(original_model_data$mean_temp_roms_30,na.rm=T)
  mean_oxy <- mean(original_model_data$mean_oxygen_roms_30,na.rm=T)
  sd_oxy <- sd(original_model_data$mean_oxygen_roms_30,na.rm=T)
  mean_depth <- mean(original_model_data$depth_trawl,na.rm=T)
  sd_depth <- sd(original_model_data$depth_trawl,na.rm=T)
  
  # create a new tibble with the projected data for the chosen gcm
  newdata <- roms %>% 
    ## NOTE: WE LOSE A TON OF DATA HERE, SOMETHING MAY BE WRONG WITH THE HABITAT DATA
    left_join(hab,by=c("lat","lon")) %>% 
    drop_na() %>% 
    mutate(depth_m= -depth_m) %>%
    dplyr::select(year,lat,lon,latitude,longitude,prop_hard_mixed,depth_m,contains(paste0("30d_",gcm)))
  
  temperature <- newdata %>% dplyr::select(contains('bt')) %>% 
    set_names('temperature') %>% 
    mutate(mean_temp_roms_30_norm=(temperature-mean_t)/sd_t)
  
  oxygen <- newdata %>% dplyr::select(contains('oxy')) %>% 
    set_names('oxygen') %>% 
    mutate(mean_oxygen_roms_30_norm=(oxygen-mean_oxy)/sd_oxy)
  
  depth_new <- newdata %>% dplyr::select(contains('depth')) %>% 
    set_names('depth') %>% 
    mutate(depth_trawl_norm=(depth-mean_depth)/sd_depth)

  newdata <- newdata %>% 
    bind_cols(temperature) %>% 
    bind_cols(oxygen) %>% 
    bind_cols(depth_new) %>% 
    dplyr::select(-temperature,-oxygen, -depth) %>%
    mutate(year=as.double(year))
  
  # years to predict
  yrs <- sort(unique(newdata$year))
  # now we can make the predictions
  predicted_cpue_km2 <- predict(modelobj,newdata,return_tmb_object=F,extra_time=yrs)
  
  predicted_cpue_km2
}
```

## Make Projection Maps

```{r}
# pretty grid for raster maps
gr <- projection_extent %>% st_make_grid(cellsize=10,what='centers') %>% st_as_sf() %>% 
  st_intersection(projection_extent)
gr_xy <- st_coordinates(gr)
```

Map a specific year from the projection

```{r}
map_year <- function(spp_preds,yr_vec=c(2000),return_pred_df=F,plot_leg=T){
  # scale for the legend, common within species
  scl <- c(0,quantile(exp(spp_preds$est),0.99))
  df <- spp_preds %>% 
    filter(year%in%yr_vec) %>% 
    group_by(longitude,latitude) %>% 
    summarise(est=mean(est,na.rm=T) %>% exp()) %>% 
    # rescale super large positive outliers for mapping purposes
    mutate(est=ifelse(est>scl[2],scl[2],est))
  # match nearest neighbors from predictions to grid
  pred_points <- df %>% dplyr::select(longitude,latitude) %>% as.matrix()
  nns <- nn2(pred_points,gr_xy,k=1)$nn.idx
  gr_pred <- gr_xy %>% as_tibble() %>% mutate(est=df$est[nns])
  
  bbox=st_bbox(projection_extent)
  
  if(return_pred_df) {out <- gr_pred %>% as_tibble()}
  
  else{
    out<-ggplot(coast)+
      geom_sf()+
      # geom_point(data=df,aes(longitude,latitude,col=exp(est)))+
      # scale_color_viridis_c()+
      geom_raster(data=gr_pred,aes(x=X,y=Y,fill=est),interpolate=F)+
      scale_fill_viridis_c(limits=scl)+
      coord_sf(datum=NA)+
      xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
      labs(x="",y="",fill="CPUE",title='')
    if(!plot_leg) out <- out + theme(legend.position = 'None')
  }
  out
}
```

Wrapper to make a map for "baseline" (1985-2010), "mid century" (2040-2060), and "end century" (2080-2100)

```{r}
make_three_maps <- function(spp_preds){
  p1 <- map_year(spp_preds,yr_vec=c(1985:2010))+labs(title='baseline')
  p2 <- map_year(spp_preds,yr_vec=c(2040:2060))+labs(title='mid century')
  p3 <- map_year(spp_preds,yr_vec=c(2080:2100))+labs(title='end of century')
  plot_grid(p1,p2,p3,nrow=1)
}
```

```{r,fig.height=5,fig.width=8}
# mp <- make_predictions(m)
# make_three_maps(mp)
```

Difference map (departure from baseline)

```{r}
make_comparison_map <- function(spp_preds,yr=2050){
  basepred <- map_year(spp_preds,yr_vec=1985:2010,return_pred_df = T)
  newpred <- map_year(spp_preds,yr_vec=yr,return_pred_df = T) %>% 
    mutate(est_comp=est) %>% dplyr::select(-est)
  both <- basepred %>% 
    left_join(newpred,by=c('X','Y')) %>% 
    mutate(diffpred=est_comp-est)
  
  bbox=st_bbox(projection_extent)
  
  out<-ggplot(coast)+
    geom_sf()+
    # geom_point(data=df,aes(longitude,latitude,col=exp(est)))+
    # scale_color_viridis_c()+
    geom_raster(data=both,aes(x=X,y=Y,fill=diffpred),interpolate=F)+
    scale_fill_gradient2(low = "red", mid = "white", high = "purple")+
    coord_sf(datum=NA)+
    # scale_fill_viridis_c(direction = -1)+
    xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
    labs(x="",y="",fill="Difference")
  out
}
```

Wrapper to make two difference maps (mid century minus baseline and end of century minus baseline)

```{r}
make_two_diff_maps <- function(spp_preds){
  p1 <- make_comparison_map(spp_preds,yr=c(2040:2060))+labs(title='mid century')
  p2 <- make_comparison_map(spp_preds,yr=c(2040:2060))+labs(title='end of century')
  plot_grid(p1,p2,nrow=1)
}
```

```{r,fig.height=5,fig.width=5.3}
# make_two_diff_maps(mp)
```

## Environmental Relationships

We can take advantage of the `visreg` package to look at the environmental response. We want to pull the data from it to back-transform the predictors into their original scale.

```{r}
plot_single_env_relationship <- function(fit,predvar="depth_trawl_norm"){
  
  # use visreg to do the fit for us
  x <- visreg(fit=fit,xvar=predvar,scale='response',plot=FALSE)
  datfit <- x$fit
  
  # if the response is not substrate, have to back-transform the predictor variable
  if(predvar=="prop_hard_mixed"){
    out = datfit %>% 
      ggplot(aes(prop_hard_mixed,visregFit,ymax=visregUpr,ymin=visregLwr))+
      geom_ribbon(fill='gray70')+
      geom_line()+
      labs(x=predvar,y="CPUE")+
      theme(panel.border = element_rect(color='black',fill=NA))
    
  } else {
      # find the mean and SD of the original predictor to back-transform
      original_predvar = str_sub(predvar,end=-6L)
      
      predvar_dat=fit$data %>% dplyr::select(all_of(original_predvar)) %>% pluck(1)
      mean_predvar <- mean(predvar_dat,na.rm=T)
      sd_predvar <- sd(predvar_dat,na.rm=T)
      predvar_trans <- (datfit %>% pluck(predvar))*sd_predvar+mean_predvar
      
      # add the transformation to the fitted relationship
      datfit <- datfit %>% 
        mutate(envtrans=predvar_trans)
      
      # output plot
      out = datfit %>% 
        ggplot(aes(envtrans,visregFit,ymax=visregUpr,ymin=visregLwr))+
        geom_ribbon(fill='gray70')+
        geom_line()+
        labs(x=original_predvar,y="CPUE")+
        theme(panel.border = element_rect(color='black',fill=NA))
  }
  
  out

}

```

```{r}
# plot_single_env_relationship(fit = m)
```

Conditional plots for environmental relationships. Important note here: the way that `visreg` seems to work is that it calculates a range of values of the focal variable (here, `predvar`), then predicts using the model and *the median values of the other variables*. So, these environmental relationships need to be viewed as conditional responses of CPUE to the given variable, conditional on the other variables being held at their median values.

```{r}
plot_all_env_relationships <- function(m){
  
  form <-m$formula %>% pluck(1) %>% as.character() %>% pluck(3)
  
  if(grepl('depth_trawl_norm',form)){
    depth_relationship <- plot_single_env_relationship(m,predvar="depth_trawl_norm")+
      labs(x="Depth (m)")
  } else depth_relationship=NULL
  
  if(grepl('prop_hard_mixed',form)){
    hab_relationship <- plot_single_env_relationship(m,predvar="prop_hard_mixed")+
      labs(x="Proportion Hard Substrate")
  } else hab_relationship=NULL
    
  if(grepl('temp_roms',form)){
    temp_relationship <- plot_single_env_relationship(m,predvar="mean_temp_roms_30_norm")+
      labs(x="Bottom Temperature")
  } else temp_relationship=NULL
  
  if(grepl('mean_oxygen_roms_30_norm',form)){
    oxy_relationship <- plot_single_env_relationship(m,predvar="mean_oxygen_roms_30_norm")+
      labs(x="Bottom Oxygen")
  } else oxy_relationship=NULL
  
  cowplot::plot_grid(temp_relationship,oxy_relationship,depth_relationship,hab_relationship,nrow=2)
}
```


```{r,fig.height=8,fig.width=8}
# plot_all_env_relationships(m)
```

## Wrap Everything

Wrap all of the above to make a "push go" function for a species to:

*   Run a model with specified params
*   Save all diagnostic plots
*   Save all environmental affinity plots
*   Make projections under 3 ESMs (Hadley, IPSL, GFDL)
*   Make maps and difference maps
*   Save projected data

```{r}
process_species_model <- function(dat=trawl_roms_utm,species,incl_depth,incl_substrate,dir.out=here::here('Janelle','models_owen')){
  tic(paste(species,"processing total time"))
  tic(paste("Fitting model for",species))
  # run model
  m <- run_sdmTMB(dat = trawl_roms_utm,spp=species,use_depth=incl_depth,time_vary = F,spatial_field = F,include_substrate = incl_substrate,hab_spline = T,env_spline = T,spline_k = 3)
  toc()
  
  # make diagnostic plots
  tic(paste("Making diagnostic plots for",species))
  diagplots <- make_diagnostic_plots(m)
  toc()
  
  # make env affinity plots
  tic(paste("Making environmental affinity plots for",species))
  envplots <- plot_all_env_relationships(m)
  toc()
  
  # make projections
  tic(paste("Making projections for",species))
  hadl <- make_predictions(m,'hadl') %>% mutate(esm='hadl')
  ipsl <- make_predictions(m,'ipsl') %>% mutate(esm='ipsl')
  gfdl <- make_predictions(m,'gfdl') %>% mutate(esm='gfdl')
  all_projections <- bind_rows(hadl,ipsl,gfdl)
  toc()
  
  # make maps
  tic(paste("Making maps for",species))
  hadl_predmap <- make_three_maps(hadl)
  hadl_diffmap <- make_two_diff_maps(hadl)
  ipsl_predmap <- make_three_maps(ipsl)
  ipsl_diffmap <- make_two_diff_maps(ipsl)
  gfdl_predmap <- make_three_maps(gfdl)
  gfdl_diffmap <- make_two_diff_maps(gfdl)
  toc()
  
  # save everything
  tic(paste("Saving everything for",species))
  save.dir=paste0(dir.out,"/",species,"/")
  if(!dir.exists(save.dir)) dir.create(save.dir)
  write_rds(all_projections,file=paste0(save.dir,"projection_3ESMs.rds"))
  ggsave(paste0(save.dir,'diagnostic_plots.png'),diagplots,h=8,w=8,bg='white')
  ggsave(paste0(save.dir,'environmental_affinities.png'),envplots,h=8,w=8,bg='white')
  ggsave(paste0(save.dir,'hadl_preds.png'),hadl_predmap,h=5,w=8,bg='white')
  ggsave(paste0(save.dir,'ipsl_preds.png'),ipsl_predmap,h=5,w=8,bg='white')
  ggsave(paste0(save.dir,'gfdl_preds.png'),gfdl_predmap,h=5,w=8,bg='white')
  ggsave(paste0(save.dir,'hadl_diffmap.png'),hadl_diffmap,h=8,w=5.3,bg='white')
  ggsave(paste0(save.dir,'ipsl_diffmap.png'),ipsl_diffmap,h=8,w=5.3,bg='white')
  ggsave(paste0(save.dir,'gfdl_diffmap.png'),gfdl_diffmap,h=8,w=5.3,bg='white')
  toc()
  
  print(paste("Model and plots finished for",species))
  toc()
}
```

# Apply to Species

California halibut: depth and substrate
```{r}
process_species_model(species='california halibut',incl_depth = T,incl_substrate = T)
```

