---
title: "Species Selection Example Models"
author: "Janelle Layton"
date: '2022-06-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages
```{r}
library(tidyverse)
library(vista)
library(mgcv)
```

# Load Data

```{r}
rmarkdown::render(here::here('scripts','sdmTMB data construction.Rmd'),quiet=TRUE)
```

# sdmTMB Model Function

```{r}
prepare_species <- function(dat,spp){
  dat_sub <- dat %>% 
    filter(species==spp) %>% 
    
    # rescale depth, oxygen, and temp to be N(0,1)
    mutate(across(c(depth_trawl,mean_temp_roms_30,mean_oxygen_roms_30),~(scale(.) %>% as.vector()),.names="{.col}_norm")) %>% 
    
    # add a year indicator
    mutate(year=lubridate::year(date))
}
```

```{r}
run_sdmTMB <- function(dat,spp,use_depth=F,time_vary=F,spatial_field=F,include_substrate=T,hab_spline=T,env_spline=T,spline_k=3){
  # filter data for species
  modeldat <- prepare_species(dat,spp=spp)
  
  # make spde
  spde <- make_mesh(modeldat,xy_cols = c('longitude','latitude'), 
                   cutoff = 20)
  
  # model formula
  formula <- paste0("cpue_kg_km2 ~ ")
  
  # substrate relationship
  substrate <- paste("prop_hard_mixed + I(prop_hard_mixed^2)")
  #wiggly habitat relationship?
  substrate <- ifelse(hab_spline, paste0("s(prop_hard_mixed,k=",spline_k,")"),
                      substrate)
  if(!include_substrate) substrate=""
  
  # make the environmental effects
  enviro <- paste("mean_temp_roms_30_norm + 
                  I(mean_temp_roms_30_norm^2) + 
                  mean_oxygen_roms_30_norm + 
                  I(mean_oxygen_roms_30_norm^2)")
  # wiggly environmental relationships?
  enviro <- ifelse(env_spline, paste0("s(mean_temp_roms_30_norm,k=",spline_k,") + ",
                                      "s(mean_oxygen_roms_30_norm,k=",spline_k,")"),
                   enviro)
  # if depth effect, add to model formula
  if(use_depth) {
    formula = paste0(formula, " + depth_trawl_norm + I(depth_trawl_norm^2)")
  }
  
  time_formula = "~ -1"
  if(time_vary) {
    time_formula = paste0(time_formula, " + ", substrate, " + ", enviro)
    time_varying = as.formula(time_formula)
    time = "year"
  } else {
    formula = paste0(formula, " + ", substrate, " + ", enviro)
    time_varying = NULL
    time = "year"
  }
  
  # fit model. EW commented out quadratic roots, since those are still experimental and won't work for all spp. Also turned
  # set.seed(41389) # for reproducibility
  # test_set = sample(1:nrow(modeldat), size = round(0.1*nrow(modeldat)), replace=FALSE)
  # modeldat$fold = 1
  # modeldat$fold[test_set] = 2
  # anisotropy off for now
  print('running model.')
  m <- try( sdmTMB(
    formula = as.formula(formula),
    time_varying = time_varying,
    mesh = spde,
    time = time,
    family = tweedie(link = "log"),
    data = modeldat,
    anisotropy = FALSE,
    spatial = spatial_field,
    #extra_time argument necessary for prediction?
    extra_time=1980:2100),
  silent=F)


  # predicted values for the 2nd fold (test)
  # m_cv$data$cv_predicted[which(m_cv$data$cv_fold==2)]
  # log likelihood values for the 2nd fold (test)
  # m_cv$data$cv_loglik[which(m_cv$data$cv_fold==2)]

    # sum(m_cv$data$cv_loglik[which(m_cv$data$cv_fold==2)])
  
  # if(class(m)!="try-error") {
  #   write_rds(m, file=here::here('model output',
  #                                paste0(spp,'.rds')))
  # }
  if(class(m)=="try-error"){
    print(paste("Error."))
  }else{
    print(paste("Model for",spp,"complete."))
  }

  # return(m)
  return(m)

}
```

## Choose a species

```{r}
spp <- "pacific ocean perch"

spp_dat <- prepare_species(trawl_roms_utm,spp)

```

# Model output

```{r}
m <- run_sdmTMB(dat=trawl_roms_utm,spp="pacific ocean perch",use_depth=F,env_spline=T,include_substrate=T)
summary(m)
```

# Normal QQ plot

```{r}
resids <- residuals(m)[1:nrow(spp_dat),1]
qqnorm(resids)
qqline(resids)
spp_dat$resid=resids
dev.copy(jpeg,'perch1_qqe.jpg', width=5, height=6, units='in', res=300)
dev.off()
```

# Spatial Predictions

```{r,fig.height=6, fig.width=5}
preds <- predict(m) %>% filter(year %in% unique(spp_dat$year)) %>% pull('est')
spp_dat$pred <- preds

# predictions, with time and space
preds_time_space <- pred_space(spp_dat,X='longitude',Y='latitude',time='year') + coord_equal()

preds_time_space

dev.copy(jpeg,'perch1_timespace.jpg', width=5, height=6, units='in', res=300)
dev.off()

```


```{r, fig.height=8, fig.width=5}
# predictions, across space but averaged over time
preds_space <- pred_space(spp_dat,X='longitude',Y='latitude',time='year',by_time = F)

preds_space

dev.copy(jpeg,'pred_perch1_space.jpg', width=5, height=6, units='in', res=300)
dev.off()

```

Spatial Residuals

```{r, fig.height=8, fig.width=5}
resids_space <- resid_space(spp_dat,X='longitude',Y='latitude',time='year')
resids_space
dev.copy(jpeg,'perch1_residspace.jpg', width=5, height=6, units='in', res=300)
dev.off()
```


# Make predictions

```{r}
make_predictions <- function(modelobj,gcm="hadl"){

  original_model_data <- modelobj$data
  # need to use the original (hindcast) environmental data to scale the projected data
  mean_t <- mean(original_model_data$mean_temp_roms_30,na.rm=T)
  sd_t <- sd(original_model_data$mean_temp_roms_30,na.rm=T)
  mean_oxy <- mean(original_model_data$mean_oxygen_roms_30,na.rm=T)
  sd_oxy <- sd(original_model_data$mean_oxygen_roms_30,na.rm=T)
  mean_depth <- mean(original_model_data$depth_trawl,na.rm=T)
  sd_depth <- sd(original_model_data$depth_trawl,na.rm=T)
  
  # create a new tibble with the projected data for the chosen gcm
  newdata <- roms %>% 
    ## NOTE: WE LOSE A TON OF DATA HERE, SOMETHING MAY BE WRONG WITH THE HABITAT DATA
    left_join(hab,by=c("lat","lon")) %>% 
    drop_na() %>% 
    mutate(depth_m= -depth_m) %>%
    dplyr::select(year,lat,lon,latitude,longitude,prop_hard_mixed,depth_m,contains(paste0("30d_",gcm)))
  
  temperature <- newdata %>% dplyr::select(contains('bt')) %>% 
    set_names('temperature') %>% 
    mutate(mean_temp_roms_30_norm=(temperature-mean_t)/sd_t)
  
  oxygen <- newdata %>% dplyr::select(contains('oxy')) %>% 
    set_names('oxygen') %>% 
    mutate(mean_oxygen_roms_30_norm=(oxygen-mean_oxy)/sd_oxy)
  
  depth_new <- newdata %>% dplyr::select(contains('depth')) %>% 
    set_names('depth') %>% 
    mutate(depth_trawl_norm=(depth-mean_depth)/sd_depth)

  newdata <- newdata %>% 
    bind_cols(temperature) %>% 
    bind_cols(oxygen) %>% 
    bind_cols(depth_new) %>% 
    dplyr::select(-temperature,-oxygen, -depth) %>%
    mutate(year=as.double(year))
  
  # years to predict
  yrs <- sort(unique(newdata$year))
  # now we can make the predictions
  predicted_cpue_km2 <- predict(modelobj,newdata,return_tmb_object=F,extra_time=yrs)
  
  predicted_cpue_km2
}
```

```{r}
preds = make_predictions(m, gcm = "hadl")
```


# mapping, annual map
```{r}
# pretty grid for raster maps
gr <- projection_extent %>% st_make_grid(cellsize=10,what='centers') %>% st_as_sf() %>% 
  st_intersection(projection_extent)
gr_xy <- st_coordinates(gr)
```

```{r}
map_year <- function(model_ens,yr_vec=c(2000),return_pred_df=F,plot_leg=T){
  # scale for the legend, common within species
  scl <- c(0,quantile(exp(model_ens$est),0.99))
  df <- model_ens %>% 
    filter(year%in%yr_vec) %>% 
    group_by(longitude,latitude) %>% 
    summarise(est=mean(est,na.rm=T) %>% exp()) %>% 
    # rescale super large positive outliers for mapping purposes
    mutate(est=ifelse(est>scl[2],scl[2],est))
  # match nearest neighbors from predictions to grid
  pred_points <- df %>% dplyr::select(longitude,latitude) %>% as.matrix()
  nns <- nn2(pred_points,gr_xy,k=1)$nn.idx
  gr_pred <- gr_xy %>% as_tibble() %>% mutate(est=df$est[nns])
  
  bbox=st_bbox(projection_extent)
  
  if(return_pred_df) {out <- gr_pred %>% as_tibble()}
  
  else{
    out<-ggplot(coast)+
      geom_sf()+
      # geom_point(data=df,aes(longitude,latitude,col=exp(est)))+
      # scale_color_viridis_c()+
      geom_raster(data=gr_pred,aes(x=X,y=Y,fill=est),interpolate=F)+
      scale_fill_viridis_c(limits=scl)+
      xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
      labs(x="",y="",fill="CPUE",title='')
    if(!plot_leg) out <- out + theme(legend.position = 'None')
  }
  out
}
```



```{r}
map_year(preds,yr_vec = 2025)
dev.copy(jpeg,'perch1_annualmap2025.jpg', width=5, height=5, units='in', res=300)
dev.off()
```


# mapping, difference map
```{r}
make_comparison_map <- function(model_ens,yr=2050){
  basepred <- map_year(model_ens,yr_vec=1980:2010,return_pred_df = T)
  newpred <- map_year(model_ens,yr_vec=yr,return_pred_df = T) %>% 
    mutate(est_comp=est) %>% dplyr::select(-est)
  both <- basepred %>% 
    left_join(newpred,by=c('X','Y')) %>% 
    mutate(diffpred=est_comp-est)
  
  bbox=st_bbox(projection_extent)
  
  out<-ggplot(coast)+
        geom_sf()+
        # geom_point(data=df,aes(longitude,latitude,col=exp(est)))+
        # scale_color_viridis_c()+
        geom_raster(data=both,aes(x=X,y=Y,fill=diffpred),interpolate=F)+
        scale_fill_gradient2(low = "red", mid = "white", high = "purple")+
        # scale_fill_viridis_c(direction = -1)+
        xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
        labs(x="",y="",fill="Difference")
  out
}
```

```{r}
make_comparison_map(preds, yr = 2025:2100)
dev.copy(jpeg,'perch1_diffmap20252100.jpg', width=5, height=5, units='in', res=300)
dev.off()
```

## Environmental Relationships

We can take advantage of the `visreg` package to look at the environmental response. We want to pull the data from it to back-transform the predictors into their original scale
```{r}
library(visreg)

plot_single_env_relationship <- function(fit,predvar="depth_trawl_norm"){
  
  # use visreg to do the fit for us
  x <- visreg(fit=fit,xvar=predvar,scale='response',plot=FALSE)
  datfit <- x$fit
  
  # if the response is not substrate, have to back-transform the predictor variable
  if(predvar=="prop_hard_mixed"){
    out = datfit %>% 
      ggplot(aes(prop_hard_mixed,visregFit,ymax=visregUpr,ymin=visregLwr))+
      geom_ribbon(fill='gray70')+
      geom_line()+
      labs(x=predvar,y="CPUE")+
      theme(panel.border = element_rect(color='black',fill=NA))
    
  } else {
      # find the mean and SD of the original predictor to back-transform
      original_predvar = str_sub(predvar,end=-6L)
      
      predvar_dat=fit$data %>% dplyr::select(all_of(original_predvar)) %>% pluck(1)
      mean_predvar <- mean(predvar_dat,na.rm=T)
      sd_predvar <- sd(predvar_dat,na.rm=T)
      predvar_trans <- (datfit %>% pluck(predvar))*sd_predvar+mean_predvar
      
      # add the transformation to the fitted relationship
      datfit <- datfit %>% 
        mutate(envtrans=predvar_trans)
      
      # output plot
      out = datfit %>% 
        ggplot(aes(envtrans,visregFit,ymax=visregUpr,ymin=visregLwr))+
        geom_ribbon(fill='gray70')+
        geom_line()+
        labs(x=original_predvar,y="CPUE")+
        theme(panel.border = element_rect(color='black',fill=NA))
  }
  
  out

}

```

```{r}
plot_single_env_relationship(fit = m)
```




Conditional plots for environmental relationships. Important note here: the way that `visreg` seems to work is that it calculates a range of values of the focal variable (here, `predvar`), then predicts using the model and *the median values of the other variables*. So, these environmental relationships need to be viewed as conditional responses of CPUE to the given variable, conditional on the other variables being held at their median values.

```{r}
depth_relationship <- plot_single_env_relationship(m,predvar="depth_trawl_norm")
oxy_relationship <- plot_single_env_relationship(m,predvar="mean_oxygen_roms_30_norm")
temp_relationship <- plot_single_env_relationship(m,predvar="mean_temp_roms_30_norm")
hab_relationship <- plot_single_env_relationship(m,predvar="prop_hard_mixed")

cowplot::plot_grid(depth_relationship,oxy_relationship,temp_relationship,hab_relationship,nrow=2)
```


```{r}
make_env <- function(modelobj){
  
  df <- modelobj$data
  f <- modelobj$formula %>% as.character() %>% pluck(3)
  
  # get means for normalization
  meanhab <- mean(df$prop_hard_mixed)
  meant <- mean(df$mean_temp_roms_30_norm)
  meano <- mean(df$mean_oxygen_roms_30_norm)
  meanlat <- mean(df$latitude)
  meanlon <- mean(df$longitude)
  
  if(grepl('prop_hard_mixed',f)){
    habdf <- tibble(prop_hard_mixed=seq(0,1,length.out=100),year=2000, mean_temp_roms_30_norm=meant, mean_oxygen_roms_30_norm=meano,longitude=meanlon,latitude=meanlat)
    
    phab <- predict(modelobj,newdata=habdf,se_fit=T)
    
    habplot <- ggplot(phab,aes(prop_hard_mixed,exp(est),ymin=exp(est-1.96*est_se), ymax=exp(est+1.96*est_se)))+
      geom_line()+
      geom_ribbon(alpha=0.5)+
      labs(title="Habitat",x="Prop Hard Bottom",y="CPUE")
    
  } else{habplot <- ggplot(df %>% slice(1:100),aes(prop_hard_mixed, cpue_kg_km2)) + geom_blank()+
      labs(title="Habitat",x="Prop Hard Bottom",y="CPUE")}
  
  
  if(grepl('mean_temp',f)){
    
    tdf <- tibble(mean_temp_roms_30_norm=seq(min(df$mean_temp_roms_30_norm)+0.2,max(df$mean_temp_roms_30_norm)-0.2,length.out=100),year=2000,prop_hard_mixed=meanhab,mean_oxygen_roms_30_norm=meano,longitude=meanlon,latitude=meanlat)
    
    pt <- predict(modelobj,newdata=tdf,se_fit=T)
    
    # rescale for plotting (un-normalize)
    pt <- pt %>% mutate(t=mean_temp_roms_30_norm*sd(df$mean_temp_roms_30,na.rm=T)+mean(df$mean_temp_roms_30,na.rm=T))
    
    tplot <- ggplot(pt,aes(t,exp(est),ymin=exp(est-1.96*est_se),ymax=exp(est+1.96*est_se)))+
      geom_line()+
      geom_ribbon(alpha=0.5)+
      labs(title="Temperature",x="Temperature",y="CPUE")
  }else{tplot <- ggplot(df %>% slice(1:100),aes(prop_hard_mixed, cpue_kg_km2)) + geom_blank()+
      labs(title="Temperature",x="Temperature",y="CPUE")}
  
  
  if(grepl('mean_oxygen',f)){
    odf <- tibble(mean_oxygen_roms_30_norm=seq(min(df$mean_oxygen_roms_30_norm)+0.2,max(df$mean_oxygen_roms_30_norm)-0.2,length.out=100),year=2000,prop_hard_mixed=meanhab,mean_temp_roms_30_norm=meant,longitude=meanlon,latitude=meanlat)
    
    po <- predict(modelobj,newdata=odf,se_fit=T)
    
    # rescale for plotting (un-normalize)
    po <- po %>% mutate(o=mean_oxygen_roms_30_norm*sd(df$mean_oxygen_roms_30,na.rm=T)+mean(df$mean_oxygen_roms_30,na.rm=T))
    
    oplot <- ggplot(po,aes(o,exp(est),ymin=exp(est-1.96*est_se),ymax=exp(est+1.96*est_se)))+
      geom_line()+
      geom_ribbon(alpha=0.5)+
      labs(title="Oxygen",x="Oxygen (mmol/m3)",y="CPUE")
    
  }else{
    oplot <- ggplot(df %>% slice(1:100),aes(prop_hard_mixed, cpue_kg_km2)) + 
      geom_blank()+
      labs(title="Oxygen",x="Oxygen (mmol/m3)",y="CPUE")}
  
  out <- plot_grid(habplot,tplot,oplot,nrow=1)
  
  # add text for model formula
  ftext <- f %>% str_replace_all("prop_hard_mixed","habitat") %>% 
    str_replace_all("mean_temp_roms_30_norm","temperature") %>% 
    str_replace_all("mean_oxygen_roms_30_norm","oxygen")
  
  formulagg <- ggdraw() +
    draw_label(paste('CPUE ~',ftext),size = 8)
  plot_grid(out,formulagg,ncol=1,rel_heights = c(1,0.2))+plot_theme
}
```

```{r}

```


