---
title: "Atlantis Grid SDM Translation"
author: "Owen Liu"
date: "9/3/2021"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r, include=FALSE}
# devtools::install_github("pbs-assess/sdmTMB")
library(sdmTMB)
library(tidyverse)
library(lubridate)
library(rnaturalearth)
library(rbgm)
library(sf)
library(here)

knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)
```

```{r}
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

# General Plan

We need to tranlsate SDM outputs into relative abundance for Atlantis. We will do this by taking projected SDM outputs, averaging CPUE by box. Next, to extrapolate beyond the trawl survey domain, we will use Atlantis polygons paired by depth to make an assumption of equal densities in paired boxes. Finally, we will multiply by polygon area, then divide by the total across all polygons, to obtain an overall relative biomass value for each box.

# Import Data

## ROMS and Atlantis Spatial Grids

```{r}
# Atlantis grid
fl <- here::here("data",'grids',"CalCurrentV3_utm.bgm")
# load the file
bgm <- read_bgm(fl)
names(bgm)
emocc_sf <- box_sf(bgm) %>% dplyr::select(box_id,label,area)
emocc_box_areas <- emocc_sf %>% st_set_geometry(NULL) %>% dplyr::select(-label)

# st_crs(emocc_sf) <- st_crs(attr(emocc_sf$geometry, "crs")$proj)

# the ROMS grid, clipped to the SDM projection extent
roms <- read_rds(here('data','roms_for_sdm_projection.rds'))
roms_sf <- roms %>%
  # just find the unique grid cells (we don't care about time series, only grid cell locations right now)
  dplyr::distinct(lat,lon,longitude,latitude) %>% 
  st_as_sf(coords=c('longitude','latitude'),crs= "+proj=utm +zone=10 +datum=WGS84 +units=km")


# load west cost land for mapping
coaststates <- ne_countries(continent="North America",returnclass = 'sf') %>% 
  filter(name %in% c('Canada','United States','Mexico')) %>% 
  st_transform(st_crs(roms_sf))
```

Convert Atlantis BGM to the same CRS as ROMS

```{r}
emocc_sf <- emocc_sf %>% 
  st_transform(st_crs(roms_sf))
```

A quick plot to show what we are starting with

```{r}
bbox=st_bbox(emocc_sf)
ggplot()+
  geom_sf(data=coaststates,fill='gray50')+
  geom_sf(data=emocc_sf,fill='blue',alpha=0.5)+
  geom_sf(data=roms_sf,size=0.1)+
  xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])
```


##  Atlantis Coupled Boxes

Credit to Pierre-Yves. NOTE: THIS NEEDS TO BE UPDATED 09/10/21

```{r}
cps <- read_csv(here('data','grids','Atlantis_polygon_couples.csv'),col_types = 'dc')
glimpse(cps)

# use just first match for now
cps <- cps %>% 
  mutate(first_match=str_sub(polygon_ref,1,2) %>% as.integer) %>% 
  dplyr::select(polygon,first_match) %>% 
  rename(box_id=polygon)
glimpse(cps)
```


## SDM Projection Functions

Load functions from projection script

```{r}
rmarkdown::render(here::here('scripts','sdmTMB_projection.Rmd'),quiet=TRUE)
```

# Spatial Join

Match the ROMS points to an Atlantis box

```{r}
roms_atlantis_spatial_join <- roms_sf %>% 
  st_join(emocc_sf)
glimpse(roms_atlantis_spatial_join)
```

Now we have a key matching ROMS points to their appropriate boxes


# Species Distribution Models

Run an example ensemble species distribution model and project using the IPSL ESM

```{r}
sable_models <- read_rds(here::here('model output','sablefish models.rds'))
sable_ens <- ensemble_predictions(sable_models,gcm='ipsl')
```

# Join to Atlantis Polygons

Write a function to extract mean predictions for each Atlantis polygon, by first joining points to polygons using the join key we created above

```{r}
sdm_to_atlantis <- function(preds){
  
  # filter to the year of interest and exponentiate the estimated dependent variable
  df <- preds %>% 
    group_by(year,lon,lat) %>% 
    summarise(est=mean(ens_est,na.rm=T) %>% exp()) %>% 
    drop_na() %>% 
    ungroup()
  
  # join the SDM/Atlantis matching key
  atlantis_matched <- df %>% 
    left_join(roms_atlantis_spatial_join,by=c('lon','lat')) %>% 
    
    # group by Atlantis box and calculate a mean
    group_by(box_id,year) %>% 
    summarise(mean_est=mean(est,na.rm=T)) %>% 
    ungroup() %>% 
    drop_na()
    
  # join the paired Atlantis boxes
  atlantis_mean_cpue_all_boxes <- atlantis_matched %>% 
    complete(box_id=full_seq(0:88,1),year) %>% 
    left_join(cps,by=c('box_id')) %>% 
    # join the estimates again, using the matched box instead of the original box
    left_join(atlantis_matched,by=c('first_match'='box_id','year')) %>% 
    # coalesce original and matched estimates
    mutate(est=coalesce(mean_est.x,mean_est.y)) %>%
    mutate(est=replace_na(est,0)) %>% 
    dplyr::select(box_id,year,est)
  
  # add in area information, scale up, then normalize across all boxes
  atlantis_est_final <- atlantis_mean_cpue_all_boxes %>% 
    left_join(emocc_box_areas,by=c('box_id')) %>% 
    group_by(year) %>% 
    # multiply by total area
    mutate(totest = est*area) %>% 
    # finally, calculate relative index
    mutate(rel_est = totest/sum(totest,na.rm=T)) %>% 
    mutate(perc_est = rel_est*100) %>% 
    ungroup()
  
  return(atlantis_est_final)
}
```

If you want to plot, can pick a year and join the spatial polygon data

```{r}
plot_sdm_year <- function(atlantis_estimates,yr=2050){
  polys<- atlantis_estimates %>% 
    filter(year==yr) %>% 
    left_join(emocc_sf,by=c('box_id','area')) %>% 
    st_as_sf()
  bbox=st_bbox(emocc_sf)

  out <- ggplot()+
    geom_sf(data=coaststates,fill='gray50')+
    geom_sf(data=polys,aes(fill=perc_est),alpha=1)+
    scale_fill_viridis_c()+
    xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
    labs(fill=paste("Percent of Total Abundance by Box,",yr))
  
  plot(out)
}
```


Test with our example model

```{r}
sable_to_atlantis <- sdm_to_atlantis(sable_ens)
plot_sdm_year(sable_to_atlantis,2013)

```


