---
title: "Climate change effects on future groundfish availability to U.S. west coast fishing ports"
author: "Owen Liu"
date: '2022-07-05'
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
# devtools::install_github("pbs-assess/sdmTMB")
library(sdmTMB)
library(tidyverse)
library(magrittr)
library(lubridate)
library(rnaturalearth)
library(sf)
library(here)
library(viridis)
# vista is Eric Ward's library for looking at outputs
library(vista)
library(cowplot)
library(ggalt)
library(RANN)
library(furrr)
library(future)
library(tictoc)
library(ggsci)
library(ggpubr)

knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)
```

```{r}
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=12,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
pal <- viridis_pal(option="A",begin=0.2,end=0.8)(4)
pal6 <- viridis_pal(option="C",begin=0.2,end=0.8)(6)
pal6_2 <- viridis_pal(option="B",begin=0.2,end=0.8)(6)
```

Load functions and data used for models. This command sources code and data from the `sdmTMB model functions` and `sdmTMB data construction` scripts.

```{r}
rmarkdown::render(here::here('scripts','sdmTMB model functions.Rmd'),quiet=TRUE)
```

# Model Species

Use `sdmTMB` to make ensemble models for each species of interest. (if these models have already been run, can just load them rather than running them again)

```{r,eval=F}
spp_to_model <- c("dover sole","sablefish","shortspine thornyhead","longspine thornyhead")
```

Run models and save cross-validation model stacking weights. (note: this chunk is turned to eval=F for now)

## Fit Models

```{r,eval=F}
for(i in 1:length(spp_to_model)){
  s <- spp_to_model[i]
  m <- model_species(s,data = trawl_roms_utm,use_substrate = F)
  write_rds(m,here::here('model output','dts paper',paste(s,'models.rds')))
  print(paste(s,'models finished'))
}
```

## Make Ensemble Predictions

We project across all 3 ESMs, weighting each equally

```{r,eval=F}
dover_models <- read_rds(here::here('model output','dts paper','dover sole models.rds'))
sable_models <- read_rds(here::here('model output','dts paper','sablefish models.rds'))
ss_models <- read_rds(here::here('model output','dts paper','shortspine thornyhead models.rds'))
ls_models <- read_rds(here::here('model output','dts paper','longspine thornyhead models.rds'))
```

IPSL
```{r,eval=F}
dover_ens_ipsl <- ensemble_predictions(dover_models,gcm='ipsl',nsims=100)
sable_ens_ipsl <- ensemble_predictions(sable_models,gcm='ipsl',nsims=100)
ss_ens_ipsl <- ensemble_predictions(ss_models,gcm="ipsl",nsims=100)
ls_ens_ipsl <- ensemble_predictions(ls_models,gcm='ipsl',nsims=100)

write_rds(dover_ens_ipsl,here('model output','dts paper','dover_ensemble_preds_ipsl.rds'))
write_rds(sable_ens_ipsl,here('model output','dts paper','sable_ensemble_preds_ipsl.rds'))
write_rds(ss_ens_ipsl,here('model output','dts paper','ss_ensemble_preds_ipsl.rds'))
write_rds(ls_ens_ipsl,here('model output','dts paper','ls_ensemble_preds_ipsl.rds'))
```

GFDL
```{r,eval=F}
dover_ens_gfdl <- ensemble_predictions(dover_models,gcm='gfdl',nsims=100)
sable_ens_gfdl <- ensemble_predictions(sable_models,gcm='gfdl',nsims=100)
ss_ens_gfdl <- ensemble_predictions(ss_models,gcm="gfdl",nsims=100)
ls_ens_gfdl <- ensemble_predictions(ls_models,gcm='gfdl',nsims=100)

write_rds(dover_ens_gfdl,here('model output','dts paper','dover_ensemble_preds_gfdl.rds'))
write_rds(sable_ens_gfdl,here('model output','dts paper','sable_ensemble_preds_gfdl.rds'))
write_rds(ss_ens_gfdl,here('model output','dts paper','ss_ensemble_preds_gfdl.rds'))
write_rds(ls_ens_gfdl,here('model output','dts paper','ls_ensemble_preds_gfdl.rds'))
```

Hadley
```{r,eval=F}
dover_ens_hadl <- ensemble_predictions(dover_models,gcm='hadl',nsims=100)
sable_ens_hadl <- ensemble_predictions(sable_models,gcm='hadl',nsims=100)
ss_ens_hadl <- ensemble_predictions(ss_models,gcm="hadl",nsims=100)
ls_ens_hadl <- ensemble_predictions(ls_models,gcm='hadl',nsims=100)

write_rds(dover_ens_hadl,here('model output','dts paper','dover_ensemble_preds_hadl.rds'))
write_rds(sable_ens_hadl,here('model output','dts paper','sable_ensemble_preds_hadl.rds'))
write_rds(ss_ens_hadl,here('model output','dts paper','ss_ensemble_preds_hadl.rds'))
write_rds(ls_ens_hadl,here('model output','dts paper','ls_ensemble_preds_hadl.rds'))
```

### Join 3 ESM Predictions

```{r,eval=F}
dover_ens_sims <-list(dover_ens_gfdl,dover_ens_ipsl,dover_ens_hadl) %>% bind_rows() %>% 
    dplyr::select(-esm) %>% 
    group_by(year,longitude,latitude,lat,lon,depth_m) %>%
    nest() %>% 
    ungroup() %>% 
    mutate(data=purrr::map(data,function(x)as.numeric(as.matrix(x))),
           species='dover')

sable_ens_sims <-list(sable_ens_gfdl,sable_ens_ipsl,sable_ens_hadl) %>% bind_rows() %>% 
    dplyr::select(-esm) %>% 
    group_by(year,longitude,latitude,lat,lon,depth_m) %>%
    nest() %>% 
    ungroup() %>% 
    mutate(data=purrr::map(data,function(x)as.numeric(as.matrix(x))),
           species='sable')

ss_ens_sims <-list(ss_ens_gfdl,ss_ens_ipsl,ss_ens_hadl) %>% bind_rows() %>% 
    dplyr::select(-esm) %>% 
    group_by(year,longitude,latitude,lat,lon,depth_m) %>%
    nest() %>% 
    ungroup() %>% 
    mutate(data=purrr::map(data,function(x)as.numeric(as.matrix(x))),
           species='ss')

ls_ens_sims <-list(ls_ens_gfdl,ls_ens_ipsl,ls_ens_hadl) %>% bind_rows() %>% 
    dplyr::select(-esm) %>% 
    group_by(year,longitude,latitude,lat,lon,depth_m) %>%
    nest() %>% 
    ungroup() %>% 
    mutate(data=purrr::map(data,function(x)as.numeric(as.matrix(x))),
           species='ls')

write_rds(sable_ens_sims,here('model output','dts paper','sable_ensemble_all_sims.rds'))
write_rds(dover_ens_sims,here('model output','dts paper','dover_ensemble_all_sims.rds'))
write_rds(ss_ens_sims,here('model output','dts paper','ss_ensemble_all_sims.rds'))
write_rds(ls_ens_sims,here('model output','dts paper','ls_ensemble_all_sims.rds'))
```

```{r,eval=F}
dover_ens_all <- purrr::map_df(list(dover_ens_gfdl,dover_ens_ipsl,dover_ens_hadl),function(df){
  df %>% pivot_longer(contains('sim'),names_to="sim",values_to="est")
}) %>% 
  group_by(year,longitude,latitude,lat,lon,depth_m) %>% 
  summarise(median_est=median(est),mean_est=mean(est),est5=quantile(est,0.05),est95=quantile(est,0.95)) %>% 
  ungroup()

sable_ens_all <- purrr::map_df(list(sable_ens_gfdl,sable_ens_ipsl,sable_ens_hadl),function(df){
  df %>% pivot_longer(contains('sim'),names_to="sim",values_to="est")
}) %>% 
  group_by(year,longitude,latitude,lat,lon,depth_m) %>% 
  summarise(median_est=median(est),mean_est=mean(est),est5=quantile(est,0.05),est95=quantile(est,0.95)) %>% 
  ungroup()

ss_ens_all <- purrr::map_df(list(ss_ens_gfdl,ss_ens_ipsl,ss_ens_hadl),function(df){
  df %>% pivot_longer(contains('sim'),names_to="sim",values_to="est")
}) %>% 
  group_by(year,longitude,latitude,lat,lon,depth_m) %>% 
  summarise(median_est=median(est),mean_est=mean(est),est5=quantile(est,0.05),est95=quantile(est,0.95)) %>% 
  ungroup()

ls_ens_all <- purrr::map_df(list(ls_ens_gfdl,ls_ens_ipsl,ls_ens_hadl),function(df){
  df %>% pivot_longer(contains('sim'),names_to="sim",values_to="est")
}) %>% 
  group_by(year,longitude,latitude,lat,lon,depth_m) %>% 
  summarise(median_est=median(est),mean_est=mean(est),est5=quantile(est,0.05),est95=quantile(est,0.95)) %>% 
  ungroup()

write_rds(sable_ens_all,here('model output','dts paper','sable_ensemble_preds_combined.rds'))
write_rds(dover_ens_all,here('model output','dts paper','dover_ensemble_preds_combined.rds'))
write_rds(ss_ens_all,here('model output','dts paper','ss_ensemble_preds_combined.rds'))
write_rds(ls_ens_all,here('model output','dts paper','ls_ensemble_preds_combined.rds'))
```

Load

```{r}
dover_ens_all <- read_rds(here('model output','dts paper','dover_ensemble_preds_combined.rds')) %>% 
  mutate(species='dover')
  # rename(dover_median=median_est,dover_lower=est5,dover_upper=est95)
sable_ens_all <- read_rds(here('model output','dts paper','sable_ensemble_preds_combined.rds'))%>% 
  mutate(species='sable')
  # rename(sable_median=median_est,sable_lower=est5,sable_upper=est95)
ss_ens_all <- read_rds(here('model output','dts paper','ss_ensemble_preds_combined.rds'))%>% 
  mutate(species='ss')
  # rename(ss_median=median_est,ss_lower=est5,ss_upper=est95)
ls_ens_all <- read_rds(here('model output','dts paper','ls_ensemble_preds_combined.rds'))%>% 
  mutate(species='ls')
  # rename(ls_median=median_est,ls_lower=est5,ls_upper=est95)

# with sims
dover_ens_sims <- read_rds(here('model output','dts paper','dover_ensemble_all_sims.rds'))
sable_ens_sims <- read_rds(here('model output','dts paper','sable_ensemble_all_sims.rds'))
ss_ens_sims <- read_rds(here('model output','dts paper','ss_ensemble_all_sims.rds'))
ls_ens_sims <- read_rds(here('model output','dts paper','ls_ensemble_all_sims.rds'))
```

Join. We need the data in slightly different forms for different analyses

```{r}
# combined ensemble abundance
four_spp <- list(dover_ens_all,sable_ens_all,ss_ens_all,ls_ens_all) %>% bind_rows() %>% 
  # join roms data
  left_join(roms,by = c("year", "longitude", "latitude", "lat", "lon"))

four_spp_sims <- list(dover_ens_sims,sable_ens_sims,ss_ens_sims,ls_ens_sims) %>% bind_rows() %>% 
  # add summary columns
  mutate(mean_est=purrr::map_dbl(data,mean),
         est5=purrr::map_dbl(data,quantile,probs=0.05),
         est95=purrr::map_dbl(data,quantile,probs=0.95))

four_spp_list <- list("dover"=bind_rows(dover_ens_ipsl,dover_ens_gfdl,dover_ens_hadl),
                      "sable"=bind_rows(sable_ens_ipsl,sable_ens_gfdl,sable_ens_hadl),
                      "ss"=bind_rows(ss_ens_ipsl,ss_ens_gfdl,ss_ens_hadl),
                      "ls"=bind_rows(ls_ens_ipsl,ls_ens_gfdl,ls_ens_hadl))
```

### Calculate Center-of-Gravity

```{r,eval=F}
#ipsl
dover_cog_ipsl <- make_cog(dover_models,gcm = 'ipsl',what="dat")
sable_cog_ipsl <- make_cog(sable_models,gcm = 'ipsl',what="dat")
ss_cog_ipsl <- make_cog(ss_models,gcm = 'ipsl',what="dat")
ls_cog_ipsl <- make_cog(ls_models,gcm = 'ipsl',what="dat")

write_rds(dover_cog,here('model output','dts paper','dover_cog_ipsl.rds'))
write_rds(sable_cog,here('model output','dts paper','sable_cog_ipsl.rds'))
write_rds(ss_cog,here('model output','dts paper','ss_cog_ipsl.rds'))
write_rds(ls_cog,here('model output','dts paper','ls_cog_ipsl.rds'))
```

# Manuscript Figures

## Figure 1

Landings

```{r}
landings_plot <- v %>% 
  group_by(Year,lab)%>% 
  filter(Year>1979) %>% 
  summarise(tot_dollars=sum(Dollars,na.rm=T)) %>% 
  ggplot(aes(Year,tot_dollars/1e6,color=lab))+geom_line(size=1.5)+
  labs(x="Year",y="Landings (Million $)",color="")+
  # scale_x_continuous(breaks=seq(2010,2020,by=2))+
  scale_color_manual(values=pal)+
  scale_x_continuous(expand=c(0,2))+
  theme(legend.position = c(0.15,0.8),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12),
        panel.border = element_rect(color='black',fill=NA))
```

Fishing footprints

```{r}
bbox=st_bbox(projection_extent)
footprints_map <- ggplot()+
  geom_sf(data=footprints,aes(fill=port_name),alpha=0.5)+
  geom_sf(data=coast,fill='gray80')+
  # geom_sf(data=isobaths,col='gray20')+
  geom_sf(data=port_coords,aes(color=port_name),size=3)+
  scale_fill_npg()+
  scale_color_npg()+
  guides(color='none')+
  # guides(color="none",fill='none')+
  coord_sf(datum=NA)+
  labs(fill="Port")+
  xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
  theme(
        legend.text=element_text(size=8),
        legend.background = element_rect(fill='white',color='black'),
        legend.position=c(0.6,0.6))
```

Overlap map (count number of species in each cell). This figure shows the number of species out of four in each grid cell that are at or above their median CPUE (across all cells) for that year.

```{r,fig.width=6,fig.height=8}

multimapdiff <- map_multispp_diff(four_spp,qlower=0.75,qupper=1)

```

Finish Figure 1

```{r}
row2 <- plot_grid(footprints_map,multimapdiff,labels=c('b','c'),nrow=1,rel_widths = c(0.8,1))
multimapcomp <- plot_grid(landings_plot,row2,nrow=2,rel_heights = c(0.3,0.7),labels='auto')
# multimapcomp

ggsave(here('model output','dts paper','fig1.png'),multimapcomp,w=6,h=8)
```

## Figure 2

Weighted distance from shore
```{r}
fourspp_reldist <- purrr::map(four_spp_list,calc_rel_dist_to_shore)
fourspp_reldist <- bind_rows(fourspp_reldist,.id='species')

#historical average
fourspp_reldist_hist <- fourspp_reldist %>% 
  filter(year %in% c(1985:2010)) %>% 
  group_by(species,lat,sim) %>% 
  summarise(mean_dist_hist=mean(dist)) %>% 
  ungroup()

# end of century
fourspp_reldist_fut <- fourspp_reldist %>% 
  filter(year %in% c(2075:2100)) %>% 
  group_by(species,lat,sim) %>% 
  summarise(mean_dist_fut=mean(dist)) %>% 
  ungroup()

# bind again
fourspp_reldist_change <- fourspp_reldist_hist %>% 
  left_join(fourspp_reldist_fut,by=c('species','lat','sim')) %>%
  mutate(change=mean_dist_fut-mean_dist_hist)
  # group_by(species,lat) %>% 
  # summarise(meanchange=mean(change),
  #           change5=quantile(change,0.05),
  #           change95=quantile(change,0.95))
```

Make the plot

```{r}
# four_spp_dist_to_shore_change <- fourspp_reldist_change %>%
#     mutate(spp_plotting=case_when(
#       species=='dover' ~ "Dover Sole",
#       species=="ls" ~ "Longspine",
#       species=="ss"~"Shortspine",
#       species=="sable" ~"Sablefish"
#     )) %>% 
#     mutate(spp_plotting=factor(spp_plotting,levels=c("Dover Sole","Sablefish","Shortspine","Longspine"))) %>% 
#   ggplot(aes(x=lat,y=meanchange,ymin=change5,ymax=change95,fill=spp_plotting,col=spp_plotting))+
#   geom_ribbon(alpha=0.2,linetype=3)+
#   geom_line(size=1)+
#   geom_hline(yintercept=0,linetype=2)+
#   # ylim(-6,25)+
#   scale_color_manual(values=pal4)+
#   scale_fill_manual(values=pal4)+
#   # scale_y_reverse(limits=c(90,20))+
#   coord_flip()+
#   # theme(legend.position = c(0.8,0.7))+
#   labs(y="Change in Distance from Shore (km)",x="Latitude",col="Species",fill="Species")+
#   theme(panel.background = element_rect(fill=NA,color='black'))
four_spp_dist_to_shore_change <- fourspp_reldist_change %>%
    mutate(spp_plotting=case_when(
      species=='dover' ~ "Dover Sole",
      species=="ls" ~ "Longspine",
      species=="ss"~"Shortspine",
      species=="sable" ~"Sablefish"
    )) %>% 
    mutate(spp_plotting=factor(spp_plotting,levels=c("Dover Sole","Sablefish","Shortspine","Longspine"))) %>% 
  ggplot(aes(x=lat,y=change,fill=spp_plotting,col=spp_plotting))+
  geom_point(alpha=0.2,size=0.7)+
  geom_hline(yintercept=0,linetype=2)+
  geom_smooth()+
  scale_color_manual(values=pal4)+
  scale_fill_manual(values=pal4)+
  # scale_y_reverse(limits=c(90,20))+
  coord_flip()+
  # theme(legend.position = c(0.8,0.7))+
  labs(y="Change in Distance from Shore (km)",x="Latitude",col="Species",fill="Species")+
  theme(panel.background = element_rect(fill=NA,color='black'))

four_spp_dist_to_shore_change
ggsave(here('model output','dts paper','four species distance from shore.png'),four_spp_dist_to_shore_change,w=6,h=6)
```

Depth change

```{r}
fourspp_depth <- plot_depth_distribution2(four_spp_sims,start_year = 2020,end_year=2100)+theme(panel.background = element_rect(fill=NA,color='black'))+theme(legend.position = 'bottom')
fourspp_depth
ggsave(here('model output','dts paper','fig_depth_change_four_spp.png'),fourspp_depth,w=6,h=4)
```


```{r,fig.width=7,fig.height=5}
leg <- get_legend(fourspp_depth)
fig_dist_depth_change<-plot_grid(four_spp_dist+theme(plot.margin=unit(c(0.5,0.5,1.5,0.5),'cm')),
                                 fourspp_depth + guides(color='none',fill='none') + theme(plot.margin=unit(c(0.5,0.5,1.5,0.5),'cm')),
                                 nrow=1,labels='auto')
fig_dist_depth_change<-ggdraw()+
  draw_plot(fig_dist_depth_change)+
  draw_plot(leg,y=-0.45)

ggsave(here('model output','dts paper','fig_dist_depth_change.png'),fig_dist_depth_change,w=8,h=5)
```

## Figure 3

Environmental state space within fishing footprints

```{r}
roms_env_space <- roms %>% 
  dplyr::select(year,lat,lon,mean_bt_30d_ipsl,mean_oxy_bottom_30d_ipsl) %>% 
  distinct() %>% 
  left_join(roms_ll,by=c('lon','lat')) %>% 
  drop_na() %>% 
  rename(t=mean_bt_30d_ipsl,oxy=mean_oxy_bottom_30d_ipsl) %>% 
  mutate(depth_bin=cut(depth_m,breaks=c(0,-100,-250,-500,-1000,-3000),
                       labels=c("1000-3000m","500-1000m","250-500m","100-250m","0-100m")),
         lat_bin=cut(lat,breaks=c(30,33,36,39,42,45,48),labels=c("30-33","33-36","36-39","39-42","42-45","45-48")))
roms_env_space_hist <- roms_env_space %>%
  # historical "mean" is 1985-2010
  filter(year%in%c(1985:2010)) %>%
  group_by(lat,lon,roms_cell,depth_m,depth_bin,lat_bin) %>% 
  summarise(t_hist=mean(t,na.rm=T),oxy_hist=mean(oxy,na.rm=T)) %>% 
  ungroup()

roms_env_space_fut <- roms_env_space %>%
  # historical "mean" is 1985-2010
  filter(year%in%c(2075:2100)) %>%
  group_by(lat,lon,roms_cell,depth_m,depth_bin,lat_bin) %>% 
  summarise(t_fut=mean(t,na.rm=T),oxy_fut=mean(oxy,na.rm=T)) %>% 
  ungroup()

roms_env_space_sf <- roms_env_space_hist %>% 
  left_join(roms_env_space_fut, c("lat", "lon", "roms_cell", "depth_m", "depth_bin", "lat_bin")) %>%
  st_as_sf(coords=c('lon','lat'),crs=4326) %>% 
  st_transform(st_crs(footprints)) %>% 
  st_join(footprints)

footprint_env_space <- roms_env_space_sf %>% 
  drop_na() %>% 
  group_by(port_name,port_group) %>% 
  summarise(mean_t_hist=mean(t_hist),mean_oxy_hist=mean(oxy_hist),
            mean_t_fut=mean(t_fut),mean_oxy_fut=mean(oxy_fut)) %>% 
  ungroup()
```

Make environmental affinities ellipses

```{r}
library(ggnewscale)
aff_polys<-affinities_ellipses(four_spp,yr_vec=1985:2010,concavity=7,threshold=0.25)
# plotting order for colors
polys_and_footprints <- ggplot()+
  geom_polygon(data=aff_polys,aes(V1,V2,col=spp_plotting),fill=NA,size=1,linetype=3)+
  scale_color_manual(values=c(pal4),name="Top 75% of\nspecies' CPUE")+
  new_scale_color()+
  geom_segment(data=footprint_env_space,aes(x=mean_t_hist,y=mean_oxy_hist,xend=mean_t_fut,yend=mean_oxy_fut,
             col=port_name),size=1,linetype=1,arrow=ggplot2::arrow(type="open",length=unit(0.1,'inches')))+
  scale_color_npg(name="Port Group\nClimate")+
  labs(x="Bottom Temperature (deg C)",y="Bottom Oxygen (mmol/m3)")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor=element_blank(),
        panel.border = element_rect(color='black',fill=NA),
        axis.text=element_text(size=12))
```

```{r,fig.width=8,fig.height=8}
# attach dumbbell plots
footprint_db_temp_plot2 <- footprint_env_space %>% 
  ggplot(aes(x=mean_t_hist,xend=mean_t_fut,y=port_name))+
  geom_dumbbell(color="#FCAE91",colour_x =NA,colour_xend = "#A50F15",
                size=1,size_x=3,size_xend = 3) +
  labs(x="Temperature",y="")+
  coord_flip()+
  theme(panel.border = element_rect(color='black',fill=NA),
        axis.text.x=element_text(color=pal_npg()(6),size=10))

footprint_db_oxy_plot2 <- footprint_env_space %>% 
  ggplot(aes(x=mean_oxy_hist,xend=mean_oxy_fut,y=port_name))+
  geom_dumbbell(color="#a3c4dc",colour_x =NA,colour_xend = "#0e668b",
                size=1,size_x=3,size_xend = 3) +
  labs(x="Oxygen",y="")+
  coord_flip()+
  theme(axis.text.x = element_blank(),panel.border = element_rect(color='black',fill=NA))

fig3 <- plot_grid(footprint_db_temp_plot2,footprint_db_oxy_plot2,polys_and_footprints,labels='auto',nrow=3,
                  hjust = 0,vjust=1,
                  rel_heights = c(0.2,0.2,0.6))
fig3
ggsave(here('model output','dts paper','fig_env_change_env_affs.png'),fig3,h=8,w=8)
```

## Figure 4

```{r}
```{r}
ens_est_overlaps <- dover_ens %>% nest(dover = c(longitude, latitude, lat, lon, depth_m, ens_est)) %>% 
  # join all the data and cogs nested by year
  left_join(sable_ens %>% nest(sable = c(longitude, latitude, lat, lon, depth_m, ens_est)),by = "year") %>% 
  left_join(ss_ens %>% nest(ss = c(longitude, latitude, lat, lon, depth_m, ens_est)),by = "year") %>% 
  left_join(ls_ens %>% nest(ls = c(longitude, latitude, lat, lon, depth_m, ens_est)),by = "year") %>% 
  left_join(dover_cog %>% nest(dover_cog=c(w.est_x,w.est_y)),by = "year")%>% 
  left_join(sable_cog %>% nest(sable_cog=c(w.est_x,w.est_y)),by = "year")%>% 
  left_join(ss_cog %>% nest(ss_cog=c(w.est_x,w.est_y)),by = "year") %>%  
  left_join(ls_cog %>% nest(ls_cog=c(w.est_x,w.est_y)),by = "year") %>% 
  
  # calculate the gic
  mutate(dover_sable_gic=purrr::pmap_dbl(list(dat1=dover,dat2=sable,cog1=dover_cog,cog2=sable_cog),gic)) %>%
  mutate(dover_ss_gic=purrr::pmap_dbl(list(dat1=dover,dat2=ss,cog1=dover_cog,cog2=ss_cog),gic)) %>% 
  mutate(dover_ls_gic=purrr::pmap_dbl(list(dat1=dover,dat2=ls,cog1=dover_cog,cog2=ls_cog),gic)) %>% 
  mutate(sable_ss_gic=purrr::pmap_dbl(list(dat1=sable,dat2=ss,cog1=sable_cog,cog2=ss_cog),gic)) %>% 
  mutate(sable_ls_gic=purrr::pmap_dbl(list(dat1=sable,dat2=ls,cog1=sable_cog,cog2=ls_cog),gic)) %>% 
  mutate(ss_ls_gic=purrr::pmap_dbl(list(dat1=ss,dat2=ls,cog1=ss_cog,cog2=ls_cog),gic)) %>% 
  
  # calculate Schoener's D
  mutate(dover_sable_sch=purrr::pmap_dbl(list(dat1=dover,dat2=sable),schoeners)) %>% 
  mutate(dover_ss_sch=purrr::pmap_dbl(list(dat1=dover,dat2=ss),schoeners)) %>% 
  mutate(dover_ls_sch=purrr::pmap_dbl(list(dat1=dover,dat2=ls),schoeners)) %>% 
  mutate(sable_ss_sch=purrr::pmap_dbl(list(dat1=sable,dat2=ss),schoeners)) %>% 
  mutate(sable_ls_sch=purrr::pmap_dbl(list(dat1=sable,dat2=ls),schoeners)) %>% 
  mutate(ss_ls_sch=purrr::pmap_dbl(list(dat1=ss,dat2=ls),schoeners)) %>% 
  
  # calculate local colocation
  mutate(dover_sable_loc=purrr::pmap_dbl(list(dat1=dover,dat2=sable),lic)) %>% 
  mutate(dover_ss_loc=purrr::pmap_dbl(list(dat1=dover,dat2=ss),lic)) %>% 
  mutate(dover_ls_loc=purrr::pmap_dbl(list(dat1=dover,dat2=ls),lic)) %>%
  mutate(sable_ss_loc=purrr::pmap_dbl(list(dat1=sable,dat2=ss),lic)) %>%  
  mutate(sable_ls_loc=purrr::pmap_dbl(list(dat1=sable,dat2=ls),lic)) %>% 
  mutate(ss_ls_loc=purrr::pmap_dbl(list(dat1=ss,dat2=ls),lic))

```
```

