---
title: "Climate change effects on future groundfish availability to U.S. west coast fishing ports"
author: "Owen Liu"
date: '2022-07-05'
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
# devtools::install_github("pbs-assess/sdmTMB")
library(sdmTMB)
library(tidyverse)
library(magrittr)
library(lubridate)
library(rnaturalearth)
library(sf)
library(here)
library(viridis)
# vista is Eric Ward's library for looking at outputs
library(vista)
library(cowplot)
library(ggalt)
library(RANN)
library(furrr)
library(future)
library(tictoc)
library(ggsci)
library(ggpubr)

knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)
```

```{r}
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=12,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
pal <- viridis_pal(option="A",begin=0.2,end=0.8)(4)
pal4 <- viridis_pal(option="A",begin=0.2,end=0.8)(4)
pal6 <- viridis_pal(option="C",begin=0.2,end=0.8)(6)
pal6_2 <- viridis_pal(option="B",begin=0.2,end=0.8)(6)
```

Load functions and data used for models. This command sources code and data from the `sdmTMB model functions` and `sdmTMB data construction` scripts.

```{r}
rmarkdown::render(here::here('scripts','sdmTMB model functions.Rmd'),quiet=TRUE)
```

# Model Species

Use `sdmTMB` to make ensemble models for each species of interest. (if these models have already been run, can just load them rather than running them again)

```{r,eval=F}
spp_to_model <- c("dover sole","sablefish","shortspine thornyhead","longspine thornyhead")
```

Run models and save cross-validation model stacking weights. (note: this chunk is turned to eval=F for now)

## Fit Models

```{r,eval=F}
for(i in 1:length(spp_to_model)){
  s <- spp_to_model[i]
  m <- model_species(s,data = trawl_roms_utm,use_substrate = F)
  write_rds(m,here::here('model output','dts paper',paste(s,'models.rds')))
  print(paste(s,'models finished'))
}
```

## Make Ensemble Predictions

We project across all 3 ESMs, weighting each equally

```{r,eval=T}
dover_models <- read_rds(here::here('model output','dts paper','dover sole models.rds'))
sable_models <- read_rds(here::here('model output','dts paper','sablefish models.rds'))
ss_models <- read_rds(here::here('model output','dts paper','shortspine thornyhead models.rds'))
ls_models <- read_rds(here::here('model output','dts paper','longspine thornyhead models.rds'))
```

IPSL
```{r,eval=F}
dover_ens_ipsl <- ensemble_predictions(dover_models,gcm='ipsl',nsims=100)
sable_ens_ipsl <- ensemble_predictions(sable_models,gcm='ipsl',nsims=100)
ss_ens_ipsl <- ensemble_predictions(ss_models,gcm="ipsl",nsims=100)
ls_ens_ipsl <- ensemble_predictions(ls_models,gcm='ipsl',nsims=100)

write_rds(dover_ens_ipsl,here('model output','dts paper','dover_ensemble_preds_ipsl.rds'))
write_rds(sable_ens_ipsl,here('model output','dts paper','sable_ensemble_preds_ipsl.rds'))
write_rds(ss_ens_ipsl,here('model output','dts paper','ss_ensemble_preds_ipsl.rds'))
write_rds(ls_ens_ipsl,here('model output','dts paper','ls_ensemble_preds_ipsl.rds'))
```

GFDL
```{r,eval=F}
dover_ens_gfdl <- ensemble_predictions(dover_models,gcm='gfdl',nsims=100)
sable_ens_gfdl <- ensemble_predictions(sable_models,gcm='gfdl',nsims=100)
ss_ens_gfdl <- ensemble_predictions(ss_models,gcm="gfdl",nsims=100)
ls_ens_gfdl <- ensemble_predictions(ls_models,gcm='gfdl',nsims=100)

write_rds(dover_ens_gfdl,here('model output','dts paper','dover_ensemble_preds_gfdl.rds'))
write_rds(sable_ens_gfdl,here('model output','dts paper','sable_ensemble_preds_gfdl.rds'))
write_rds(ss_ens_gfdl,here('model output','dts paper','ss_ensemble_preds_gfdl.rds'))
write_rds(ls_ens_gfdl,here('model output','dts paper','ls_ensemble_preds_gfdl.rds'))
```

Hadley
```{r,eval=F}
dover_ens_hadl <- ensemble_predictions(dover_models,gcm='hadl',nsims=100)
sable_ens_hadl <- ensemble_predictions(sable_models,gcm='hadl',nsims=100)
ss_ens_hadl <- ensemble_predictions(ss_models,gcm="hadl",nsims=100)
ls_ens_hadl <- ensemble_predictions(ls_models,gcm='hadl',nsims=100)

write_rds(dover_ens_hadl,here('model output','dts paper','dover_ensemble_preds_hadl.rds'))
write_rds(sable_ens_hadl,here('model output','dts paper','sable_ensemble_preds_hadl.rds'))
write_rds(ss_ens_hadl,here('model output','dts paper','ss_ensemble_preds_hadl.rds'))
write_rds(ls_ens_hadl,here('model output','dts paper','ls_ensemble_preds_hadl.rds'))
```

### Join 3 ESM Predictions

```{r,eval=F}
dover_ens_sims <-list(dover_ens_gfdl,dover_ens_ipsl,dover_ens_hadl) %>% bind_rows() %>% 
    dplyr::select(-esm) %>% 
    group_by(year,longitude,latitude,lat,lon,depth_m) %>%
    nest() %>% 
    ungroup() %>% 
    mutate(data=purrr::map(data,function(x)as.numeric(as.matrix(x))),
           species='dover')

sable_ens_sims <-list(sable_ens_gfdl,sable_ens_ipsl,sable_ens_hadl) %>% bind_rows() %>% 
    dplyr::select(-esm) %>% 
    group_by(year,longitude,latitude,lat,lon,depth_m) %>%
    nest() %>% 
    ungroup() %>% 
    mutate(data=purrr::map(data,function(x)as.numeric(as.matrix(x))),
           species='sable')

ss_ens_sims <-list(ss_ens_gfdl,ss_ens_ipsl,ss_ens_hadl) %>% bind_rows() %>% 
    dplyr::select(-esm) %>% 
    group_by(year,longitude,latitude,lat,lon,depth_m) %>%
    nest() %>% 
    ungroup() %>% 
    mutate(data=purrr::map(data,function(x)as.numeric(as.matrix(x))),
           species='ss')

ls_ens_sims <-list(ls_ens_gfdl,ls_ens_ipsl,ls_ens_hadl) %>% bind_rows() %>% 
    dplyr::select(-esm) %>% 
    group_by(year,longitude,latitude,lat,lon,depth_m) %>%
    nest() %>% 
    ungroup() %>% 
    mutate(data=purrr::map(data,function(x)as.numeric(as.matrix(x))),
           species='ls')

write_rds(sable_ens_sims,here('model output','dts paper','sable_ensemble_all_sims.rds'))
write_rds(dover_ens_sims,here('model output','dts paper','dover_ensemble_all_sims.rds'))
write_rds(ss_ens_sims,here('model output','dts paper','ss_ensemble_all_sims.rds'))
write_rds(ls_ens_sims,here('model output','dts paper','ls_ensemble_all_sims.rds'))
```

```{r,eval=F}
dover_ens_all <- purrr::map_df(list(dover_ens_gfdl,dover_ens_ipsl,dover_ens_hadl),function(df){
  df %>% pivot_longer(contains('sim'),names_to="sim",values_to="est")
}) %>% 
  group_by(year,longitude,latitude,lat,lon,depth_m) %>% 
  summarise(median_est=median(est),mean_est=mean(est),est5=quantile(est,0.05),est95=quantile(est,0.95)) %>% 
  ungroup()

sable_ens_all <- purrr::map_df(list(sable_ens_gfdl,sable_ens_ipsl,sable_ens_hadl),function(df){
  df %>% pivot_longer(contains('sim'),names_to="sim",values_to="est")
}) %>% 
  group_by(year,longitude,latitude,lat,lon,depth_m) %>% 
  summarise(median_est=median(est),mean_est=mean(est),est5=quantile(est,0.05),est95=quantile(est,0.95)) %>% 
  ungroup()

ss_ens_all <- purrr::map_df(list(ss_ens_gfdl,ss_ens_ipsl,ss_ens_hadl),function(df){
  df %>% pivot_longer(contains('sim'),names_to="sim",values_to="est")
}) %>% 
  group_by(year,longitude,latitude,lat,lon,depth_m) %>% 
  summarise(median_est=median(est),mean_est=mean(est),est5=quantile(est,0.05),est95=quantile(est,0.95)) %>% 
  ungroup()

ls_ens_all <- purrr::map_df(list(ls_ens_gfdl,ls_ens_ipsl,ls_ens_hadl),function(df){
  df %>% pivot_longer(contains('sim'),names_to="sim",values_to="est")
}) %>% 
  group_by(year,longitude,latitude,lat,lon,depth_m) %>% 
  summarise(median_est=median(est),mean_est=mean(est),est5=quantile(est,0.05),est95=quantile(est,0.95)) %>% 
  ungroup()

write_rds(sable_ens_all,here('model output','dts paper','sable_ensemble_preds_combined.rds'))
write_rds(dover_ens_all,here('model output','dts paper','dover_ensemble_preds_combined.rds'))
write_rds(ss_ens_all,here('model output','dts paper','ss_ensemble_preds_combined.rds'))
write_rds(ls_ens_all,here('model output','dts paper','ls_ensemble_preds_combined.rds'))
```

### Load

```{r}
# separate projections, with sims
dover_ens_separate <- list.files(here('model output','dts paper'),full.names = T) %>% str_subset('dover_ensemble_preds') %>% 
  str_subset("combined",negate=T) %>% map_df(read_rds)
sable_ens_separate <- list.files(here('model output','dts paper'),full.names = T) %>%
  str_subset('sable_ensemble_preds') %>% 
  str_subset("combined",negate=T) %>% map_df(read_rds)
ss_ens_separate <- list.files(here('model output','dts paper'),full.names = T) %>%
  str_subset('ss_ensemble_preds') %>% 
  str_subset("combined",negate=T) %>% map_df(read_rds)
ls_ens_separate <- list.files(here('model output','dts paper'),full.names = T) %>%
  str_subset('ls_ensemble_preds') %>% 
  str_subset("combined",negate=T) %>% map_df(read_rds)

# combined across ESMs
dover_ens_all <- read_rds(here('model output','dts paper','dover_ensemble_preds_combined.rds')) %>% 
  mutate(species='dover')
  # rename(dover_median=median_est,dover_lower=est5,dover_upper=est95)
sable_ens_all <- read_rds(here('model output','dts paper','sable_ensemble_preds_combined.rds'))%>% 
  mutate(species='sable')
  # rename(sable_median=median_est,sable_lower=est5,sable_upper=est95)
ss_ens_all <- read_rds(here('model output','dts paper','ss_ensemble_preds_combined.rds'))%>% 
  mutate(species='ss')
  # rename(ss_median=median_est,ss_lower=est5,ss_upper=est95)
ls_ens_all <- read_rds(here('model output','dts paper','ls_ensemble_preds_combined.rds'))%>% 
  mutate(species='ls')
  # rename(ls_median=median_est,ls_lower=est5,ls_upper=est95)

# with sims
dover_ens_sims <- read_rds(here('model output','dts paper','dover_ensemble_all_sims.rds'))
sable_ens_sims <- read_rds(here('model output','dts paper','sable_ensemble_all_sims.rds'))
ss_ens_sims <- read_rds(here('model output','dts paper','ss_ensemble_all_sims.rds'))
ls_ens_sims <- read_rds(here('model output','dts paper','ls_ensemble_all_sims.rds'))
```

Join. We need the data in slightly different forms for different analyses

```{r}
# separate ESMs
four_spp_list <- list("dover"=dover_ens_separate,
                      "sable"=sable_ens_separate,
                      "ss"=ss_ens_separate,
                      "ls"=ls_ens_separate)
# combined ensemble abundance
four_spp <- list(dover_ens_all,sable_ens_all,ss_ens_all,ls_ens_all) %>% bind_rows() %>% 
  # join roms data
  left_join(roms,by = c("year", "longitude", "latitude", "lat", "lon"))

# with separate sims
four_spp_sims <- list(dover_ens_sims,sable_ens_sims,ss_ens_sims,ls_ens_sims) %>% bind_rows() %>% 
  # add summary columns
  mutate(mean_est=purrr::map_dbl(data,mean),
         est5=purrr::map_dbl(data,quantile,probs=0.05),
         est95=purrr::map_dbl(data,quantile,probs=0.95))


```

### Calculate Center-of-Gravity

```{r,eval=F}
#ipsl
# dover_cog_ipsl <- make_cog(dover_models,gcm = 'ipsl',what="dat")
# sable_cog_ipsl <- make_cog(sable_models,gcm = 'ipsl',what="dat")
# ss_cog_ipsl <- make_cog(ss_models,gcm = 'ipsl',what="dat")
# ls_cog_ipsl <- make_cog(ls_models,gcm = 'ipsl',what="dat")
# 
# write_rds(dover_cog,here('model output','dts paper','dover_cog_ipsl.rds'))
# write_rds(sable_cog,here('model output','dts paper','sable_cog_ipsl.rds'))
# write_rds(ss_cog,here('model output','dts paper','ss_cog_ipsl.rds'))
# write_rds(ls_cog,here('model output','dts paper','ls_cog_ipsl.rds'))
```

# Manuscript Figures

## Figure 1

Landings

```{r}
landings_plot <- v %>% 
  group_by(Year,lab)%>% 
  filter(Year>1979) %>% 
  summarise(tot_dollars=sum(Dollars,na.rm=T)) %>% 
  ggplot(aes(Year,tot_dollars/1e6,color=lab))+geom_line(size=1.5)+
  labs(x="Year",y="Landings (Million $)",color="")+
  # scale_x_continuous(breaks=seq(2010,2020,by=2))+
  scale_color_manual(values=pal)+
  scale_x_continuous(expand=c(0,2))+
  theme(legend.position = c(0.15,0.8),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12),
        panel.border = element_rect(color='black',fill=NA))
```

Fishing footprints

```{r}
bbox=st_bbox(projection_extent)
footprints_map <- ggplot()+
  geom_sf(data=footprints,aes(fill=port_name),alpha=0.5)+
  geom_sf(data=coast,fill='gray80')+
  # geom_sf(data=isobaths,col='gray20')+
  geom_sf(data=port_coords,aes(color=port_name),size=3)+
  scale_fill_npg()+
  scale_color_npg()+
  guides(color='none')+
  # guides(color="none",fill='none')+
  coord_sf(datum=NA)+
  labs(fill="Port")+
  xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
  theme(
        legend.text=element_text(size=8),
        legend.background = element_rect(fill='white',color='black'),
        legend.position=c(0.6,0.6))
```

Overlap map (count number of species in each cell). This figure shows the number of species out of four in each grid cell that are at or above their median CPUE (across all cells) for that year.

```{r,fig.width=6,fig.height=8}

multimapdiff <- map_multispp_diff(four_spp,qlower=0.75,qupper=1)

```

Finish Figure 1

```{r}
row2 <- plot_grid(footprints_map,multimapdiff,labels=c('b','c'),nrow=1,rel_widths = c(0.8,1))
multimapcomp <- plot_grid(landings_plot,row2,nrow=2,rel_heights = c(0.3,0.7),labels='auto')
# multimapcomp

ggsave(here('model output','dts paper','fig1.png'),multimapcomp,w=6,h=8)
```

## Figure 2

Weighted distance from shore
```{r}
fourspp_reldist <- purrr::map(four_spp_list,calc_rel_dist_to_shore)
fourspp_reldist <- bind_rows(fourspp_reldist,.id='species')

#historical average
fourspp_reldist_hist <- fourspp_reldist %>% 
  filter(year %in% c(1985:2010)) %>% 
  group_by(species,lat,esm,sim) %>% 
  summarise(mean_dist_hist=mean(dist)) %>% 
  ungroup()

# end of century
fourspp_reldist_fut <- fourspp_reldist %>% 
  filter(year %in% c(2075:2100)) %>% 
  group_by(species,lat,esm,sim) %>% 
  summarise(mean_dist_fut=mean(dist)) %>% 
  ungroup()

# bind again
fourspp_reldist_change <- fourspp_reldist_hist %>% 
  left_join(fourspp_reldist_fut,by=c('species','lat','esm','sim')) %>%
  mutate(change=mean_dist_fut-mean_dist_hist)
  # group_by(species,lat) %>% 
  # summarise(meanchange=mean(change),
  #           change5=quantile(change,0.05),
  #           change95=quantile(change,0.95))
```

Make the plot

```{r}
# four_spp_dist_to_shore_change <- fourspp_reldist_change %>%
#     mutate(spp_plotting=case_when(
#       species=='dover' ~ "Dover Sole",
#       species=="ls" ~ "Longspine",
#       species=="ss"~"Shortspine",
#       species=="sable" ~"Sablefish"
#     )) %>% 
#     mutate(spp_plotting=factor(spp_plotting,levels=c("Dover Sole","Sablefish","Shortspine","Longspine"))) %>% 
#   ggplot(aes(x=lat,y=meanchange,ymin=change5,ymax=change95,fill=spp_plotting,col=spp_plotting))+
#   geom_ribbon(alpha=0.2,linetype=3)+
#   geom_line(size=1)+
#   geom_hline(yintercept=0,linetype=2)+
#   # ylim(-6,25)+
#   scale_color_manual(values=pal4)+
#   scale_fill_manual(values=pal4)+
#   # scale_y_reverse(limits=c(90,20))+
#   coord_flip()+
#   # theme(legend.position = c(0.8,0.7))+
#   labs(y="Change in Distance from Shore (km)",x="Latitude",col="Species",fill="Species")+
#   theme(panel.background = element_rect(fill=NA,color='black'))
four_spp_dist_to_shore_change <- fourspp_reldist_change %>%
    mutate(spp_plotting=case_when(
      species=='dover' ~ "Dover Sole",
      species=="ls" ~ "Longspine",
      species=="ss"~"Shortspine",
      species=="sable" ~"Sablefish"
    )) %>% 
    mutate(spp_plotting=factor(spp_plotting,levels=c("Dover Sole","Sablefish","Shortspine","Longspine"))) %>% 
  mutate(ESM=toupper(esm)) %>% 
  ggplot(aes(x=lat,y=change,fill=spp_plotting,col=spp_plotting))+
  geom_point(alpha=0.2,size=0.7)+
  geom_hline(yintercept=0,linetype=2)+
  geom_smooth()+
  scale_y_continuous(limits=c(-20,20))+
  facet_wrap(~ESM,nrow=1)+
  scale_color_manual(values=pal4)+
  scale_fill_manual(values=pal4)+
  # scale_y_reverse(limits=c(90,20))+
  coord_flip()+
  # theme(legend.position = c(0.8,0.7))+
  labs(y="Change in Distance from Shore (km)",x="Latitude",col="Species",fill="Species")+
  theme(panel.background = element_rect(fill=NA,color='black'))

four_spp_dist_to_shore_change
```
```{r}
ggsave(here('model output','dts paper','four species distance from shore.png'),four_spp_dist_to_shore_change,w=8,h=6)
```


Depth change

```{r}
fourspp_depth <- purrr::map(four_spp_list,calc_depth_frac,start_year=2020,end_year=2100) %>% bind_rows(.id='species')

fourspp_depth_plot <- fourspp_depth %>% 
    mutate(spp_plotting=case_when(
      species=='dover' ~ "Dover Sole",
      species=="ls" ~ "Longspine",
      species=="ss"~"Shortspine",
      species=="sable" ~"Sablefish"
    )) %>% 
    mutate(spp_plotting=factor(spp_plotting,levels=c("Dover Sole","Sablefish","Shortspine","Longspine"))) %>% 
  mutate(ESM=toupper(esm)) %>% 
  group_by(species,year,esm) %>% 
  slice_sample(n=25) %>% 
  ungroup() %>% 
  ggplot(aes(x=year,y=frac,col=spp_plotting,fill=spp_plotting))+
  geom_point(alpha=0.2,size=0.7)+
  geom_smooth()+
  scale_color_manual(values=pal4)+
  scale_fill_manual(values=pal4)+
  facet_wrap(~ESM,nrow=1)+
  # scale_color_manual(values=c("#2271B2","#d55e00"))+
  labs(x="Year",y="Proportion < 700 fathoms",title="",fill="Species",col="Species")+
  theme(axis.text.x=element_text(size=10),
        axis.text.y=element_text(size=10),
        panel.background = element_rect(fill=NA,color='black'),
        legend.position="bottom")
fourspp_depth_plot
```

```{r}
ggsave(here('model output','dts paper','fig_depth_change_four_spp.png'),fourspp_depth_plot,w=6,h=4)
```


```{r,fig.width=7,fig.height=5}
leg <- get_legend(fourspp_depth_plot)
fig_dist_depth_change<-plot_grid(four_spp_dist_to_shore_change+theme(plot.margin=unit(c(0.5,0.5,0.1,0.5),'cm'))+guides(color='none',fill='none'),
                                 fourspp_depth_plot + guides(color='none',fill='none') + theme(plot.margin=unit(c(0,0.5,1.5,0.5),'cm')),
                                 nrow=2,labels='auto')
fig_dist_depth_change<-ggdraw()+
  draw_plot(fig_dist_depth_change)+
  draw_plot(leg,y=-0.45)

ggsave(here('model output','dts paper','fig_dist_depth_change.png'),fig_dist_depth_change,w=8,h=7.5)
```

## Figure 3

Environmental state space within fishing footprints

```{r}
roms_env_space <- roms %>% 
  dplyr::select(year,lat,lon,mean_bt_30d_ipsl,mean_oxy_bottom_30d_ipsl) %>% 
  distinct() %>% 
  left_join(roms_ll,by=c('lon','lat')) %>% 
  drop_na() %>% 
  rename(t=mean_bt_30d_ipsl,oxy=mean_oxy_bottom_30d_ipsl) %>% 
  mutate(depth_bin=cut(depth_m,breaks=c(0,-100,-250,-500,-1000,-3000),
                       labels=c("1000-3000m","500-1000m","250-500m","100-250m","0-100m")),
         lat_bin=cut(lat,breaks=c(30,33,36,39,42,45,48),labels=c("30-33","33-36","36-39","39-42","42-45","45-48")))
roms_env_space_hist <- roms_env_space %>%
  # historical "mean" is 1985-2010
  filter(year%in%c(1985:2010)) %>%
  group_by(lat,lon,roms_cell,depth_m,depth_bin,lat_bin) %>% 
  summarise(t_hist=mean(t,na.rm=T),oxy_hist=mean(oxy,na.rm=T)) %>% 
  ungroup()

roms_env_space_fut <- roms_env_space %>%
  # historical "mean" is 1985-2010
  filter(year%in%c(2075:2100)) %>%
  group_by(lat,lon,roms_cell,depth_m,depth_bin,lat_bin) %>% 
  summarise(t_fut=mean(t,na.rm=T),oxy_fut=mean(oxy,na.rm=T)) %>% 
  ungroup()

roms_env_space_sf <- roms_env_space_hist %>% 
  left_join(roms_env_space_fut, c("lat", "lon", "roms_cell", "depth_m", "depth_bin", "lat_bin")) %>%
  st_as_sf(coords=c('lon','lat'),crs=4326) %>% 
  st_transform(st_crs(footprints)) %>% 
  st_join(footprints)

footprint_env_space <- roms_env_space_sf %>% 
  drop_na() %>% 
  group_by(port_name,port_group) %>% 
  summarise(mean_t_hist=mean(t_hist),mean_oxy_hist=mean(oxy_hist),
            mean_t_fut=mean(t_fut),mean_oxy_fut=mean(oxy_fut)) %>% 
  ungroup()
```

Make environmental affinities ellipses

```{r}
library(ggnewscale)
aff_polys<-affinities_ellipses(four_spp,yr_vec=1985:2010,concavity=7,threshold=0.25)
# plotting order for colors
polys_and_footprints <- ggplot()+
  geom_polygon(data=aff_polys,aes(V1,V2,col=spp_plotting),fill=NA,size=1,linetype=3)+
  scale_color_manual(values=c(pal4),name="Top 75% of\nspecies' CPUE")+
  new_scale_color()+
  geom_segment(data=footprint_env_space,aes(x=mean_t_hist,y=mean_oxy_hist,xend=mean_t_fut,yend=mean_oxy_fut,
             col=port_name),size=1,linetype=1,arrow=ggplot2::arrow(type="open",length=unit(0.1,'inches')))+
  scale_color_npg(name="Port Group\nClimate")+
  labs(x="Bottom Temperature (deg C)",y="Bottom Oxygen (mmol/m3)")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor=element_blank(),
        panel.border = element_rect(color='black',fill=NA),
        axis.text=element_text(size=12))
```

```{r,fig.width=8,fig.height=8}
# attach dumbbell plots
footprint_db_temp_plot2 <- footprint_env_space %>% 
  ggplot(aes(x=mean_t_hist,xend=mean_t_fut,y=port_name))+
  geom_dumbbell(color="#FCAE91",colour_x =NA,colour_xend = "#A50F15",
                size=1,size_x=3,size_xend = 3) +
  labs(x="Temperature",y="")+
  coord_flip()+
  theme(panel.border = element_rect(color='black',fill=NA),
        axis.text.x=element_text(color=pal_npg()(6),size=10))

footprint_db_oxy_plot2 <- footprint_env_space %>% 
  ggplot(aes(x=mean_oxy_hist,xend=mean_oxy_fut,y=port_name))+
  geom_dumbbell(color="#a3c4dc",colour_x =NA,colour_xend = "#0e668b",
                size=1,size_x=3,size_xend = 3) +
  labs(x="Oxygen",y="")+
  coord_flip()+
  theme(axis.text.x = element_blank(),panel.border = element_rect(color='black',fill=NA))

fig3 <- plot_grid(footprint_db_temp_plot2,footprint_db_oxy_plot2,polys_and_footprints,labels='auto',nrow=3,
                  hjust = 0,vjust=1,
                  rel_heights = c(0.2,0.2,0.6))
fig3
ggsave(here('model output','dts paper','fig_env_change_env_affs.png'),fig3,h=8,w=8)
```

## Figure 4

Calculate GIC for each species combination, year, simulation, and ESM

```{r}
tic("Ran GIC for sims")
dover_sable_gic<-crossing(gcm=c('gfdl','ipsl','hadl'),simnum=1:100) %>%
  mutate(gic=purrr::pmap(list(gcm,simnum),gic_sims,sims_df1=dover_ens_separate,sims_df2=sable_ens_separate))

dover_ss_gic<-crossing(gcm=c('gfdl','ipsl','hadl'),simnum=1:100) %>%
  mutate(gic=purrr::pmap(list(gcm,simnum),gic_sims,sims_df1=dover_ens_separate,sims_df2=ss_ens_separate))

dover_ls_gic<-crossing(gcm=c('gfdl','ipsl','hadl'),simnum=1:100) %>%
  mutate(gic=purrr::pmap(list(gcm,simnum),gic_sims,sims_df1=dover_ens_separate,sims_df2=ls_ens_separate))

sable_ss_gic<-crossing(gcm=c('gfdl','ipsl','hadl'),simnum=1:100) %>%
  mutate(gic=purrr::pmap(list(gcm,simnum),gic_sims,sims_df1=sable_ens_separate,sims_df2=ss_ens_separate))

sable_ls_gic<-crossing(gcm=c('gfdl','ipsl','hadl'),simnum=1:100) %>%
  mutate(gic=purrr::pmap(list(gcm,simnum),gic_sims,sims_df1=sable_ens_separate,sims_df2=ls_ens_separate))

ss_ls_gic<-crossing(gcm=c('gfdl','ipsl','hadl'),simnum=1:100) %>%
  mutate(gic=purrr::pmap(list(gcm,simnum),gic_sims,sims_df1=ss_ens_separate,sims_df2=ls_ens_separate))
toc()
```
Calculate LIC and Schoener's D.

```{r}
tic("Ran LIC and D for sims")
dover_sable_lic <- lic_D_sims(dover_ens_separate,sable_ens_separate)
dover_ss_lic <- lic_D_sims(dover_ens_separate,ss_ens_separate)
dover_ls_lic <- lic_D_sims(dover_ens_separate,ls_ens_separate)
sable_ss_lic <- lic_D_sims(sable_ens_separate,ss_ens_separate)
sable_ls_lic <- lic_D_sims(sable_ens_separate,ls_ens_separate)
ss_ls_lic <- lic_D_sims(ss_ens_separate,ls_ens_separate)
toc()
```
Join and plot

```{r}
dover_sable_overlap <- dover_sable_gic %>% 
  unnest(cols=c(gic)) %>% 
  mutate(year=rep(1980:2100,300)) %>% 
  left_join(dover_sable_lic %>% mutate(simnum=str_remove(sim,'sim') %>% as.numeric()),
            by=c('gcm'='esm','simnum','year'))
dover_ss_overlap <- dover_ss_gic %>% 
  unnest(cols=c(gic)) %>% 
  mutate(year=rep(1980:2100,300)) %>% 
  left_join(dover_ss_lic %>% mutate(simnum=str_remove(sim,'sim') %>% as.numeric()),
            by=c('gcm'='esm','simnum','year'))
dover_ls_overlap <- dover_ls_gic %>% 
  unnest(cols=c(gic)) %>% 
  mutate(year=rep(1980:2100,300)) %>% 
  left_join(dover_ls_lic %>% mutate(simnum=str_remove(sim,'sim') %>% as.numeric()),
            by=c('gcm'='esm','simnum','year'))
sable_ss_overlap <- sable_ss_gic %>% 
  unnest(cols=c(gic)) %>% 
  mutate(year=rep(1980:2100,300)) %>% 
  left_join(sable_ss_lic %>% mutate(simnum=str_remove(sim,'sim') %>% as.numeric()),
            by=c('gcm'='esm','simnum','year'))
sable_ls_overlap <- sable_ls_gic %>% 
  unnest(cols=c(gic)) %>% 
  mutate(year=rep(1980:2100,300)) %>% 
  left_join(sable_ls_lic %>% mutate(simnum=str_remove(sim,'sim') %>% as.numeric()),
            by=c('gcm'='esm','simnum','year'))
ss_ls_overlap <- ss_ls_gic %>% 
  unnest(cols=c(gic)) %>% 
  mutate(year=rep(1980:2100,300)) %>% 
  left_join(ss_ls_lic %>% mutate(simnum=str_remove(sim,'sim') %>% as.numeric()),
            by=c('gcm'='esm','simnum','year'))
```

```{r}
# make a good color palette
# 3 sets of 3 colors for the 3 ESMs and 3 metrics
#"#790149","#9F0162","#EF0096",
overlaps_cols <- c("#005745","#009175","#00CBA7",
                   "#7E0018","#CD022D","#FF6E3A",
                   "#00306F","#005FCC","#00C2F9")
overlap_plot_foo <- function(overlap_df,show_legend=FALSE){
  p <- overlap_df %>% 
    filter(year>2019) %>% 
    pivot_longer(cols=c('gic','lic','D'),names_to='metric',values_to='value') %>% 
    unite(esm_metric,gcm,metric,remove=F) %>% 
    unite(sim_esm_metric,simnum,esm_metric,remove=F) %>% 
    ggplot(aes(year,value,color=esm_metric))+
    # geom_point(alpha=0.2,size=0.7)+
    geom_line(aes(group=sim_esm_metric),color='gray80')+
    geom_smooth(se=F)+
    scale_color_manual(values=overlaps_cols)+
    labs(x="Year",y="Overlap",color="Metric")+
    theme(panel.background = element_rect(color="black",fill=NA),
          legend.position = 'bottom')
  if(!show_legend) p+guides(color='none')
  else p
}

dover_sable_overlap_plot <- overlap_plot_foo(dover_sable_overlap,show_legend = T)
dover_ss_overlap_plot <- overlap_plot_foo(dover_ss_overlap)
dover_ls_overlap_plot <- overlap_plot_foo(dover_ls_overlap)
sable_ss_overlap_plot <- overlap_plot_foo(sable_ss_overlap)
sable_ls_overlap_plot <- overlap_plot_foo(sable_ls_overlap)
ss_ls_overlap_plot <- overlap_plot_foo(ss_ls_overlap)
overlaps_legend <- get_legend(dover_sable_overlap_plot)

r1 <- plot_grid(dover_sable_overlap_plot+guides(color='none'),sable_ss_overlap_plot,sable_ls_overlap_plot,dover_ss_overlap_plot,dover_ls_overlap_plot,ss_ls_overlap_plot,nrow=2,labels='auto')
r2=plot_grid(NULL,NULL,overlaps_legend,NULL,NULL,nrow=1)

overlaps_all_fig4 <- plot_grid(r1,r2,nrow=2,rel_heights = c(1,0.2))

ggsave(here('model output','dts paper','fig_overlaps_all.png'),overlaps_all_fig4,h=8,w=8)
```

## Figure 5

```{r,fig.height=8,fig.width=8}
dover_sable_lic_D <- LIC_D_prop_change(dover_ens_separate,sable_ens_separate)
dover_ss_lic_D <- LIC_D_prop_change(dover_ens_separate,ss_ens_separate)
dover_ls_lic_D <- LIC_D_prop_change(dover_ens_separate,ls_ens_separate)
sable_ss_lic_D <- LIC_D_prop_change(sable_ens_separate,ss_ens_separate)
sable_ls_lic_D <- LIC_D_prop_change(sable_ens_separate,ls_ens_separate)
ss_ls_lic_D <- LIC_D_prop_change(ss_ens_separate,ls_ens_separate)

# together
LIC_D_fps_all <- dover_sable_lic_D %>% mutate(spp="Sablefish/Dover Sole") %>% 
  bind_rows(dover_ss_lic_D %>% mutate(spp="Dover Sole/Shortspine")) %>%  
  bind_rows(dover_ls_lic_D %>% mutate(spp="Dover Sole/Longspine")) %>%  
  bind_rows(sable_ss_lic_D %>% mutate(spp="Sablefish/Shortspine")) %>%  
  bind_rows(sable_ls_lic_D %>% mutate(spp="Sablefish/Longspine")) %>%  
  bind_rows(ss_ls_lic_D %>% mutate(spp="Shortspine/Longspine")) %>% 
  mutate(spp=factor(spp,levels=c("Sablefish/Dover Sole","Sablefish/Shortspine","Sablefish/Longspine",
                                 "Dover Sole/Shortspine","Dover Sole/Longspine","Shortspine/Longspine")))

D_fps_plot <- LIC_D_fps_all %>% 
  mutate(ESM=toupper(esm)) %>% 
  ggplot(aes(year,fct_rev(spp),fill=D_prop/100,col=D_prop/100))+
  geom_tile()+
  # geom_point(size=0.2)+
  # geom_smooth(se=F)+
  facet_grid(port_name~ESM)+
  scale_x_continuous(expand=c(0,0))+
  # scale_y_continuous(limits=c(0,1))+
  scale_fill_gradient2(midpoint=0.5)+
  scale_color_gradient2(midpoint=0.5)+
  # geom_hline(yintercept=0.5,linetype=2)+
  labs(x="Year",y="Species Pair",fill="Proportion of Simulations\nwith Increased Overlap",
       color="Proportion of Simulations\nwith Increased Overlap")+
  theme(panel.background = element_rect(fill=NA,color='black'),
        legend.position = 'bottom',
        axis.text.x = element_text(angle=45,vjust=0.8,hjust=0.8),
        legend.key.width = unit(0.75,'inches'),
        panel.spacing = unit(0.2,'inches'),
        strip.text = element_text(size=8),
        panel.border = element_rect(color='black',fill=NA,size=1.5))

D_fps_plot

ggsave(here('model output','dts paper','fig_overlaps_footprints.png'),D_fps_plot,w=8,h=8)
```

## Figure 6

Try a new way with smoothed histograms

```{r,fig.height=8,fig.width=8}
allspp_reldens <- purrr::map2_df(list(dover_ens_separate,sable_ens_separate,ss_ens_separate,ls_ens_separate),
                                 c('dover','sable','shortspine','longspine'),
                                 function(x,y){
                                   calc_footprint_reldens_sims(x) %>% mutate(spp=y)
                                 }) %>%
  mutate(ESM=toupper(esm)) %>% 
  mutate(spp_plotting=case_when(
      spp=='dover' ~ "Dover Sole",
      spp=="longspine" ~ "Longspine",
      spp=="shortspine"~"Shortspine",
      spp=="sable" ~"Sablefish"
    )) %>% 
  mutate(spp_plotting=factor(spp_plotting,levels=c("Sablefish","Dover Sole","Shortspine","Longspine"))) %>% 
  # mutate(delta_perc=ifelse(delta_perc>200|delta_perc < -200,NA,delta_perc)) %>% 
  drop_na()

reldens_fig <- allspp_reldens %>% 
  dplyr::select(sim,ESM,spp_plotting,delta_perc,port_name) %>% 
  unite(fct,spp_plotting,port_name,sep="\n",remove=F) %>% 
  mutate(fct=factor(fct,levels=unique(fct))) %>% 
  ggplot(aes(delta_perc,fill=ESM))+
  geom_density(alpha=0.5)+
  geom_vline(xintercept=0,linetype=2)+
  facet_wrap(~fct,scales='free',ncol=6)+
  # xlim(-200,200)+
  scale_color_manual(values=viridis_pal(option="D",begin=0.2,end=0.8)(3))+
  scale_fill_manual(values=viridis_pal(option="D",begin=0.2,end=0.8)(3))+
  labs(x="Percent Change in CPUE",y="",fill="CCROMS Model")+
  theme(panel.background = element_rect(fill=NA,color='black'),
        axis.text.y=element_blank(),
        axis.text.x=element_text(angle=45,vjust=0.8,hjust=0.8),
        axis.ticks.x = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor=element_blank())


  # nest(dat=c(sim,ESM,port_name,delta_perc)) %>% 
  # mutate(p=purrr::map(dat,function(d){
  #     d %>%
  #     ggplot(aes(delta_perc,fill=ESM))+
  #     geom_density(alpha=0.5)+
  #     geom_vline(xintercept=0,linetype=2)+
  #     facet_wrap(~port_name,scales='free_y',nrow=1)+
  #     # xlim(-200,200)+
  #     scale_color_npg()+
  #     scale_fill_npg()+
  #     labs(x="Percent Change in CPUE",y="",fill="CCROMS Model")+
  #     theme(panel.background = element_rect(fill=NA,color='black'),
  #           axis.text.y=element_blank(),
  #           panel.grid.major = element_blank(),
  #           panel.grid.minor=element_blank())
  # }))

# plot_grid(plotlist = reldens_fig$p,nrow=4)

reldens_fig
ggsave(here('model output','dts paper','fig_cpue_change_footprints.png'),reldens_fig,w=8,h=8)
```

