---
title: "Organize Habitat Data"
author: "Owen Liu"
date: "6/24/2021"
output: html_document
---

```{r setup, include=FALSE}
# devtools::install_github("pbs-assess/sdmTMB")
library(raster)
library(rasterVis)
library(tidync)
library(tidyverse)
library(lubridate)
library(sf)
library(here)
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)
```


**NOTE: THIS NEEDS TO BE UPDATED IF WE ARE GOING TO USE THE SUBSTRATE DATA, BECAUSE THE DEPTH LAYER HAS CHANGED, APRIL 2022**

# Purpose

Organize and interpolate habitat (substrate) data from Joe Bizarro, Blake Feist, and Lewis Barnett into habitat variables that can be used in sdmTMB models.

# Data Structure

There are two versions of substrate data. One is a 25m by 25m grid, where each grid cell is a 1, 2, or 3 denoting hard, mixed, or soft substrate, respectively.

The second version was put together by Blake Feist (README is in this repo), and uses the actual trawl tow lines to measure the proportion of each towline that traverses each type of substrate.

For now we will organie both. Version 1 will be used in SDM projections, but summarizing it to the resolution of the ROMS grid, while Version 2 will be used in SDM model fitting.

# Organize Version 1

Import version 1 of the substrate data, and summarize onto the ROMS grid. Calculate the proportion of each ROMS cell covered by each substrate type.

## Import Data

```{r}
# version 1, converted to a point feature
habdat <- raster(here('data','substrate','bef','wc_sub_hms_i_UTM.tif'))

# ROMS depths, from Mike Jacox email June 11, 2021
# CHANGED APRIL 2022, code will break from here
roms_depths <- tidync(here('data','grids','bottom_layer_depth.nc')) %>% hyper_tibble() %>% 
  mutate(depth_m=-bld)

roms_ll_key <- roms_depths %>% arrange(desc(lat),lon) %>% 
  mutate(roms_cell=row_number()) %>% 
  dplyr::select(roms_cell,lon,lat,depth_m)

write_rds(roms_ll_key,here::here('data','roms_latlon_depth_key_topleft_start.rds'))
```

```{r}
res_roms <- 0.1
roms_lat <- unique(roms_depths$lat)
roms_lon <- unique(roms_depths$lon)
# make the raster, with small adjustment because R usually interprets grid corners, not grid centers
roms_grid <- raster(nrows=length(roms_lat),ncols=length(roms_lon),
                    xmn=min(roms_lon)-res_roms/2,xmx=max(roms_lon)+res_roms/2,
                    ymn=min(roms_lat)-res_roms/2,ymx=max(roms_lat)+res_roms/2,crs=4326,
                    # vals=roms_depths$depth_m)
                    # vals=roms_depths %>% arrange(lon,desc(lat)) %>% pull('depth_m'))
                    vals=roms_depths %>% arrange(desc(lat),lon) %>% pull('depth_m'))

# ROMS grid, converted to polygons
# roms_grid <- raster(here::here('data','substrate','bottom_layer_depth.grd'))
roms_cellnums <- setValues(roms_grid,1:length(roms_grid))
roms_sf <- rasterToPolygons(roms_cellnums) %>% st_as_sf()
```

```{r}
# cell xy
# cellxy <- xyFromCell(roms_grid,cell=1:length(roms_grid))
# #save a ROMS latitude/longitude key
# #we want to save the upper left corner of the cell, not the centroid
# roms_ll_key <- tibble(roms_cell=1:length(roms_grid),lon=cellxy[,1]-0.05,lat=cellxy[,2]+0.05,depth_m=getValues(roms_grid)*-1)
# 
# roms_ll_key %>% ggplot(aes(lon,lat,col=depth_m))+geom_point()
# # save the key
# # write_rds(roms_ll_key,here::here('data','roms_latlon_depth_key_topleft_start.rds'))
```


## Transform and Resample ROMS data

Spatially transform the ROMS grid into the same UTM projection as the substrate raster. We do this in sort of a weird way by rasterizing the ROMS sf object.

```{r}
roms_utm <- roms_sf %>% st_transform(crs(habdat)) %>% 
  # rasterize
  fasterize::fasterize(habdat,field='layer')
roms_utm
```

Do a cross-tabulation, counting the overlap of cells by type between the ROMS and habitat layers:

```{r}
crosstab <- crosstab(roms_utm,habdat,long=T,progress='text')
```

Finally, summarize the crosstabulation to obtain the proportion of each habitat type in each represented ROMS cell. At this step, we combine habitat categories 1 and 2 (hard and mixed substrate) into one category.

```{r}
prop_hab_ROMS <- crosstab %>% 
  rename(substrate=wc_sub_hms_i_UTM,
         roms_cell=layer) %>% 
  group_by(roms_cell) %>% 
  mutate(totcells=sum(Freq)) %>% 
  ungroup() %>%
  mutate(substrate2=ifelse(substrate==3,"soft","hard")) %>%
  mutate(prop=Freq/totcells) %>% 
  group_by(roms_cell,substrate2) %>% 
  summarise(prop=sum(prop)) %>% 
  ungroup() %>% 
  # add one last variable for just "proportion hard substrate"
  mutate(prop_hard=ifelse(substrate2=="soft",1-prop,prop)) %>% 
  mutate(prop_hard=round(prop_hard,digits=5))

glimpse(prop_hab_ROMS)
```

## Visualize

Re-attach the proportion of hard habitat to the original ROMS grid, then plot.

```{r}
prop_hard_ROMS <- prop_hab_ROMS %>% distinct(roms_cell,prop_hard)
glimpse(prop_hard_ROMS)
roms_prop_hard_grid <- roms_cellnums
values(roms_prop_hard_grid) <- NA
roms_prop_hard_grid[prop_hard_ROMS$roms_cell] <- prop_hard_ROMS$prop_hard

levelplot(roms_prop_hard_grid)
```

# Organize Version 2

```{r}
# version 2
habdat2 <- read_tsv(here::here('data','substrate','bef','Haul line substrate classification based on Bizzarro & Yoklavich.txt'))
glimpse(habdat2)
```

For this dataset, all we do is sum up the proportion of hard and mixed into one category.

```{r}
habdat2 <- habdat2 %>% 
  mutate(prop_hard_mixed=Proportion_hard+Proportion_mixed)
```

# Save

```{r}
write_rds(prop_hard_ROMS,here('data','substrate','prop_hard_by_ROMS_cell.rds'))
write_rds(habdat2,here('data','substrate','substrate_by_trawlID.rds'))
```

