---
title: "ROMS hypervolumes"
author: "Owen Liu"
date: '2022-05-26'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(hypervolume)
library(terra)
library(here)
library(tictoc)
library(viridis)
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=12,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

```{r}
#SDM projection extent, to mask the rasters
proj_ext <- read_rds(here::here('data','cropped_domain_for_projected_sdms.rds')) %>% vect()
```


```{r}
## NOTE: SEEING IF HYPERVOLUME CODE WILL RUN FASTER IF WE REPLACE SOME FUNCTIONS WITH TERRA INSTEAD OF RASTER:
hypervolume_project <- function (hv, rasters, type = "probability", verbose = TRUE,...){
  raster.values = data.frame(terra::values(rasters))
  na.rows <- raster.values %>% rowid_to_column("idx") %>% filter(if_any(everything(),is.na)) %>% pull(idx)
  if (type == "probability") {
    projected.values <- hypervolume_estimate_probability(hv = hv, 
                                                         points = raster.values, verbose = verbose, ...)
  }
  else if (type == "inclusion") {
    projected.values = hypervolume_inclusion_test(hv, raster.values, 
                                                  verbose = verbose, ...)
  }
  else {
    stop("Unsupported 'type' argument")
  }
  
  raster.out <- rasters[[1]]
  if (length(attr(projected.values, "NAIdx")) > 0) {
    terra::values(raster.out)[-na.rows] <- projected.values
    terra::values(raster.out)[na.rows] <- NA
  }
  else {
    terra::values(raster.out) <- projected.values
  }
  return(raster.out)
}
```


# Purpose

Using code modified from James Smith, calculate hypervolumes and novelty for the 3 ROMS-ESM models for bottom temperature and bottom oxygen in the California Current. For historical conditions, use 1980-2010 (ROMS hindcast period), then apply to 2011 - 2100

First, we do some organizing of the data.

# Organize ROMS data

We need to organize the historical and projected ROMS data into data frames and rasters for fitting of hypervolumes

Could do this with a masked projection area (e.g., trawl survey footprint)

```{r}
roms_fls <- list.files(here('data','hypervolumes','projection rasters'),full.names=T)

# for each ESM type, make a raster stack for the appropriate years
gfdl_bt_historical <- roms_fls %>% str_subset('gfdl_bt') %>% str_subset('.grd') %>% rast() %>% subset(1:31) %>% mask(proj_ext)
ipsl_bt_historical <- roms_fls %>% str_subset('ipsl_bt') %>% str_subset('.grd') %>% rast() %>% subset(1:31)%>% mask(proj_ext)
had_bt_historical <- roms_fls %>% str_subset('had_bt') %>% str_subset('.grd') %>% rast() %>% subset(1:31)%>% mask(proj_ext)
gfdl_oxy_historical <- roms_fls %>% str_subset('gfdl_oxy') %>% str_subset('.grd') %>% rast() %>% subset(1:31)%>% mask(proj_ext)
ipsl_oxy_historical <- roms_fls %>% str_subset('ipsl_oxy') %>% str_subset('.grd') %>% rast() %>% subset(1:31)%>% mask(proj_ext)
had_oxy_historical <- roms_fls %>% str_subset('had_oxy') %>% str_subset('.grd') %>% rast() %>% subset(1:31)%>% mask(proj_ext)

# for each ESM type, make a raster stack for the appropriate years
gfdl_bt_future <- roms_fls %>% str_subset('gfdl_bt') %>% str_subset('.grd') %>% rast() %>% subset(32:121)%>% mask(proj_ext)
ipsl_bt_future <- roms_fls %>% str_subset('ipsl_bt') %>% str_subset('.grd') %>% rast() %>% subset(32:121)%>% mask(proj_ext)
had_bt_future <- roms_fls %>% str_subset('had_bt') %>% str_subset('.grd') %>% rast() %>% subset(32:121)%>% mask(proj_ext)
gfdl_oxy_future <- roms_fls %>% str_subset('gfdl_oxy') %>% str_subset('.grd') %>% rast() %>% subset(32:121)%>% mask(proj_ext)
ipsl_oxy_future <- roms_fls %>% str_subset('ipsl_oxy') %>% str_subset('.grd') %>% rast() %>% subset(32:121)%>% mask(proj_ext)
had_oxy_future <- roms_fls %>% str_subset('had_oxy') %>% str_subset('.grd') %>% rast() %>% subset(32:121)%>% mask(proj_ext)

latlon <- xyFromCell(rast(roms_fls[1]), 1:ncell(rast(roms_fls[1]))) %>% as_tibble() %>%
  set_names(c('lon','lat'))
```

Re-cast rasters to dataframes

```{r}
recast_stack <- function(rs,varname){
  # yrs=1979+(1:nlyr(rs))
  out <- rs %>%
    terra::values() %>% 
    as_tibble(.name_repair = "unique") %>% 
    pivot_longer(everything(),names_to = 'year',values_to = varname) %>% 
    dplyr::select(2) #pull out the value variable
    # mutate(year=rep(yrs,ncell(rs)))
  out
}

historical_all <- map2(.x=list(gfdl_bt_historical,ipsl_bt_historical,had_bt_historical,
                               gfdl_oxy_historical,ipsl_oxy_historical,had_oxy_historical),
                       .y=c("gfdl_bt","ipsl_bt","had_bt","gfdl_oxy","ipsl_oxy","had_oxy"),
                       recast_stack) %>% bind_cols() %>% 
  # add year and lat/lons
  mutate(year=rep(1980:2010,ncell(gfdl_bt_historical))) %>% 
  mutate(lat=rep(latlon$lat,each=length(1980:2010)),
         lon=rep(latlon$lon,each=length(1980:2010)))
```

Normalize values

```{r}
env_vars <- setdiff(names(historical_all),c("year","lat","lon"))

means <- historical_all %>% 
  summarise(across(all_of(env_vars),mean,na.rm=T)) #'global' values used to standardise all your data

SDs <- historical_all %>% 
  summarise(across(all_of(env_vars),sd,na.rm=T))

hist_data_s <- historical_all %>% 
  dplyr::select(all_of(env_vars)) %>% 
  scale(center=means,scale=SDs) %>% as_tibble()

# without NAs
hist_data_s_complete <- hist_data_s %>% drop_na()
```

# Make Hypervolumes

Make hypervolumes for each ESM for bottom temperature and bottom oxygen

```{r}
calc_hv <- function(df,esm,max_sample=1e5){
  dfsub <- df %>% dplyr::select(contains(esm))
  set.seed(123)
  if(nrow(dfsub)>max_sample){
    dfsub <- dfsub %>% slice_sample(n=max_sample)
  }
  tic("Calculated hypervolume.")
  hv <- hypervolume(data=dfsub,method='svm')
  toc()
  return(hv)
}

gfdl_hv <- calc_hv(hist_data_s_complete,esm="gfdl")
ipsl_hv <- calc_hv(hist_data_s_complete,esm='ipsl')
had_hv <- calc_hv(hist_data_s_complete,esm='had')

# gut checks- took 76 minutes to run this with the full data (~1m rows)
# try with a subset?
# 100k sample took 35s
# 300k sample took 466s
# maybe should consider 100k subset, but with replicates??
```

Pair plots of environmental niche; dark red = data subset; small red = random points inside hypervolume

```{r}
plot(ipsl_hv)
plot(gfdl_hv)
plot(had_hv)
```


# Calculate Novelty

Using the hypervolumes created above, calculate novelty for future scenarios

```{r}
# library(future)
# library(furrr)
# future::plan(sequential)
# future_rasters should be a stack of temperature and oxygen rasters that you want to calculate novelty for (here, 2011-2100)
calc_novelty_ts <- function(hv,esm,future_bt,future_oxy){
  # stack temperature and oxygen for each year
  tic("Splitting rasters.")
  yrslist <- map(2011:2100,function(y){
    n=paste0('year',y)
    ind=y-2010
    rx=c(future_bt[[ind]],future_oxy[[ind]])
    names(rx)=c('bt','oxy')
    # scale using global means
    rx$bt <- scale(rx[[1]],center=pull(means,paste0(esm,'_bt')),scale = pull(SDs,paste0(esm,'_bt')))
    rx$oxy <- scale(rx[[2]],center=pull(means,paste0(esm,'_oxy')),scale = pull(SDs,paste0(esm,'_oxy')))
    rx
  })
  toc()
  tic("Running inclusion test on each year.")
  # for each year, calculate inclusion of these future raster values in historical hypervolume
  out <- map(yrslist,function(x){
    hp <- hypervolume_project(hv, rasters=x,
                                type="inclusion",
                                fast.or.accurate="accurate")
    excl <- sum(terra::values(hp)==0,na.rm = T)  #number of cells 'excluded';
    excl_prop <- excl/sum(terra::values(hp)>=0,na.rm=T) #proportion of cells excluded;
    return(list(excl=excl,prop=excl_prop,rast=hp))
  })
  toc()
  return(out)
}
```

```{r}
# testing
# future_bt=gfdl_bt_future
# future_oxy=gfdl_oxy_future
# esm='gfdl'
# yrslist <- map(2011:2100,function(y){
#     n=paste0('year',y)
#     ind=y-2010
#     rx=c(future_bt[[ind]],future_oxy[[ind]])
#     names(rx)=c('bt','oxy')
#     # scale using global means
#     rx$bt <- scale(rx[[1]],center=pull(means,paste0(esm,'_bt')),scale = pull(SDs,paste0(esm,'_bt')))
#     rx$oxy <- scale(rx[[2]],center=pull(means,paste0(esm,'_oxy')),scale = pull(SDs,paste0(esm,'_oxy')))
#     rx
#   })
# 
# xy <- values(yrslist[[1]]) %>% as_tibble() %>% slice_sample(n=10000)
# xy %>% ggplot(aes(oxy,bt))+geom_point()
# 
# hp <- hypervolume_project(gfdl_hv, rasters=yrslist[[75]],
#                                 type="inclusion",
#                                 fast.or.accurate="accurate")
# excl <- sum(terra::values(hp)==0,na.rm = T)  #number of cells 'excluded';
# excl_prop <- excl/sum(terra::values(hp)>=0,na.rm=T) #proportion of cells excluded;
# rm(excl,excl_prop,xy,future_bt,future_oxy,hp,esm)
```


```{r}
gfdl_proj <- calc_novelty_ts(hv=gfdl_hv,esm='gfdl',future_bt = gfdl_bt_future,future_oxy = gfdl_oxy_future)
ipsl_proj <- calc_novelty_ts(hv=ipsl_hv,esm='ipsl',future_bt = ipsl_bt_future,future_oxy = ipsl_oxy_future)
had_proj <- calc_novelty_ts(hv=had_hv,esm='had',future_bt = had_bt_future,future_oxy = had_oxy_future)
# benchmarking- this took about 3.5 minutes with the masked data
```

# Visualize

```{r}
inclusion_ts <- function(hvout,plot=F){
  excl_ts <- tibble(yr=2011:2100,
                    excl=purrr::map_dbl(hvout,function(x)pluck(x,'excl')),
                    prop=purrr::map_dbl(hvout,function(x)pluck(x,'prop')))
  
  if(plot) {
    out <- excl_ts %>% 
      ggplot(aes(yr,prop))+
      geom_line()+
      labs(x="Year",y="Proportion Novel")
  }
  else{out<-excl_ts}

  return(out)
}

library(animation)

animate_ts <- function(hvout,fname){
  r <- purrr::map(hvout,function(x)pluck(x,'rast')) %>%
    reduce(c)
  animation::saveGIF(terra::animate(r,main=as.character(2011:2100),n=1),movie.name =paste0(fname,'.gif'))
}
```
Apply

```{r}
gfdl_ts <- inclusion_ts(gfdl_proj,plot=F) %>% mutate(esm='gfdl')
ipsl_ts <- inclusion_ts(ipsl_proj,plot=F) %>% mutate(esm='ipsl')
had_ts <- inclusion_ts(had_proj,plot=F) %>% mutate(esm='hadley')
```
# Save
```{r}
novelty_3ESMs <- bind_rows(gfdl_ts,ipsl_ts,had_ts) %>% 
  ggplot(aes(yr,prop,col=esm))+
  geom_line(size=1.5)+
  ylim(0,0.6)+
  labs(x="Year",y="Proportion Novel")

novelty_3ESMs

ggsave(here('plots','novelty_3ESMs.png'),novelty_3ESMs,w=7,h=4)
#gifs

animate_ts(gfdl_proj,fname=here::here('plots','gfdl_novelty'))
animate_ts(ipsl_proj,fname=here::here('plots','ipsl_novelty'))
animate_ts(had_proj,fname=here::here('plots','had_novelty'))
```



