---
title: "Atlantis SDMs with GLORYS"
author: "Owen Liu"
date: "1/4/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r, include=FALSE}
# devtools::install_github("pbs-assess/sdmTMB")
library(sdmTMB)
library(rbgm)
library(tidyverse)
library(lubridate)
library(rnaturalearth)
library(sf)
library(here)
library(cowplot)
library(RANN)
library(furrr)
library(future)
library(tictoc)

knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)
```

```{r}
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

```{r}
# set up parallel processing?
# future::plan(sequential)
# future::plan(multisession,workers=4)
```

```{r}
# Atlantis grid
fl <- here::here("data",'grids',"CalCurrentV3_utm.bgm")
# load the file
bgm <- read_bgm(fl)
names(bgm)
emocc_sf <- box_sf(bgm) %>% dplyr::select(box_id,label,area,botz) %>% 
  st_transform("+proj=utm +zone=10 +datum=WGS84 +units=km")
# areas of boxes
emocc_box_areas <- emocc_sf %>% st_set_geometry(NULL) %>% dplyr::select(-label)
```

```{r}
# GLORYS temp/oxygen data
glorys <- read_rds(here('data','GLORYS_bottom_temp_o2_SDMs.rds')) %>% 
  # convert to UTM zone 10
  st_as_sf(coords=c('longitude','latitude'),crs=4326) %>% 
  st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km") %>% 
  mutate(latitude = sf::st_coordinates(.)[,2],
         longitude = sf::st_coordinates(.)[,1])
```

```{r}
# spatial background map
# load west cost land for mapping
coast <- ne_countries(continent="North America",scale=50,returnclass = 'sf') %>% 
  st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km") %>% 
  st_crop(emocc_sf)
# coast <- ne_states(country='United States of America',returnclass = 'sf') %>% 
#   filter(name %in% c('California','Oregon','Washington','Nevada')) %>%
#   st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km")
```


# Purpose

Produce groundfish species distribution models for use in the new version of the Atlantis ecosystem model. We will use `sdmTMB` to fit SDMs to high-resolution ROMS and trawl survey data, establishing the environmental niches of groundfish relative to bottom temperature and oxygen. Then, we use a GLORYS/ESM hybrid oceanography product to project species distributions out to 2100 in the California Current. Finally, we will translate those SDMs on to the Atlantis polygon/grid structure.

Note: Model runs are described and performed in the `sdmTMB models.Rmd` script. Here, we apply the new GLORYS data for projection.

# Define Projection Extent

Here, we set the boundaries for the `sdmTMB` projection. We use three criteria: 1) we crop the GLORYS data to the extent of the Atlantis polygons; 2) for all species except lanternfish/myctophids, we want to we exclude offshore Atlantis polygons from the projection; 3) for lanternfish, we will keep all boxes.

```{r}
# trawl IDs and depths
trawls <- read_rds(here::here('data','trawl','trawlID.rds'))
max_depth <- max(trawls$depth_trawl,na.rm=T)

# glorys bbox with areas that are too deep removed
atlantis_glorys_extent <- glorys %>% 
  filter(depth<max_depth) %>% 
  st_bbox() %>% 
  st_as_sfc()

# atlantis polygons, summarized/unioned
emocc_merged <- emocc_sf %>% summarise()
emocc_no_offshore_merged <- emocc_sf %>% 
  filter(botz!=min(botz)) %>% 
  summarise()

# plot our starting point
ggplot()+
  geom_sf(data=coast)+
  geom_sf(data=atlantis_glorys_extent,fill='blue',alpha=0.5)+
  geom_sf(data=emocc_no_offshore_merged,fill='red',alpha=0.5)

# crop glorys bbox with atlantis polygons
ext <- atlantis_glorys_extent %>% 
  st_intersection(emocc_no_offshore_merged)
ext_offshore <- atlantis_glorys_extent %>% 
  st_intersection(emocc_merged)

# plot the final domain
ggplot()+
  geom_sf(data=coast)+
  geom_sf(data=ext_offshore,fill='yellow',alpha=0.5)+
  geom_sf(data=ext,fill='purple',alpha=0.5)
```

Use this projection extent to crop the GLORYS data used for projection

```{r}
glorys_crop <- glorys %>% 
  # crop to SDM extent
  st_intersection(ext_offshore) %>%
  st_set_geometry(NULL)
glorys_no_offshore_crop <- glorys %>% 
  # crop to SDM extent
  st_intersection(ext) %>%
  st_set_geometry(NULL)
```

# Load Fitted Models

Start with an example species: Dover sole. See `sdmTMB models.Rmd` for details of the ensemble modelling approach.

```{r,eval=F}
## dover sole models
dover_models <- read_rds(here::here('model output','dover sole no substrate.rds'))
names(dover_models)
```

# GLORYS Projection Function

```{r,eval=F}
make_predictions_glorys <- function(modelobj){

  original_model_data <- modelobj$data
  # need to use the original (hindcast) environmental data to scale the projected data
  mean_t <- mean(original_model_data$mean_temp_roms_30,na.rm=T)
  sd_t <- sd(original_model_data$mean_temp_roms_30,na.rm=T)
  mean_oxy <- mean(original_model_data$mean_oxygen_roms_30,na.rm=T)
  sd_oxy <- sd(original_model_data$mean_oxygen_roms_30,na.rm=T)
  
  # create a new tibble with the projected data from GLORYS
  newdata <- glorys_crop %>% 
    # No habitat data
    # left_join(hab,by=c("lat","lon")) %>% 
    # drop_na() %>% 
    dplyr::select(year,latitude,longitude,depth,temperature,o2)
  
  temperature <- newdata %>%
    # keep variable names from model fitting to ROMS
    mutate(mean_temp_roms_30_norm=(temperature-mean_t)/sd_t)
  
  oxygen <- newdata %>% 
    mutate(mean_oxygen_roms_30_norm=(o2-mean_oxy)/sd_oxy)

  newdata <- newdata %>% 
    left_join(temperature,by = c("year", "latitude", "longitude", "depth", "temperature", "o2")) %>% 
    left_join(oxygen,by = c("year", "latitude", "longitude", "depth", "temperature", "o2")) %>% 
    dplyr::select(-temperature,-o2) %>%
    mutate(year=as.double(year))
  
  # years to predict
  yrs <- sort(unique(newdata$year))
  # now we can make the predictions
  predicted_cpue_km2 <- predict(modelobj,newdata,return_tmb_object=T,extra_time=yrs)
  
  predicted_cpue_km2
}
```

Test it
```{r,eval=F}
dover_future1 <- make_predictions_glorys(dover_models %>% pluck('model',1))
glimpse(dover_future1$data)
```

# Ensemble Predictions Function

To create ensemble predictions, we calculate predictions for each species' model separately, then weight them using the weights established in model fitting (see `sdmTMB::sdmTMB_stacking()`).

```{r}
# this function is meant to be run with a list of models that all are fit on the same data, with a common set of predictors
ensemble_predictions <- function(model_df,use_offshore=F){
  
  # use the first model in the list to pull out the data
  original_model_data <- model_df %>% pluck('model',1,'data')
  # need to use the original (hindcast) environmental data to scale the projected data
  mean_t <- mean(original_model_data$mean_temp_roms_30,na.rm=T)
  sd_t <- sd(original_model_data$mean_temp_roms_30,na.rm=T)
  mean_oxy <- mean(original_model_data$mean_oxygen_roms_30,na.rm=T)
  sd_oxy <- sd(original_model_data$mean_oxygen_roms_30,na.rm=T)
  
  # create a new tibble with the projected data from GLORYS
  if(use_offshore){glorys_to_use <- glorys_crop}else{glorys_to_use <- glorys_no_offshore_crop}
  newdata <- glorys_to_use %>% 
    # No habitat data
    # left_join(hab,by=c("lat","lon")) %>% 
    # drop_na() %>% 
    dplyr::select(year,latitude,longitude,depth,temperature,o2)
  
  temperature <- newdata %>%
    # keep variable names from model fitting to ROMS
    mutate(mean_temp_roms_30_norm=(temperature-mean_t)/sd_t)
  
  oxygen <- newdata %>% 
    mutate(mean_oxygen_roms_30_norm=(o2-mean_oxy)/sd_oxy)

  newdata <- newdata %>% 
    left_join(temperature,by = c("year", "latitude", "longitude", "depth", "temperature", "o2")) %>% 
    left_join(oxygen,by = c("year", "latitude", "longitude", "depth", "temperature", "o2")) %>% 
    dplyr::select(-temperature,-o2) %>%
    mutate(year=as.double(year))
  
  # years to predict
  yrs <- sort(unique(newdata$year))
  
  # model weights
  w <- model_df %>% pluck('weight')
  w <- ifelse(all(is.na(w)),rep(1,length(w)),w)
  
  model_list <- model_df %>% pluck('model')
  
  # make the predictions for all 8 models
  all_predictions <- purrr::map2_df(model_list,w,function(m,weight){
    preds <- predict(m,newdata=newdata,return_tmb_object=F,extra_time=yrs) %>% 
      mutate(weight=weight)
    preds
  })
  ens_preds <- all_predictions %>% 
    group_by(year,longitude,latitude,depth) %>% 
    summarise(ens_est=weighted.mean(est,weight)) %>% 
    ungroup()
  
  ens_preds
}
```

```{r,eval=F}
dover_ens_preds <- ensemble_predictions(dover_models)
glimpse(dover_ens_preds)
```

## Plot Ensemble Predictions

```{r}
map_year <- function(model_ens,yr_vec=c(2013),return_pred_df=F,plot_leg=T,use_offshore=F){
  # domain to use
  if(use_offshore){sf_to_use <- ext_offshore}else{sf_to_use <- ext}
  gr <-  sf_to_use %>% st_make_grid(cellsize=10,what='centers') %>% st_as_sf() %>% 
    st_intersection(sf_to_use)
  
  gr_xy <- st_coordinates(gr)
  # scale for the legend, common within species
  scl <- c(0,quantile(exp(model_ens$ens_est),0.99))
  df <- model_ens %>% 
    filter(year%in%yr_vec) %>% 
    group_by(longitude,latitude) %>% 
    summarise(est=mean(ens_est,na.rm=T) %>% exp()) %>% 
    # rescale super large positive outliers for mapping purposes
    mutate(est=ifelse(est>scl[2],scl[2],est))
  # match nearest neighbors from predictions to grid
  pred_points <- df %>% dplyr::select(longitude,latitude) %>% as.matrix()
  nns <- nn2(pred_points,gr_xy,k=1)$nn.idx
  gr_pred <- gr_xy %>% as_tibble() %>% mutate(est=df$est[nns])
  
  bbox=st_bbox(emocc_sf)
  
  if(return_pred_df) {out <- gr_pred %>% as_tibble()}
  
  else{
    out<-ggplot(coast)+
      geom_sf()+
      # geom_point(data=df,aes(longitude,latitude,col=exp(est)))+
      # scale_color_viridis_c()+
      geom_raster(data=gr_pred,aes(x=X,y=Y,fill=est),interpolate=F)+
      geom_sf(data=emocc_sf %>% summarise(),fill=NA)+
      scale_fill_viridis_c(limits=scl)+
      xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
      labs(x="",y="",fill="CPUE",title='')+
      theme(legend.position=c(0.7,0.7))
    if(!plot_leg) out <- out + theme(legend.position = 'None')
  }
  out
}
```

```{r,eval=F}
dover2013 <- map_year(dover_ens_preds)
dover2013
```


# Calculate Proportional Abundance

Using Atlantis polygons, calculate proportional abundance of the species by Atlantis polygon and year

Match the GLORYS points to an Atlantis box

```{r}
glorys_atlantis_spatial_join <- glorys %>% 
  st_join(emocc_sf) %>% 
  st_set_geometry(NULL) %>% 
  distinct(longitude,latitude,box_id,label,area)
glimpse(glorys_atlantis_spatial_join)
```

# Join to Atlantis Polygons

Write a function to extract mean predictions for each Atlantis polygon, by first joining points to polygons using the join key we created above

```{r}
# the `preds` input here is the dataframe of ensemble predictions
ensemble_to_atlantis <- function(preds){
  
  # filter to the year of interest and exponentiate the estimated dependent variable
  df <- preds %>% 
    group_by(year,longitude,latitude) %>% 
    summarise(est=mean(ens_est,na.rm=T) %>% exp()) %>% 
    drop_na() %>% 
    ungroup()
  
  # join the SDM/Atlantis matching key
  atlantis_matched <- df %>% 
    left_join(glorys_atlantis_spatial_join,by=c('longitude','latitude')) %>% 
    drop_na() %>% 
    
    # group by Atlantis box and calculate a mean
    group_by(box_id,year) %>% 
    summarise(mean_est=mean(est,na.rm=T)) %>% 
    ungroup()
    
  # join the paired Atlantis boxes
  atlantis_mean_cpue_all_boxes <- atlantis_matched %>% 
    complete(box_id=full_seq(0:88,1),year,fill=list(mean_est=0)) %>%
    rename(est=mean_est) %>% 
    dplyr::select(box_id,year,est)
  
  # add in area information, scale up, then normalize across all boxes
  atlantis_est_final <- atlantis_mean_cpue_all_boxes %>% 
    left_join(emocc_box_areas,by=c('box_id')) %>% 
    group_by(year) %>% 
    # multiply by total area
    mutate(totest = est*area) %>% 
    # finally, calculate relative index
    mutate(rel_est = totest/sum(totest,na.rm=T)) %>% 
    mutate(perc_est = rel_est*100) %>% 
    ungroup()
  
  return(atlantis_est_final)
}
```

If you want to plot, can pick a year and join the spatial polygon data

```{r}
plot_sdm_year <- function(atlantis_estimates,sppname="",yr=2013){
  polys<- atlantis_estimates %>% 
    filter(year==yr) %>% 
    mutate(perc_est=na_if(perc_est,0)) %>% 
    left_join(emocc_sf,by=c('box_id','area')) %>% 
    st_as_sf()
  bbox=st_bbox(emocc_sf)
  
  t <- paste(sppname,yr)

  out <- ggplot()+
    geom_sf(data=coast,fill='gray50')+
    geom_sf(data=polys,aes(fill=perc_est),col=NA,alpha=1)+
    geom_sf(data=emocc_sf %>% summarise(),fill=NA,col='black')+
    scale_fill_viridis_c(na.value="white")+
    xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
    labs(title=paste("Percent of Total Abundance by Box,",t),fill="Percent")+
    theme(legend.position=c(0.7,0.7))
  
  plot(out)
}
```

Test with our example model

```{r,eval=F}
dover_to_atlantis <- ensemble_to_atlantis(dover_ens_preds)
plot_sdm_year(dover_to_atlantis,yr=2013)
```

```{r}
# write_rds(dover_to_atlantis,here('model output','atlantis sdms','dover sole.rds'))
```

# Combined Function

Write a function that wraps the above to go straight from a set of models to a dataframe of relative abundance in Atlantis polygons.

```{r}
sdm_to_atlantis <- function(model_df,use_offshore=F,save_rds=T){
  # make ensemble predictions
  n <- unique(model_df$spp)
  ens <- model_df %>% 
    ensemble_predictions(use_offshore=use_offshore)
  out <- ens %>% ensemble_to_atlantis()
  
  # save plots of 2013 CPUE and Atlantis proportions
  cpue_plot <- map_year(ens,use_offshore=use_offshore)
  perc_plot <- plot_sdm_year(out,sppname=n,yr=2013)
  
  ggsave(here('model output','atlantis sdms',paste0(n,'_cpue_2013.png')),cpue_plot,h=6,w=5)
  ggsave(here('model output','atlantis sdms',paste0(n,'_atlantis_2013.png')),perc_plot,h=6,w=5)
  
  if(save_rds){
    write_rds(out,here('model output','atlantis sdms',paste0(n,'.rds')))
  }
  return(out)
}
```

```{r,eval=F}
dover <- sdm_to_atlantis(dover_models)
```

# Apply to Atlantis Species

Source the sdmTMB modelling script, make a list of species to model, then try to fit ensemble models. For now, we do not include substrate as a covariate because GLORYS data does not have it.

```{r, message=F, echo=F, include=F}
rmarkdown::render(here::here('scripts','sdmTMB models.Rmd'),quiet=TRUE)
```

## Atlantis Demersal Fish Species

Some Atlantis groups are groups of species (e.g., "Shallow small rockfish"). Therefore, we will make a matching key and a modified version of the trawl survey data that combines these groups.

```{r}
atlantis_groups <- read_csv(here('data','atlantis_demersal_groups.csv'),col_types = 'ccc')
trawl_atlantis <- trawl_roms_utm %>% 
  left_join(atlantis_groups,by=c('species'='trawl_name')) %>% 
  filter(!is.na(code))
unique(trawl_atlantis$code)
unique(trawl_atlantis$group_name)

# groups by Atlantis code/group and summarize
trawl_atlantis <- trawl_atlantis %>% 
  group_by(date,trawl_id,station,depth_trawl,mean_temp_roms_30,mean_oxygen_roms_30,prop_hard_mixed,latitude,longitude,code,group_name) %>% 
  summarise(cpue_kg_km2=sum(cpue_kg_km2)) %>% 
  #rename so the models work
  rename(species=code) %>% 
  ungroup()
```

## Run Models

```{r}
nms <- unique(trawl_atlantis$species)
# for all groups except FBP (lanternfish, deep vertical migrators), don't use offshore boxes for prediction
nms <- nms %>% str_subset("FBP",negate=T)

for(i in nms){
  tic(paste("Processing",i))
  model_df <- model_species(i,data=trawl_atlantis,use_substrate = F)
  atlantis_df <- sdm_to_atlantis(model_df,use_offshore = F)
  write_rds(atlantis_df,here('model output','atlantis sdms',paste0(i,'.rds')))
  toc()
}

# model FBP last
tic(paste("Processing FBP"))
model_df <- model_species("FBP",data=trawl_atlantis,use_substrate = F)
atlantis_df <- sdm_to_atlantis(model_df,use_offshore=T)
write_rds(atlantis_df,here('model output','atlantis sdms','FBP.rds'))
toc()
```

## Plots

```{r,fig.width=10,fig.height=10}
# plts<-purrr::map(nms,function(fl){
#   x <- read_rds(here('model output','atlantis sdms',paste0(fl,'.rds')))
#   plot_sdm_year(x,fl,yr=2013)
# })
# 
# library(cowplot)
# 
# plot_grid(plotlist=plts[1:4],nrow=2,ncol=2)
# plot_grid(plotlist=plts[5:8],nrow=2,ncol=2)
# plot_grid(plotlist=plts[9:12],nrow=2,ncol=2)
# plot_grid(plotlist=plts[13:16],nrow=2,ncol=2)
# plot_grid(plotlist=plts[17:20],nrow=2,ncol=2)
# plts[21]
```


