---
title: "Run sdmTMB"
author: "Owen Liu"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
# devtools::install_github("pbs-assess/sdmTMB")
library(sdmTMB)
library(tidyverse)
library(lubridate)
library(sf)
library(here)
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)
```

# Purpose

Using sdmTMB, fit models for west coast groundfish based on hindcast ROMS oceanographic data.

# Import Data

## Trawl Data

Trawl data, including a matching key to link to ROMS data. The hindcast ROMS data has values for all trawl survey locations for all times, but we just want the values matched to the actual trawl survey times/dates.

```{r}
trawl <- read_rds(here('data','nwfsc_trawl_data.rds')) %>% 
  # convert date from character to date
  mutate(date=as_date(date))
# roms time refernce
roms_time <- read_rds(here('data','roms_time_date_reference.rds'))
trawl_locs <- read_rds(here::here('data','trawl','trawlID.rds')) %>% 
  # add a dummy indicator of a "real" trawl survey location in time/space
  mutate(date=as_datetime(trawl_time,origin="1900-01-01")) %>%
  mutate(date=as_date(date)) %>%
  mutate(realTrawl=1) %>% 
  # join the roms_time reference
  left_join(roms_time) %>% 
  # select the variable we'll use to match
  select(station,date,time,lon_trawl,lat_trawl,depth_trawl,realTrawl)

trawl <- trawl %>% 
  left_join(trawl_locs,by=c("date"="date","longitude_dd"="lon_trawl","latitude_dd"="lat_trawl","depth"="depth_trawl"))
```

Filter the trawl survey data to fit within the time frame for which we have ROMS hindcast data (1980-2010)

```{r}
trawl <- trawl %>%
  filter(realTrawl==1) %>% 
  select(-realTrawl) %>% 
  # rename time to something more useful
  rename(roms_hindcast_day=time)
```

## ROMS data

Hindcast ROMS data matched to trawl survey locations and times

```{r}
roms <- read_rds(here::here('data','joined_30d_lagged_t_o.rds')) %>%
  # join the trawl_locs and filter by actual trawl locations and times
  left_join(trawl_locs) %>% 
  filter(realTrawl==1) %>% 
  select(-realTrawl) %>% 
  # rename time to something more useful
  rename(roms_hindcast_day=time)
```


Look at the form of these data

```{r}
glimpse(trawl)
glimpse(roms)
```

For the ROMS data (for now), we are using modelled temperature and oxygen, lagged 30 days from each trawl survey location and time.

```{r}
roms_thin <- roms %>% 
  dplyr::select(station,lon_trawl,lat_trawl,depth_trawl,mean_temp_roms_30,mean_oxygen_roms_30)
```

# Join Trawl and ROMS

Join the two datasets together, such that we have the appropriately-matched ROMS outputs

```{r}
trawl_roms <- trawl %>% left_join(roms,by = c("date", "station", "roms_hindcast_day")) %>% 
  #clean up some columns
  dplyr::select(date,station,lon_trawl,lat_trawl,depth_trawl,mean_temp_roms_30,mean_oxygen_roms_30,species,cpue_kg_km2) %>% 
  # drop any rows with NAs
  drop_na()
```


# Prepare Data for sdmTMB

Convert the trawl spatial data to UTM. 

```{r}
# convert to UTM
trawl_roms_utm <- trawl_roms %>%
  # convert to sf object
  st_as_sf(coords=c('lon_trawl','lat_trawl'),crs=4326) %>%
  # transform to UTM zone 10
  st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km") %>%
  # add new coords as vars
  mutate(latitude = sf::st_coordinates(.)[,2],
         longitude = sf::st_coordinates(.)[,1]) %>%
  # convert back to normal df
  st_set_geometry(NULL)
```

# Functions to Run a Model

We'll write two functions, one to prepare a specific species for an sdmTMB model, and another to actually run an sdmTMB model with custom options

## Prepare Species' Data

This function selects a species' data from the trawl survey data, converts the spatial data to UTM, does a couple of filters for missing data, and then joins the ROMS hindcast data to it by time and location.

```{r}
prepare_species <- function(dat,spp){
  dat_sub <- dat %>% 
    filter(species==spp) %>% 
    
    # rescale depth, oxygen, and temp to be N(0,1)
    mutate(across(c(depth_trawl,mean_temp_roms_30,mean_oxygen_roms_30),~(scale(.) %>% as.vector()),.names="{.col}_norm")) %>% 
    
    # add a year indicator
    mutate(year=lubridate::year(date))
}
```

Try an example for sablefish

```{r}
sablefish_dat <- prepare_species(trawl_roms_utm,spp="sablefish")
glimpse(sablefish_dat)
```

## sdmTMB Model Function

Write a function that runs sdmTMB. It wil call the previous function to make the appropriate species data. For now, the environmental variable names are not generic (always `mean_temp_roms_30_norm` and `mean_oxygen_roms_30_norm`)

```{r}
run_sdmTMB <- function(dat,spp,nknots=250,use_depth=F,time_varying=F,spatial_only=T){
  
  # pull the species' data
  modeldat <- prepare_species(dat,spp)
  
  # make spde
  spde <- make_mesh(modeldat,xy_cols = c('longitude','latitude'), 
                    n_knots = nknots)
  
  # model formula
  formula <- paste0("cpue_kg_km2 ~ -1")
  # if depth effect, add to model formla
  if(use_depth) {
    formula = paste0(formula, " + depth + I(depth^2)")
  }
  time_formula = "~ -1"
  if(time_varying) {
    time_formula = paste0(time_formula, " + ", 
                          "mean_temp_roms_30_norm+mean_temp_roms_30_norm")
    time_varying = as.formula(time_formula)
    time = "year"
  } else {
    formula = paste0(formula, " + ", 
                     "mean_temp_roms_30_norm+mean_temp_roms_30_norm")
    time_varying = NULL
    time = "year"
  }
  # formula = paste0(formula, " + as.factor(year)")
  
  # fit model
  m <- try( sdmTMB( 
    formula = as.formula(formula),
    #formula = cpue_kg_km2 ~ as.factor(year) + depth,
    time_varying = time_varying,
    spde = spde,
    time = time,
    family = tweedie(link = "log"),
    data = modeldat,
    anisotropy = TRUE,
    spatial_only = spatial_only,
    quadratic_roots = TRUE
  ), 
  silent=TRUE)
  
  if(class(m)!="try-error") {
    write_rds(m, path=here::here('model output',
                                 paste0(spp,'.rds')))
  }
  print(paste("Model for",spp,"complete."))
  
  return(m)
}
```

Try

```{r}
sable <- run_sdmTMB(trawl_roms_utm,spp="sablefish")
```

Crashing.
