---
title: "sdmTMB output plotting"
author: "Owen Liu"
date: "7/9/2021"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r, include=FALSE}
# devtools::install_github("pbs-assess/sdmTMB")
library(sdmTMB)
library(tidyverse)
library(lubridate)
library(sf)
library(here)
# vista is Eric Ward's library for looking at outputs
library(vista)

knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)
```

# Purpose

Using the outputs from species distribution models in the script `sdmTMB models`, visualize the outputs to investigate the behavior of the models.

```{r}
# sdmTMB test (vignette)
# data(pcod)
# m <- sdmTMB(data = pcod,
# formula = density ~ 0 + as.factor(year),
# time = "year", spde = make_mesh(pcod, c("X", "Y"), n_knots = 80, type = "kmeans"),
# family = tweedie(link = "log"), anisotropy = TRUE,
# include_spatial = FALSE)
# pcod$pred = predict(m)$est
# pcod$resid = residuals(m)
# pred_space(pcod,time="year")
```

## Load Models

```{r}
sdms <- read_rds(here::here('model output','four species sdmTMB test.rds')) %>% 
  # split the nested data frame of models into a list of 4 (one list of models per species)
  group_split(spp) %>% 
  set_names(c('canary rockfish','darkblotched rockfish','sablefish','shortspine thornyhead'))
```

In the testing of sdmTMB, 8 different models were run for four different species. For the purposes of testing diagnostic plots, let's test the diagnostic plots on some of these 32 models.

```{r}
# here is one of the sablefish models
sable_model <- sdms %>% pluck('sablefish','model',8)
sable_data <- sable_model$data
sable_data$pred = predict(sable_model)$est
sable_data$resid = residuals(sable_model)

pred_space(sable_data,time="year",X="longitude",Y="latitude")+
  coord_equal()
pred_time(sable_data,time="year")
```


# Run Diagnostic plots

```{r}
d <- diagnostic_plots(sable_data,time="year",X="longitude",Y="latitude")
```


