---
title: "Atlantis SDMs with GLORYS"
author: "Owen Liu"
date: "1/4/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r, include=FALSE}
# devtools::install_github("pbs-assess/sdmTMB")
library(sdmTMB)
library(rbgm)
library(tidyverse)
library(lubridate)
library(rnaturalearth)
library(sf)
library(here)
library(cowplot)
library(RANN)
library(furrr)
library(future)
library(viridis)
library(tictoc)

knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)
```

```{r}
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

```{r}
# set up parallel processing?
# future::plan(sequential)
# future::plan(multisession,workers=4)
```

```{r}
# Atlantis grid
fl <- here::here("data",'grids',"CalCurrentV3_utm.bgm")
# load the file
bgm <- read_bgm(fl)
names(bgm)
emocc_sf <- box_sf(bgm) %>% dplyr::select(box_id,label,area,botz) %>% 
  st_transform("+proj=utm +zone=10 +datum=WGS84 +units=km")
# areas of boxes
emocc_box_areas <- emocc_sf %>% st_set_geometry(NULL) %>% dplyr::select(-label)
```

```{r}
# GLORYS temp/oxygen data
glorys <- read_rds(here('data','GLORYS_bottom_temp_o2_SDMs.rds')) %>% 
  # convert to UTM zone 10
  st_as_sf(coords=c('longitude','latitude'),crs=4326) %>% 
  st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km") %>% 
  mutate(latitude = sf::st_coordinates(.)[,2],
         longitude = sf::st_coordinates(.)[,1])
```

```{r}
# spatial background map
# load west cost land for mapping
coast <- ne_countries(continent="North America",scale=50,returnclass = 'sf') %>% 
  st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km") %>% 
  st_crop(emocc_sf)
# coast <- ne_states(country='United States of America',returnclass = 'sf') %>% 
#   filter(name %in% c('California','Oregon','Washington','Nevada')) %>%
#   st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km")
```


# Purpose

Produce groundfish species distribution models for use in the new version of the Atlantis ecosystem model. We will use `sdmTMB` to fit SDMs to high-resolution ROMS and trawl survey data, establishing the environmental niches of groundfish relative to bottom temperature and oxygen. Then, we use a GLORYS/ESM hybrid oceanography product to project species distributions out to 2100 in the California Current. Finally, we will translate those SDMs on to the Atlantis polygon/grid structure.

Note: Model runs are described and performed in the `sdmTMB models.Rmd` script. Here, we apply the new GLORYS data for projection.

# Calculate Distance from Shore

For some applications of the GLORYS projection outputs, we want a measure of how far each cell is from shore. We calculate that here.

```{r}
coast %>% ggplot()+geom_sf()
glorys_xy <- glorys %>% st_set_geometry(NULL) %>% distinct(longitude,latitude) %>% 
  st_as_sf(coords=c('longitude','latitude'),crs=st_crs(glorys),remove=F)
glorys_xy %>% ggplot()+geom_sf()
coastsumm <- coast %>% summarise()
glorys_coast_dist <- glorys_xy %>% st_distance(coastsumm)
glorys_dist_to_coast <- glorys_xy %>% 
  mutate(km_to_coast=as.numeric(glorys_coast_dist)) %>%
  st_set_geometry(NULL)
glorys_dist <- glorys %>% left_join(glorys_dist_to_coast,by=c('longitude','latitude'))

glorys_dist_to_coast %>% 
  ggplot(aes(longitude,latitude,col=km_to_coast))+
  geom_point(size=0.5)

# polygon for <= 100km offshore
glorys_100km_buffer1 <- coastsumm %>% 
  st_buffer(100) %>% 
  st_difference(coastsumm) %>% 
  st_crop(emocc_sf %>% summarise())
ggplot(glorys_100km_buffer1)+
  geom_sf()

write_rds(glorys_dist_to_coast,here('data','grids','glorys_distance_to_coast.rds'))
write_rds(glorys_100km_buffer1,here('data','grids','coast_100km_buffer.rds'))
```


# Define Projection Extent

Here, we set the boundaries for the `sdmTMB` projection. We use three criteria: 1) we crop the GLORYS data to the extent of the Atlantis polygons; 2) for all species except lanternfish/myctophids, we want to we exclude offshore Atlantis polygons from the projection; 3) for lanternfish, we will keep all boxes.

```{r}
# trawl IDs and depths
trawls <- read_rds(here::here('data','trawl','trawlID.rds'))
max_depth <- max(trawls$depth_trawl,na.rm=T)

# glorys bbox with areas that are too deep removed
glorys_full <- glorys %>% summarise() %>% st_convex_hull()%>% 
  st_difference(coast %>% summarise())

# roms domain (for display purposes)
roms_domain <- read_rds(here::here('data','roms_latlon_depth_key_topleft_start.rds')) %>% 
  mutate(roms_cell=row_number()) %>% 
  st_as_sf(coords=c('lon','lat'),crs=4326) %>%
  summarise() %>% 
  st_convex_hull() %>%
  st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km") %>% 
  st_difference(coast %>% summarise())

atlantis_glorys_extent <- glorys %>% 
  filter(depth<max_depth) %>% 
  st_bbox() %>% 
  st_as_sfc() %>% 
  st_difference(coast %>% summarise())

# atlantis polygons, summarized/unioned
emocc_merged <- emocc_sf %>% summarise()
emocc_no_offshore_merged <- emocc_sf %>% 
  filter(botz!=min(botz)) %>% 
  summarise()

# plot our starting point(s)
ggplot()+
  geom_sf(data=coast)+
  geom_sf(data=glorys_full,fill='blue',alpha=0.5)+
  geom_sf(data=emocc_sf,fill='red',alpha=0.5)+
  geom_sf(data=roms_domain,fill='darkgreen',alpha=0.5)

ggplot()+
  geom_sf(data=coast)+
  geom_sf(data=atlantis_glorys_extent,fill='blue',alpha=0.5)+
  geom_sf(data=emocc_no_offshore_merged,fill='red',alpha=0.5)


# crop glorys bbox with atlantis polygons
ext <- atlantis_glorys_extent %>% 
  st_intersection(emocc_no_offshore_merged)
ext_offshore <- atlantis_glorys_extent %>% 
  st_intersection(emocc_merged)

# plot the final domain
ggplot()+
  geom_sf(data=coast)+
  geom_sf(data=ext_offshore,fill='yellow',alpha=0.5)+
  geom_sf(data=ext,fill='purple',alpha=0.5)
```

Use this projection extent to crop the GLORYS data used for projection

```{r}
glorys_crop <- glorys %>% 
  # crop to SDM extent
  st_intersection(ext_offshore) %>%
  st_set_geometry(NULL)
glorys_no_offshore_crop <- glorys %>% 
  # crop to SDM extent
  st_intersection(ext) %>%
  st_set_geometry(NULL)
```

# Load Fitted Models

Start with an example species: Dover sole. See `sdmTMB models.Rmd` for details of the ensemble modelling approach.

```{r,eval=F}
## dover sole models
dover_models <- read_rds(here::here('model output','dover sole no substrate.rds'))
names(dover_models)
```

# GLORYS Projection Function

```{r,eval=T}
make_predictions_glorys <- function(modelobj){

  original_model_data <- modelobj$data
  # need to use the original (hindcast) environmental data to scale the projected data
  mean_t <- mean(original_model_data$mean_temp_roms_30,na.rm=T)
  sd_t <- sd(original_model_data$mean_temp_roms_30,na.rm=T)
  mean_oxy <- mean(original_model_data$mean_oxygen_roms_30,na.rm=T)
  sd_oxy <- sd(original_model_data$mean_oxygen_roms_30,na.rm=T)
  
  # create a new tibble with the projected data from GLORYS
  newdata <- glorys_crop %>% 
    # No habitat data
    # left_join(hab,by=c("lat","lon")) %>% 
    # drop_na() %>% 
    dplyr::select(year,latitude,longitude,depth,temperature,o2)
  
  temperature <- newdata %>%
    # keep variable names from model fitting to ROMS
    mutate(mean_temp_roms_30_norm=(temperature-mean_t)/sd_t)
  
  oxygen <- newdata %>% 
    mutate(mean_oxygen_roms_30_norm=(o2-mean_oxy)/sd_oxy)

  newdata <- newdata %>% 
    left_join(temperature,by = c("year", "latitude", "longitude", "depth", "temperature", "o2")) %>% 
    left_join(oxygen,by = c("year", "latitude", "longitude", "depth", "temperature", "o2")) %>% 
    dplyr::select(-temperature,-o2) %>%
    mutate(year=as.double(year))
  
  # years to predict
  yrs <- sort(unique(newdata$year))
  # now we can make the predictions
  predicted_cpue_km2 <- predict(modelobj,newdata,return_tmb_object=T,extra_time=yrs)
  
  predicted_cpue_km2
}
```

Test it
```{r,eval=F}
dover_future1 <- make_predictions_glorys(dover_models %>% pluck('model',1))
glimpse(dover_future1$data)
```

# Diagnostic Plots

```{r}
create_diag <- function(m,return_what="plots"){
  x <- m$data
  x$pred <- predict(m)$est
  x$resid <- residuals(m)[,1]
  x <- x %>% select(year,longitude,latitude,cpue_kg_km2,pred,resid) %>% 
    rename(X=longitude,Y=latitude,time=year) %>% 
    filter(time %in% 2003:2010) %>% 
    mutate(logcpue=log(cpue_kg_km2+1))
  
  if(return_what=="plots") vista::diagnostic_plots(x)
  else(x)
}
```

# Ensemble Predictions Function

To create ensemble predictions, we calculate predictions for each species' model separately, then weight them using the weights established in model fitting (see `sdmTMB::sdmTMB_stacking()`).

```{r}
# this function is meant to be run with a list of models that all are fit on the same data, with a common set of predictors
ensemble_predictions <- function(model_df,use_offshore=F){
  
  # use the first model in the list to pull out the data
  original_model_data <- model_df %>% pluck('model',1,'data')
  # need to use the original (hindcast) environmental data to scale the projected data
  mean_t <- mean(original_model_data$mean_temp_roms_30,na.rm=T)
  sd_t <- sd(original_model_data$mean_temp_roms_30,na.rm=T)
  mean_oxy <- mean(original_model_data$mean_oxygen_roms_30,na.rm=T)
  sd_oxy <- sd(original_model_data$mean_oxygen_roms_30,na.rm=T)
  
  # create a new tibble with the projected data from GLORYS
  if(use_offshore){glorys_to_use <- glorys_crop}else{glorys_to_use <- glorys_no_offshore_crop}
  newdata <- glorys_to_use %>% 
    # No habitat data
    # left_join(hab,by=c("lat","lon")) %>% 
    # drop_na() %>% 
    dplyr::select(year,latitude,longitude,depth,temperature,o2)
  
  temperature <- newdata %>%
    # keep variable names from model fitting to ROMS
    mutate(mean_temp_roms_30_norm=(temperature-mean_t)/sd_t)
  
  oxygen <- newdata %>% 
    mutate(mean_oxygen_roms_30_norm=(o2-mean_oxy)/sd_oxy)

  newdata <- newdata %>% 
    left_join(temperature,by = c("year", "latitude", "longitude", "depth", "temperature", "o2")) %>% 
    left_join(oxygen,by = c("year", "latitude", "longitude", "depth", "temperature", "o2")) %>% 
    dplyr::select(-temperature,-o2) %>%
    mutate(year=as.double(year))
  
  # years to predict
  yrs <- sort(unique(newdata$year))
  
  # model weights
  w <- model_df %>% pluck('weight')
  w <- ifelse(all(is.na(w)),rep(1,length(w)),w)
  
  model_list <- model_df %>% pluck('model')
  
  # make the predictions for all 8 models
  all_predictions <- purrr::map2_df(model_list,w,function(m,weight){
    preds <- predict(m,newdata=newdata,return_tmb_object=F,extra_time=yrs) %>% 
      mutate(weight=weight)
    preds
  })
  ens_preds <- all_predictions %>% 
    group_by(year,longitude,latitude,depth) %>% 
    summarise(ens_est=weighted.mean(est,weight)) %>% 
    ungroup()
  
  ens_preds
}
```

```{r}
create_diag_ensemble <- function(model_df,return_what="plots"){
  all_preds <- purrr::map2_df(model_df$model,model_df$weight,function(df,wt){
    create_diag(df,return_what = 'df') %>% mutate(weight=wt)})
  all_preds %>% 
    mutate(weighted_pred=pred*weight) %>% 
    group_by(time,X,Y,logcpue) %>% 
    summarise(enspred=sum(weighted_pred)) %>% 
    ungroup()
}
```

## Plot Ensemble Predictions

```{r}
map_year <- function(model_df,yr_vec=c(2013),return_pred_df=F,plot_leg=T,use_offshore=F){
  model_ens <- ensemble_predictions(model_df)
  # domain to use
  if(use_offshore){sf_to_use <- ext_offshore}else{sf_to_use <- ext}
  gr <-  sf_to_use %>% st_make_grid(cellsize=10,what='centers') %>% st_as_sf() %>% 
    st_intersection(sf_to_use)
  
  gr_xy <- st_coordinates(gr)
  # scale for the legend, common within species
  scl <- c(0,stats::quantile(exp(model_ens$ens_est),0.99))
  df <- model_ens %>% 
    filter(year%in%yr_vec) %>% 
    group_by(longitude,latitude) %>% 
    summarise(est=mean(ens_est,na.rm=T) %>% exp()) %>% 
    # rescale super large positive outliers for mapping purposes
    mutate(est=ifelse(est>scl[2],scl[2],est))
  # match nearest neighbors from predictions to grid
  pred_points <- df %>% dplyr::select(longitude,latitude) %>% as.matrix()
  nns <- nn2(pred_points,gr_xy,k=1)$nn.idx
  gr_pred <- gr_xy %>% as_tibble() %>% mutate(est=df$est[nns])
  
  bbox=st_bbox(emocc_sf)
  
  if(return_pred_df) {out <- gr_pred %>% as_tibble()}
  
  else{
    out<-ggplot(coast)+
      geom_sf()+
      # geom_point(data=df,aes(longitude,latitude,col=exp(est)))+
      # scale_color_viridis_c()+
      geom_raster(data=gr_pred,aes(x=X,y=Y,fill=est),interpolate=F)+
      geom_sf(data=emocc_sf %>% summarise(),fill=NA)+
      scale_fill_viridis_c(limits=scl)+
      xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
      labs(x="",y="",fill="CPUE",title='')+
      theme(legend.position=c(0.7,0.7))
    if(!plot_leg) out <- out + theme(legend.position = 'None')
  }
  out
}
```

```{r}
# map_year <- function(model_df,yr_vec=c(2013),return_pred_df=F,plot_leg=T,use_offshore=F,n_slices=1){
#   model_ens <- ensemble_predictions(model_df)
#   # domain to use
#   if(use_offshore){sf_to_use <- ext_offshore}else{sf_to_use <- ext}
#   gr <-  sf_to_use %>% st_make_grid(cellsize=10,what='centers') %>% st_as_sf() %>% 
#     st_intersection(sf_to_use)
#   
#   gr_xy <- st_coordinates(gr)
#   # scale for the legend, common within species
#   scl <- c(0,quantile(exp(model_ens$ens_est),0.99))
#   df <- model_ens %>% 
#     filter(year%in%yr_vec) %>% 
#     group_by(longitude,latitude) %>% 
#     summarise(est=mean(ens_est,na.rm=T) %>% exp()) %>% 
#     # rescale super large positive outliers for mapping purposes
#     mutate(est=ifelse(est>scl[2],scl[2],est))
#   # match nearest neighbors from predictions to grid
#   pred_points <- df %>% dplyr::select(longitude,latitude) %>% as.matrix()
#   nns <- nn2(pred_points,gr_xy,k=1)$nn.idx
#   gr_pred <- gr_xy %>% as_tibble() %>% mutate(est=df$est[nns]) %>% 
#     mutate(zone=ntile(Y,n_slices))
#   
#   # bbox=st_bbox(gr)
#   
#   if(return_pred_df) {out <- gr_pred %>% as_tibble()}
#   
#   else{
#     outs <- purrr::map(1:n_slices,function(x){
#       tdf <- gr_pred %>% filter(zone==x)
#       bbox=tdf %>% st_as_sf(coords=c("X","Y"),crs="+proj=utm +zone=10 +datum=WGS84 +units=km") %>% st_bbox()
#       p<-ggplot()+
#         # geom_point(data=df,aes(longitude,latitude,col=exp(est)))+
#         # scale_color_viridis_c()+
#         geom_raster(data=tdf,aes(x=X,y=Y,fill=est),interpolate=F)+
#         geom_sf(data=emocc_sf %>% summarise(),fill=NA)+
#         geom_sf(data=coast)+
#         scale_fill_viridis_c(limits=scl)+
#         coord_sf(datum=NA)+
#         xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
#         labs(x="",y="",fill="CPUE",title=yr_vec)+
#         theme(legend.position='right',
#               legend.text = element_text(size=8))
#     # if(!plot_leg) out <- out + theme(legend.position = 'None')
#     })
#    out <- plot_grid(plotlist=rev(outs),ncol=1,axis='l') 
#   }
#   out
# }
```

```{r,eval=F}
dover2013 <- map_year(dover_models,plot_leg=F)
dover2100 <- map_year(dover_models,yr_vec = 2100)
plot_grid(dover2013,dover2100,nrow=1)
# dover2013
# dover2100
```

## Center of Gravity

```{r}
make_cog <- function(model_df,what="plot"){
  # future::plan(multisession)
  
  sppname <- unique(model_df$spp) %>% tools::toTitleCase()
  
  # model weights
  w <- model_df %>% pluck('weight')
  if(all(is.na(w))) w <- rep(1,nrow(model_df))
  
  # tic('making predictions')
  predictions_list <- model_df %>% pluck('model') %>% 
    purrr::map(make_predictions_glorys) 
  # toc()
  
  # tic('making cogs')
  
  cogs <- predictions_list %>% purrr::map_dfr(get_cog)
  # toc()
  
  nyr <- length(unique(cogs$year))
  ws <- rep(w,each=nyr)
  
  
# reorganize output (X and Y coords)
  cog_x <- cogs %>% 
    filter(coord=="X")%>% 
    rename_with(~paste0(.,"_x"),all_of(c('est','lwr','upr','se'))) %>% 
    dplyr::select(-coord) %>% 
    # add weights
    mutate(weight=ws) %>% 
    group_by(year) %>% 
    summarise(w.est_x=weighted.mean(est_x,w=weight)) %>% 
    ungroup()
  cog_y <- cogs %>% 
    filter(coord=="Y") %>% 
    rename_with(~paste0(.,"_y"),all_of(c('est','lwr','upr','se')))%>% 
    dplyr::select(-coord) %>% 
    # add weights
    mutate(weight=ws) %>% 
    group_by(year) %>% 
    summarise(w.est_y=weighted.mean(est_y,w=weight)) %>% 
    ungroup()
  cog_2d <- cog_x %>% left_join(cog_y,by='year') %>% 
    # fix for weird non-overlap between ROMS fitting and GLORYs projections
    mutate(year=year+26) %>% 
    drop_na()
  
  cog_x_plot <- cog_2d %>% 
    # ggplot(aes(year,w.est_x,ymax=upr_x,ymin=lwr_x))+
    ggplot(aes(year,w.est_x))+
    geom_line()+
    # geom_ribbon(fill='red',alpha=0.5)+
    scale_x_continuous(expand=c(0,5))+
    labs(x="Year",y="Eastings (km)",title="Center of Gravity, X")
  
  cog_y_plot <- cog_2d %>% 
    # ggplot(aes(year,w.est_y,ymax=upr_y,ymin=lwr_y))+
    ggplot(aes(year,w.est_y))+
    geom_line()+
    # geom_ribbon(fill='red',alpha=0.5)+
    scale_x_continuous(expand=c(0,5))+
    labs(x="Year",y="Northings (km)",title="Center of Gravity, Y")
  
  cog2d_timeseries <- plot_grid(cog_x_plot,cog_y_plot,nrow=2)
  
  cog_2d_sf <- cog_2d %>% 
    drop_na() %>% 
    st_as_sf(coords=c('w.est_x','w.est_y'),crs="+proj=utm +zone=10 +datum=WGS84 +units=km")

  bbox <- st_bbox(ext_offshore)

  cog_spatial <- ggplot()+
      geom_sf(data=coast)+
      geom_sf(data=cog_2d_sf,aes(col=year))+
      labs(title=paste(sppname,"Center of Gravity"),col="Year")+
      xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])

  out <- plot_grid(cog_spatial,cog2d_timeseries,ncol=2)
  if(what=="plot") return(out) else return(cog_2d)
}
```

## Depth Distribution

```{r}
plot_depth_distribution <- function(model_df){
  ens_preds <- ensemble_predictions(model_df)
  y <- ens_preds %>% 
    filter(year==2013|year==2100) %>%
    mutate(cpue=exp(ens_est)) %>% 
    group_by(year,depth) %>% 
    summarise(cpue=mean(cpue,na.rm=T)) %>% 
    ungroup() %>% 
    group_by(year) %>% 
    mutate(prop_cpue=cpue/sum(cpue,na.rm=T)) %>%
    arrange(depth) %>% 
    mutate(cum_cpue=cumsum(prop_cpue)) %>% 
    ungroup()
  y %>% 
    ggplot(aes(-depth,cum_cpue,col=factor(year)))+
    geom_line(size=2)+
    coord_flip()+
    scale_color_manual(values=c("#2271B2","#d55e00"))+
    labs(x="Depth (m)",y="Cumulative CPUE",col="Year",title=paste0(unique(model_df$spp)," Predicted Depth Distribution"))
}
```

## Environmental Relationships

```{r}
latlons <- glorys_crop %>% distinct(latitude,longitude)
cropto <- glorys_crop %>% 
  st_as_sf(coords=c("temperature","o2")) %>% 
  ungroup() %>% 
  summarise() %>% 
  st_convex_hull()

pred_template <- crossing(
  mean_temp_roms_30_norm=seq(min(glorys_crop$temperature)+0.2,max(glorys_crop$temperature)-0.2,length.out=100),
  mean_oxygen_roms_30_norm=seq(min(glorys_crop$o2)+0.2,max(glorys_crop$o2)-0.2,length.out=150)) %>% 
  bind_cols(latlons %>% slice_sample(n=15000,replace=T)) %>% mutate(year=2006)
pred_template_sf <- pred_template %>% 
      st_as_sf(coords=c("mean_temp_roms_30_norm","mean_oxygen_roms_30_norm"))
tempr <- raster::raster(xmn=0,xmx=23,ymn=0,ymx=300,nrows=150,ncol=100)


rasterize_affinities <- function(model_df, template_raster=tempr,return_what="plot"){
  
  original_model_data <- model_df %>% pluck('model',1,'data')
  # need to use the original (hindcast) environmental data to scale the projected data
  mean_t <- mean(original_model_data$mean_temp_roms_30,na.rm=T)
  sd_t <- sd(original_model_data$mean_temp_roms_30,na.rm=T)
  mean_oxy <- mean(original_model_data$mean_oxygen_roms_30,na.rm=T)
  sd_oxy <- sd(original_model_data$mean_oxygen_roms_30,na.rm=T)
  
  # create a new tibble with the projected data from GLORYS
  newdata <- crossing(
    temperature=seq(0,23,length.out=100),
    o2=seq(0,300,length.out=150)) %>% 
    bind_cols(latlons %>% slice_sample(n=15000,replace=T)) %>% mutate(year=2003) %>% 
    # keep variable names from model fitting to ROMS
    mutate(mean_temp_roms_30_norm=(temperature-mean_t)/sd_t,
           mean_oxygen_roms_30_norm=(o2-mean_oxy)/sd_oxy)
  
  preds <- purrr::map_df(1:nrow(model_df),function(x){
    m <- model_df %>% pluck('model',x)
    w <- model_df %>% pluck('weight',x)
    ests <- m %>% predict(newdata=newdata,se_fit=F) %>% 
      dplyr::select(longitude:est) %>% 
      mutate(weight=w)
    ests
  })
  
  preddf <- preds %>% 
    mutate(cpue=exp(est)) %>% 
    mutate(w.cpue=cpue*weight) %>% 
    group_by(temperature,o2) %>% 
    summarise(w.est=sum(w.cpue)) %>% 
    ungroup() %>% 
    st_as_sf(coords=c("temperature","o2"),remove=F) %>% 
    st_intersection(cropto)
    # st_set_geometry(NULL)
  
  ul <- quantile(preddf$w.est,0.99,na.rm=T)
  ll <- quantile(preddf$w.est,0.01,na.rm=T)
  preddf <- preddf %>% 
    mutate(w.est=ifelse(w.est>ul,ul,w.est)) %>% 
    rename(cpue=w.est)
  spp_r <- preddf %>% raster::rasterize(template_raster,field="cpue",fun=mean)
  spp_df <- spp_r %>% raster::rasterToPoints() %>% as_tibble() %>% set_names(c('temp','oxy','cpue'))
  # make the plot
    
  mainp <- spp_df %>% 
    ggplot()+
    geom_tile(aes(temp,oxy,fill=cpue))+
    # geom_sf(data=cropto,color='black',fill=NA)+
    # coord_sf(datum=NA)+
    scale_fill_viridis(na.value='white',name="CPUE")+
    labs(x="",y="",title='')+
    annotate('text',3,270,label=unique(model_df$spp))+
    theme(panel.background = element_rect(fill='white'),
          # panel.background = element_rect(fill=viridis_pal()(1)),
          # legend.position = c(0.1,0.7),
          # legend.background = element_rect(fill='white'),
          panel.grid.major = element_blank(),
          panel.grid.minor=element_blank())
  
  if(return_what=="legend") return(get_legend(mainp))
  else{
      
  m1 <- spp_df %>% group_by(temp) %>% summarise(m=sum(cpue)) %>% 
    ggplot(aes(temp,m))+geom_area(fill='grey50')+theme_void()+
    theme(plot.margin = margin(0,-0.1,-0.1,0,'cm'))
  
  m2 <- spp_df %>% group_by(oxy) %>% summarise(m=sum(cpue)) %>% 
    ggplot(aes(oxy,m))+geom_area(fill='grey50')+theme_void()+coord_flip()+
    theme(plot.margin = margin(0,0,0,0,'cm'))
  
  col1 <- plot_grid(m1,NULL,mainp+theme(legend.position='none',plot.margin = margin(0,0,0,0,'pt')),nrow=3,align='v',axis='lr',rel_heights = c(0.2,-0.07,1))
  # grobs <- ggplotGrob(row1)
  # g <- grid::rectGrob(gp = grid::gpar(col=0))
  # row1 <- plot_grid(NULL,m1,NULL,nrow=1,rel_widths = c(0.5,5,1.15))
  col2 <- plot_grid(NULL,m2,NULL,nrow=3,rel_heights = c(0.25,1,0.15))
  out <- plot_grid(col1,col2,ncol=2,rel_widths =c(1,0.25))
  # out <- plot_grid(m1,NULL,mainp+rremove('legend')+rremove(),m2,ncol=2,align='hv',rel_widths = c(1,0.25),rel_heights = c(0.25,1))
  out
  }

}
```

## Hindcast Trawl Survey

Compare to trawl survey observations, but using GLORYS projection
```{r}
trawl_sf <- trawls %>% 
  st_as_sf(coords=c('lon_trawl','lat_trawl'),crs=4326) %>% st_transform(st_crs(glorys))
glorys_grd <- glorys_crop %>% 
  distinct(latitude,longitude) %>% 
  st_as_sf(coords=c('longitude','latitude'),crs=st_crs(glorys),remove=F) %>% 
  mutate(grdid=row_number())
  
nns <- nngeo::st_nn(trawl_sf,glorys_grd,maxdist = 50000) %>% unlist()
trawl_glorys_join_key <- trawls %>% 
  mutate(glorys_nn=nns) %>% 
  left_join(glorys_grd %>% st_set_geometry(NULL),by=c("glorys_nn"="grdid")) %>% 
  dplyr::select(station,glorys_nn)
```

```{r}
# plot_survey_comparison <- function(model_df){
#   spp=unique(model_df$spp)
#   ens_preds <- ensemble_predictions(model_df) %>% 
#     left_join(glorys_grd %>% st_set_geometry(NULL)) %>%
#     dplyr::select(-longitude,-latitude)
#   # join to trawl data
#   dat <- trawl_atlantis %>%
#     filter(species==spp) %>% 
#     mutate(year=year(date)) %>% 
#     left_join(trawl_glorys_join_key) %>% 
#     left_join(ens_preds,by=c('year','glorys_nn'='grdid')) %>% 
#     drop_na() %>% 
#     mutate(cpue_pred=exp(ens_est))
# }
```

# Calculate Proportional Abundance

Using Atlantis polygons, calculate proportional abundance of the species by Atlantis polygon and year

Match the GLORYS points to an Atlantis box

```{r}
glorys_atlantis_spatial_join <- glorys %>% 
  st_join(emocc_sf) %>% 
  st_set_geometry(NULL) %>% 
  distinct(longitude,latitude,box_id,label,area)
glimpse(glorys_atlantis_spatial_join)
```

# Join to Atlantis Polygons

Write a function to extract mean predictions for each Atlantis polygon, by first joining points to polygons using the join key we created above

```{r}
# the `preds` input here is the dataframe of ensemble predictions
ensemble_to_atlantis <- function(preds){
  
  # filter to the year of interest and exponentiate the estimated dependent variable
  df <- preds %>% 
    group_by(year,longitude,latitude) %>% 
    summarise(est=mean(ens_est,na.rm=T) %>% exp()) %>% 
    drop_na() %>% 
    ungroup()
  
  # join the SDM/Atlantis matching key
  atlantis_matched <- df %>% 
    left_join(glorys_atlantis_spatial_join,by=c('longitude','latitude')) %>% 
    drop_na() %>% 
    
    # group by Atlantis box and calculate a mean
    group_by(box_id,year) %>% 
    summarise(mean_est=mean(est,na.rm=T)) %>% 
    ungroup()
    
  # join the paired Atlantis boxes
  atlantis_mean_cpue_all_boxes <- atlantis_matched %>% 
    complete(box_id=full_seq(0:88,1),year,fill=list(mean_est=0)) %>%
    rename(est=mean_est) %>% 
    dplyr::select(box_id,year,est)
  
  # add in area information, scale up, then normalize across all boxes
  atlantis_est_final <- atlantis_mean_cpue_all_boxes %>% 
    left_join(emocc_box_areas,by=c('box_id')) %>% 
    group_by(year) %>% 
    # multiply by total area
    mutate(totest = est*area) %>% 
    # finally, calculate relative index
    mutate(rel_est = totest/sum(totest,na.rm=T)) %>% 
    mutate(perc_est = rel_est*100) %>% 
    ungroup()
  
  return(atlantis_est_final)
}
```

If you want to plot, can pick a year and join the spatial polygon data

```{r}
plot_sdm_year <- function(atlantis_estimates,sppname="",yr=2013){
  polys<- atlantis_estimates %>% 
    filter(year==yr) %>% 
    mutate(perc_est=na_if(perc_est,0)) %>% 
    left_join(emocc_sf,by=c('box_id','area')) %>% 
    st_as_sf()
  bbox=st_bbox(emocc_sf)
  
  t <- paste(sppname,yr)

  out <- ggplot()+
    geom_sf(data=coast,fill='gray50')+
    geom_sf(data=polys,aes(fill=perc_est),col=NA,alpha=1)+
    geom_sf(data=emocc_sf %>% summarise(),fill=NA,col='black')+
    scale_fill_viridis_c(na.value="white")+
    xlim(bbox[1],bbox[3])+ylim(bbox[2],bbox[4])+
    labs(title=paste("Percent of Total Abundance by Box,",t),fill="Percent")+
    theme(legend.position=c(0.7,0.7))
  
  plot(out)
}
```

Test with our example model

```{r,eval=F}
dover_to_atlantis <- ensemble_to_atlantis(dover_ens_preds)
plot_sdm_year(dover_to_atlantis,yr=2013)
```

```{r}
# write_rds(dover_to_atlantis,here('model output','atlantis sdms','dover sole.rds'))
```

# Combined Function

Write a function that wraps the above to go straight from a set of models to a dataframe of relative abundance in Atlantis polygons.

```{r}
sdm_to_atlantis <- function(model_df,use_offshore=F,save_rds=T){
  # make ensemble predictions
  n <- unique(model_df$spp)
  ens <- model_df %>% 
    ensemble_predictions(use_offshore=use_offshore)
  out <- ens %>% ensemble_to_atlantis()
  
  # save plots of 2013 CPUE and Atlantis proportions
  cpue2013_plot <- map_year(model_df,use_offshore=use_offshore)
  perc2013_plot <- plot_sdm_year(out,sppname=n,yr=2013)
  cpue2100_plot <- map_year(model_df,yr_vec=2100,use_offshore=use_offshore)
  perc2100_plot <- plot_sdm_year(out,sppname=n,yr=2100)
  
  ggsave(here('model output','atlantis sdms',paste0(n,'_cpue_2013.png')),cpue2013_plot,h=6,w=5)
  ggsave(here('model output','atlantis sdms',paste0(n,'_atlantis_2013.png')),perc2013_plot,h=6,w=5)
    
  ggsave(here('model output','atlantis sdms',paste0(n,'_cpue_2100.png')),cpue2100_plot,h=6,w=5)
  ggsave(here('model output','atlantis sdms',paste0(n,'_atlantis_2100.png')),perc2100_plot,h=6,w=5)
  
  if(save_rds){
    write_rds(out,here('model output','atlantis sdms',paste0(n,'.rds')))
  }
  return(out)
}
```

```{r,eval=F}
dover <- sdm_to_atlantis(dover_models)
```

# Apply to Atlantis Species

Source the sdmTMB modelling script, make a list of species to model, then try to fit ensemble models. For now, we do not include substrate as a covariate because GLORYS data does not have it.

```{r, message=F, echo=F, include=F}
rmarkdown::render(here::here('scripts','sdmTMB models.Rmd'),quiet=TRUE)
```

## Atlantis Demersal Fish Species

Some Atlantis groups are groups of species (e.g., "Shallow small rockfish"). Therefore, we will make a matching key and a modified version of the trawl survey data that combines these groups.

```{r}
atlantis_groups <- read_csv(here('data','atlantis_demersal_groups.csv'),col_types = 'ccc')
trawl_atlantis <- trawl_roms_utm %>% 
  left_join(atlantis_groups,by=c('species'='trawl_name')) %>% 
  filter(!is.na(code))
unique(trawl_atlantis$code)
unique(trawl_atlantis$group_name)

# groups by Atlantis code/group and summarize
trawl_atlantis <- trawl_atlantis %>% 
  group_by(date,trawl_id,station,depth_trawl,mean_temp_roms_30,mean_oxygen_roms_30,prop_hard_mixed,latitude,longitude,code,group_name) %>% 
  summarise(cpue_kg_km2=sum(cpue_kg_km2)) %>% 
  #rename so the models work
  rename(species=code) %>% 
  ungroup()
```


## Run Models

```{r,eval=F}
nms <- unique(trawl_atlantis$species)
# for all groups except FBP (lanternfish, deep vertical migrators), don't use offshore boxes for prediction
nms <- nms %>% str_subset("FBP",negate=T)

for(i in nms){
  tic(paste("Processing",i))
  model_df <- model_species(i,data=trawl_atlantis,use_substrate = F)
  atlantis_df <- sdm_to_atlantis(model_df,use_offshore = F)
  write_rds(atlantis_df,here('model output','atlantis sdms',paste0(i,'.rds')))
  toc()
}

# model FBP last
tic(paste("Processing FBP"))
model_df <- model_species("FBP",data=trawl_atlantis,use_substrate = F)
atlantis_df <- sdm_to_atlantis(model_df,use_offshore=T)
write_rds(atlantis_df,here('model output','atlantis sdms','FBP.rds'))
toc()
```

## Plots

```{r,fig.width=10,fig.height=10}
# plts<-purrr::map(nms,function(fl){
#   x <- read_rds(here('model output','atlantis sdms',paste0(fl,'.rds')))
#   plot_sdm_year(x,fl,yr=2013)
# })
# 
# library(cowplot)
# 
# plot_grid(plotlist=plts[1:4],nrow=2,ncol=2)
# plot_grid(plotlist=plts[5:8],nrow=2,ncol=2)
# plot_grid(plotlist=plts[9:12],nrow=2,ncol=2)
# plot_grid(plotlist=plts[13:16],nrow=2,ncol=2)
# plot_grid(plotlist=plts[17:20],nrow=2,ncol=2)
# plts[21]
```

```{r,eval=F}
# dover example
dover <- read_rds(here('model output','atlantis sdms','FDP.rds'))
dover2013 <- plot_sdm_year(dover,yr=2013)
dover2100 <- plot_sdm_year(dover,yr=2100)
```

