---
title: "Set up Fishing Footprints for Atlantis"
author: "Owen Liu"
date: '2022-09-09'
output: 
  html_document:
    toc: true
    toc_float: true
---

# Setup

```{r setup, include=FALSE}
library(tidyverse)
library(rbgm)
library(here)
library(rnaturalearth)
library(sf)
library(units)
select <- dplyr::select
options(dplyr.summarise.inform=F)
```

```{r}
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)

options(dplyr.summarise.inform = FALSE)
```

# Import Data

Groundfish fishing footprints by port

```{r} 
# import footprints
fp <- read_sf(here('data','atlantis','logbook_all_pvcs.shp'))
```

Tribal Usual and Accustomed Areas, with areas deeper than 700 fathoms masked out

```{r}
# tribal UAs
tribes <- read_sf(here('data','atlantis','tribal_U&As_with_700fm_cutoff.shp')) %>% 
  st_transform(st_crs(fp))
```

Generic fleet radii

```{r}
fleet_rad <- read_csv(here('data','atlantis','fisheries_radii.csv'))
```
Background and coastline map

```{r}
# load west cost land for mapping
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>%
  st_transform(st_crs(fp))
```

Atlantis .bgm

```{r}
# import Atlantis bgm as sf
# load the file
bgm <- read_bgm(here('data','grids','CalCurrentV3_utm.bgm'))
names(bgm)
atlantis_sf <- box_sf(bgm) %>% select(label,.bx0,box_id,poly_area=area)
st_crs(atlantis_sf) <- st_crs(attr(atlantis_sf$geometry, "crs")$proj)
# transform to same crs as the footprints
atlantis_sf <- atlantis_sf %>% st_transform(st_crs(fp))
```
Atlantis fleet definitions
```{r}
# fleets (with a manually-added set of columns to match to catch data)
fleets <- read_csv(here::here('data','atlantis','Atlantis_54fleets_definitions.csv'))
```


# Organize Footprints

## Get most recent groundfish footprints
```{r}
## filter footprints to just the most recent period (period 4, catch shares)
# unless the port does not have a footprints for period 4, then use the most recent one
fp4 <- fp %>%  
  group_by(port_name) %>% 
  slice_max(n=1,order_by = period) %>% 
  ungroup()
```


## Edit tribal areas

IMPORTANT UPDATE: MAKAH ARE THE ONLY TRIBE THAT USE TRAWL. 
So, tribal trawl will use the Makah footprint. Tribal non-trawl will use the combined (all-tribe U&As) footprints.

```{r}
# IMPORTANT UPDATE: MAKAH ARE THE ONLY TRIBE THAT USE TRAWL
# so, tribal trawl will use the Makah footprint
# tribal non-trawl will use the combined (all-tribe U&As) footprints
tribes_trawl_fp <- tribes %>% 
  filter(UA=="Makah")
tribes_nontrawl_fp <- tribes %>% 
  summarise() %>% 
  mutate(UA="all tribes")
tribes_trawl_nontrawl <- bind_rows(tribes_trawl_fp,tribes_nontrawl_fp) %>% 
  mutate(fleet=c("Tribal trawl","Tribal non-trawl"))
```

## Build Radii for Generic Fleets
For the generic Atlantis fleets, we use radii from a focal port to establish fishing areas. Here, we build the polygons with those radii.

```{r}
fleet_rad_polys <- fleet_rad %>% 
  filter(!is.na(km)) %>% 
  st_as_sf(coords=c("Longitude","Latitude"),crs=4326) %>% 
  # convert to planar crs
  st_transform(st_crs(fp)) %>% 
  # buffer using designated radii
  st_buffer(dist=.$km*1000) %>% 
  # mask out the coastline
  st_difference(coaststates %>% summarise())
```
## Visualize footprints pre-join

Tribal

```{r}
tbbox=st_bbox(tribes_trawl_nontrawl)
ggplot()+
  geom_sf(data=coaststates)+
  geom_sf(data=tribes_trawl_nontrawl,aes(fill=fleet,linetype=fleet))+
  coord_sf(xlim=c(tbbox[1],tbbox[3]),ylim=c(tbbox[2],tbbox[4]))+
  facet_wrap(~fleet)
```
Groundfish trawl
```{r,fig.width=6,fig.height=8}
bbox <- st_bbox(atlantis_sf)
ggplot()+
  geom_sf(data=coaststates)+
  geom_sf(data=atlantis_sf,fill='lightblue')+
  geom_sf(data=fp4,aes(fill=port_name))+
  labs(fill="IOPAC port")+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))
```
Generic fleets

```{r,fig.height=8,fig.width=6}
ggplot()+
  geom_sf(data=coaststates)+
  geom_sf(data=atlantis_sf,fill='lightblue')+
  geom_sf(data=fleet_rad_polys %>% filter(grepl("Coastal",Name)),aes(fill=Name),alpha=0.5)+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))
ggplot()+
  geom_sf(data=coaststates)+
  geom_sf(data=atlantis_sf,fill='lightblue')+
  geom_sf(data=fleet_rad_polys %>% filter(!grepl("Coastal",Name)),aes(fill=Name),alpha=0.5)+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))
```

# Bind footprints

Join these disparate polygon datasets and give them consistent names that match Atlantis fleet names

```{r}
# fix groundfish trawl footprint fleet names
fp_trawl <- fp4 %>% 
  left_join(fleets,by=c("port_name"="iopac")) %>% 
  filter(fishery_type=="bottom trawl") %>% 
  select(Code,Index,Name)
```

```{r}
# fix tribal fleet names
fp_tribes <- tribes_trawl_nontrawl %>% 
  # just do this manually
  mutate(Index=c(51,52),Code=c("TribalTrawl","TribalNonTrawl"),Name=c("TribalTrawl","TribalNonTrawl")) %>% 
  # mutate(area=st_area(.))%>% 
  select(Code,Index,Name)
```

```{r}
# fix generic fleet names
fp_generic <- fleet_rad_polys %>% 
  select(Code,Index,Name)
```

```{r}
# join
fp_all <- bind_rows(fp_trawl,fp_tribes,fp_generic) %>% 
  arrange(Index)
```

# Spatial Overlay on Atlantis Polygons

Groundfish footprints

```{r}
fp_atlantis_join <- fp_all %>% 
  st_intersection(atlantis_sf) %>% 
  # calculate area of each individual footprint/polygon overlap
  mutate(area_overlap=st_area(.)) %>% 
  # then, calculate each of these areas as a proportion of total polygon area
  mutate(prop_poly=area_overlap/poly_area) %>% 
  mutate(prop_poly=as.numeric(prop_poly)) %>% 
  mutate(code_name=paste0("MPA",Code," 89"))

# arrange final output as data frames (port by polygon by proportion)
fp_prop_polys<- fp_atlantis_join %>% 
  mutate(prop_poly=round(prop_poly,digits=6)) %>% 
  st_set_geometry(NULL) %>%
  select(Index,code_name,box_id,prop_poly) %>% 
  #fill in zeroes for other, non-represented boxes
  complete(nesting(Index,code_name),box_id=full_seq(c(0,88),1),fill=list(prop_poly=0)) %>% 
  arrange(Index,code_name,box_id) %>% 
  dplyr::select(code_name,box_id,prop_poly) %>% 
  pivot_wider(names_from="box_id",values_from = "prop_poly")
```

# Visualize Results

Example overlays of footprint polygons and Atlantis polygons

```{r}
fp_atlantis_join %>% 
  filter(code_name=="11_AstoriaGroundfishTrawl") %>% 
  ggplot()+geom_sf(aes(fill=prop_poly))+
  labs(fill="Proportion of\nAtlantis Polygon")

fp_atlantis_join %>% 
  filter(code_name=="52_TribalNonTrawl") %>% 
  ggplot()+geom_sf(aes(fill=prop_poly))+
  labs(fill="Proportion of\nAtlantis Polygon")
```

# Write output

```{r}
options(scipen = 100)
# options(scipen = 0)
# write!
write.table(fp_prop_polys,here::here('data','atlantis','mpas_by_fleet.txt'),quote=F,row.names=F,col.names=F)

write_sf(fleet_rad_polys,here::here('data','atlantis','cps_footprints.shp'))
```

