---
title: "SDMs and Climate Novelty"
author: "Owen Liu"
date: '2022-08-17'
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(cowplot)

knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform=FALSE)
```

```{r}
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=10),
        legend.title = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=10,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

# Purpose

Compare outputs from projected species distribution models to measures of climate novelty calculated using hypervolumes. In short, we ask: in the future, where will projected species distributions overlap with novel climates (where, in this case, novel climate means a novel combination of bottom oxygen and bottom temperature encountered in a given spatial cell).

# Load Novelty Data

We have measures of projected novelty for each year from 2011-2100, relative to a temperature/oxygen hypervolume created for the period 1980-2010. The novelty calculation was performed separately for the Hadley, GFDL, and IPSL downscaled ROMS models for the California Current (see Pozo-Buil et al. 2021)

```{r}
novelty <- read_rds(here('model output','climate novelty','novelty_inclusion_3esms.rds'))
```

```{r}
# get area of roms cells
# roms_fls <- list.files(here('data','hypervolumes','projection rasters'),full.names=T)
# roms_area <- terra::cellSize(rast(roms_fls[[1]]),unit='km') %>% 
#   as.data.frame(cells=T)
# write_rds(roms_area,here('data','roms_cell_area_km2.rds'))
roms_area <- read_rds(here('data','roms_cell_area_km2.rds'))
```

And here is some example projected species data for comparison

```{r}
dogfish <- read_rds(here('model output','pacific spiny dogfish','projection_3ESMS.rds'))
```

Take a look at the data
```{r}
glimpse(novelty)
glimpse(dogfish)
```

# Join Function

Write a function to join together the novelty data and the species distribution projections, while keeping track of the 3 ROMS models

```{r}
join_novelty_sdm <- function(novelty_df,sdm_df){
  sdm_df <- sdm_df %>% 
    dplyr::select(-est_non_rf,est_rf,epsilon_st)
  novelty_long <- novelty_df %>% 
    pivot_longer(contains('novel'),names_to='esm',values_to='is_novel') %>% 
    mutate(esm=str_replace(esm,'novel_',''))
  novelty_sdm_join <- sdm_df %>% 
    left_join(novelty_long,by=c('lon','lat','year','esm')) %>% 
    filter(year>2010)
  novelty_sdm_join
}
```

```{r}
dogfish_novelty <- join_novelty_sdm(novelty,dogfish)
glimpse(dogfish_novelty)
```

# Summary Function

We can write another function to summarize the data by year, calculating what percent of the species' distribution (e.g., number of cells or proportion of total CPUE) is projected to occur in novel climates.

```{r}
summarise_sdm_novelty <- function(sdm_novelty_df){
  sdm_novelty_summ <- sdm_novelty_df %>% 
    # bring in area of ROMS cells to calculate biomass
    left_join(roms_area,by=c('roms_cell'='cell')) %>% 
    mutate(biomass_kg=exp(est)*area) %>% 
    # then summarise by year and ROMS model
    group_by(esm,year) %>% 
    summarise(biomass_novel=sum(biomass_kg[is_novel]),
              total_biomass=sum(biomass_kg),
              novel_cells=sum(is_novel),
              total_cells=n()) %>% 
    mutate(prop_novel_biomass=biomass_novel/total_biomass,
           prop_novel_cells=novel_cells/total_cells) %>% 
    ungroup()
  
  sdm_novelty_summ
}
```

```{r}
dogfish_novelty_summ <- summarise_sdm_novelty(dogfish_novelty)
```

# Apply

Apply to all selected, projected species' models

```{r}
fps <- list.files(here('Janelle','selected models'),recursive = T,full.names = T) %>% 
  str_subset("projection_3ESMs.rds")
head(fps)
# make sure this is in the right order
spp_names <- c("big skate","longnose skate","pacific spiny dogfish","spotted ratfish","arrowtooth flounder","curlfin sole","dover sole","english sole","pacific sanddab","petrale sole","rex sole","slender sole","lingcod","longspine thornyhead","pacific grendadier","sablefish","shortspine thornyhead","aurora rockfish","bocaccio","canary rockfish","chilipepper","darkblotched rockfish","greenstriped rockfish","pacific ocean perch","sharpchin rockfish","shortbelly rockfish","shortraker rockfish","splitnose rockfish","stripetail rockfish","widow rockfish","yelloweye rockfish","yellowtail rockfish")

# do the calculations
sdm_novelty <- purrr::map2_df(spp_names,fps,function(x,y){
  df <- read_rds(y)
  out <- join_novelty_sdm(novelty_df = novelty,sdm_df=df) %>% 
    summarise_sdm_novelty() %>% 
    mutate(species=x)
  out
})
spp_cats <- tibble(species=spp_names,type=c(rep("elasmobranch",4),rep("flatfish",8),rep("other",5),rep("rockfish",15)))
sdm_novelty <- sdm_novelty %>% 
  ungroup() %>% 
  left_join(spp_cats,by="species")
```

# Visualize

```{r}
# example for dogfish- total biomass in novel climates
dogfish_novelty_biomass_novel_ts <- dogfish_novelty_summ %>% 
  ggplot(aes(year,biomass_novel/1000,color=esm))+
  geom_line(size=1.5)+
  labs(x="Year",y="Biomass in Novel Climate (MT)",color="Model")

# proportion of biomass in novel climates
dogfish_novelty_propnovel_ts <- dogfish_novelty_summ %>% 
  ggplot(aes(year,prop_novel_biomass,color=esm))+
  geom_line(size=1.5)+
  ylim(0,1)+
  labs(x="Year",y="Proportion of Biomass in Novel Climate",color="Model")

dogfish_novelty_biomass_novel_ts
dogfish_novelty_propnovel_ts
```

Summary figures across species

```{r,fig.width=8,fig.height=10}
hov_prop_bio_elasmo <- sdm_novelty %>% 
  filter(year>2020,type=="elasmobranch") %>% 
  ggplot(aes(year,species,fill=prop_novel_biomass),col=NA)+
  geom_tile()+
  scale_fill_gradient(low="#440154FF",high="#FDE725FF",limits=c(0,1),guide="none")+
  labs(fill="Proportion of\nBiomass",x="",title="")+
  facet_wrap(~esm)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle=45,hjust=0.5))+
  theme(plot.margin = unit(c(0, 0.2, 0, 0), "cm"))
hov_prop_bio_flatfish <- sdm_novelty %>% 
  filter(year>2020,type=="flatfish") %>% 
  ggplot(aes(year,species,fill=prop_novel_biomass),col=NA)+
  geom_tile()+
  scale_fill_gradient(low="#440154FF",high="#FDE725FF",limits=c(0,1),guide="none")+
  labs(fill="Proportion of\nBiomass",x="",title="")+
  facet_wrap(~esm)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle=45,hjust=0.5))+
  theme(plot.margin = unit(c(0, 0.2, 0, 0), "cm"))
hov_prop_bio_other <- sdm_novelty %>% 
  filter(year>2020,type=="other") %>% 
  ggplot(aes(year,species,fill=prop_novel_biomass),col=NA)+
  geom_tile()+
  scale_fill_gradient(low="#440154FF",high="#FDE725FF",limits=c(0,1),guide="none")+
  labs(fill="Proportion of\nBiomass",x="",title="")+
  facet_wrap(~esm)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle=45,hjust=0.5))+
  theme(plot.margin = unit(c(0, 0.2, 0, 0), "cm"))
hov_prop_bio_rockfish <- sdm_novelty %>% 
  filter(year>2020,type=="rockfish") %>% 
  ggplot(aes(year,species,fill=prop_novel_biomass),col=NA)+
  geom_tile()+
  scale_fill_gradient(low="#440154FF",high="#FDE725FF",limits=c(0,1))+
  labs(fill="Proportion of\nBiomass",x="",title="")+
  facet_wrap(~esm)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle=45,hjust=0.5),
        legend.position = 'bottom',
        legend.key.width = unit(2,'cm'))+
  theme(plot.margin = unit(c(0, 0.2, 0, 0), "cm"))
hov_prop_bio <- plot_grid(hov_prop_bio_elasmo,NULL,hov_prop_bio_other,NULL,hov_prop_bio_flatfish,NULL,hov_prop_bio_rockfish,nrow=7,rel_heights = c(4,-1,5,-1,8,-1,14))
hov_prop_bio
```

Mean rate of habitat "loss" by species

```{r}
calc_hab_loss <- function(df){
  m <- lm(prop_novel_biomass~year,data=df)
  coef(m)['year']
}
sdm_habloss <- sdm_novelty %>% 
  ungroup() %>% 
  group_by(species,type,esm) %>% 
  nest() %>% 
  mutate(yeareff=purrr::map_dbl(data,calc_hab_loss)) %>% 
  mutate(perc_loss_yr=yeareff*100) %>% 
  dplyr::select(species,type,esm,perc_loss_yr) %>% 
  ungroup()
```

```{r,fig.width=8}
habloss_plot <- sdm_habloss %>% 
  mutate(spp_ordered=fct_reorder(species,perc_loss_yr)) %>% 
  ggplot(aes(spp_ordered,perc_loss_yr,fill=type))+
  geom_col()+
  coord_flip()+
  geom_hline(yintercept = 0,color='black')+
  scale_fill_manual(values=pnw_palette("Bay",4))+
  facet_wrap(~esm,nrow=1)+
  labs(y=expression(paste(Delta,"Novel Biomass per Year (%)")),x="Species",fill="Group")+
  theme(panel.border = element_rect(color='black',fill=NA))
habloss_plot
```


